### 2018-05-08 article/ms.tex
article frame

\documentclass[twocolumn]{aastex61}
\usepackage{bm}

\newcommand\teff{T_{\rm eff}}
\newcommand\logg{\log{g}}
\newcommand\feh{[\rm{Fe}/\rm{H}]}
\newcommand\mh{[\rm{M}/\rm{H}]}

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\Gaia}{\project{Gaia}}
\newcommand{\gaia}{\project{gaia}}

\received{2018 XX XX}
\revised{2018 XX XX}
\accepted{2018 XX XX}

%\submitjournal{AAS Journals}

\shorttitle{Short title}
\shortauthors{Casey et al.}

\begin{document}

\title{Binarity and other things in Gaia}

\correspondingauthor{Andrew R. Casey}
\email{andrew.casey@monash.edu}

\author[0000-0003-0174-0564]{Andrew R. Casey}
\affiliation{School of Physics \& Astronomy, Monash University, Clayton 3800, 
    Victoria, Australia}
\affiliation{Faculty of Information Technology, Monash University, Clayton 3800, 
    Victoria, Australia}

\author[0000-0003-2866-9403]{David W. Hogg}
\affiliation{Center for Cosmology and Particle Physics, Department of Physics,
    New York University, 726 Broadway, New York, NY 10003, USA}
\affiliation{Center for Data Science, New York University, 60 Fifth Ave, 
    New York, NY 10011, USA}
\affiliation{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 
    Heidelberg}
\affiliation{Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}

\begin{abstract}
Velociraptors can't see you if you don't move. (\Gaia can see you if you can.)
\end{abstract}

\keywords{stars: dynamics}

\section{Introduction} \label{sec:intro}

\begin{itemize}
\item the binary fraciton is important
\item Gaia provides exquisite astrometry and RV measurements over many years
\item The errors in these quantities are well-characterised
\item When a star is in a binary system (or triary, etc), there will be excess scatter relative to a single star of the primary's properties
\item the excess RV noise (and in DR1 or DR3, excess astrometric noise) can be used to infer the presence of a companion
\end{itemize}

\section{Model}

% What is provided by Gaia?
% astrometric excess, number of measurements, brightness, etc
% what is the error model that Gaia uses?
% WE COULD USE A HIErARCHICAL MODEL FOR THE ERRORS
% Thus for a single star we would expect, so we can calculate
% p(\ra, \dec|data)

% For stars with a companion, if the companion is large enough to induce an effect on the primary then this would produce astrometric noise that would  be observable in the

% Astrometric model:

% Can we do this for DR1 already?

% When would we expect to see astrometric noise and not RV noise, and vice-versa? in what situations?

% we can calculate probability that it is consistent with a single  star,... p-value?

% From the Gaia expected error distributions: how many stars might we be able to discriminate between?

\section{Results}

\section{Discussion}

\software{numpy, scipy, astropy, Stan}

\acknowledgements
- Gaia/DPAC
- Gaia archive?
- Gaia DR2 hack week at Flatiron

\end{document}

### 2018-05-08 .ci/upload.sh

@@ -1,0 +1,11 @@
+#!/bin/bash -x
+
+# Push to GitHub
+if [ -n "$GITHUB_API_KEY" ]; then
+  cd $TRAVIS_BUILD_DIR
+  git checkout --orphan pdf
+  git rm -rf .
+  git add -f ms.pdf
+  git -c user.name='travis' -c user.email='travis' commit -m "building the paper"
+  git push -q -f https://andycasey:$GITHUB_API_KEY@github.com/andycasey/velociraptor pdf
+fi

### 2018-05-08 data/download.py
things

@@ -10,7 +10,7 @@ columns = (
     "source_id",
     "ra", "dec", "l", "b",
     "parallax", "parallax_error",
-    "phot_g_mean_mag",
+    "phot_g_mean_mag", "phot_bp_mean_mag", "phot_rp_mean_mag",
     "phot_g_mean_flux", "phot_g_mean_flux_error",
     "phot_bp_mean_flux", "phot_bp_mean_flux_error", "phot_bp_n_obs",
     "phot_rp_mean_flux", "phot_rp_mean_flux_error", "phot_rp_n_obs",
@@ -20,7 +20,8 @@ columns = (
     "a_g_val", "a_g_percentile_lower", "a_g_percentile_upper",
     "radial_velocity", "radial_velocity_error", "rv_nb_transits",
     "rv_template_teff", "rv_template_logg", "rv_template_fe_h",
-    "astrometric_weight_al", "astrometric_gof_al", "astrometric_chi2_al"
+    "astrometric_weight_al", "astrometric_gof_al", "astrometric_chi2_al",
+    "phot_variable_flag"
 )
 
 job = Gaia.launch_job_async("""
@@ -34,7 +35,7 @@ job = Gaia.launch_job_async("""
     AND     MOD(random_index, 10) = 0
     """.format(", ".join(columns)))
 subset = job.get_results()
-subset.write("rv_floor_cal_subset-result.fits")
+subset.write("rv_calibration_floor_subset-result.fits")
 
 
 job = Gaia.launch_job_async("""
@@ -47,4 +48,4 @@ job = Gaia.launch_job_async("""
     AND     radial_velocity_error < 20
     """.format(", ".join(columns)))
 sources = job.get_results()
-sources.write("rv_floor_cal-result.fits")
+sources.write("rv_calibration_floor-result.fits")

### 2018-05-08 model.py

@@ -30,7 +30,7 @@ def _savefig(fig, basename, dry_run=True, **kwargs):
 
 #gaia_sources = Table.read("data/rv_floor_cal-result.fits")
 
-subset = Table.read("data/rv_floor_cal_subset-result.fits.gz")
+subset = Table.read("data/rv_calibration_floor_subset-result.fits.gz")
 
 # Take a random subset of stars.
 subset_indices = np.random.choice(len(subset), size=10000, replace=False)
@@ -131,22 +131,31 @@ data = dict(N=N, rv_variance=rv_variance[qc], flux=rp_flux[qc])
 
 
 def get_mu_design_matrix(subset):
-    return np.array([
+    dm = np.array([
         np.ones(len(subset)),
         subset["phot_rp_mean_flux"]**-1,
         subset["phot_rp_mean_flux"]**-2,
     ])
+    #(subset["phot_bp_mean_flux"] - subset["phot_rp_mean_flux"])**-1,
+    #(subset["phot_bp_mean_flux"] - subset["phot_rp_mean_flux"])**-2,
+    initial = np.array([1e-2, 1e5, 1e10])# 1e-2, 1e-2])
+    return (dm, initial)
+
 
 def get_sigma_design_matrix(subset):
-    return np.array([
+    dm = np.array([
         np.ones(len(subset)),
         subset["phot_rp_mean_flux"]**-1,
         subset["phot_rp_mean_flux"]**-2,
     ])
+    #    (subset["phot_bp_mean_flux"] - subset["phot_rp_mean_flux"])**-1,
+    #    (subset["phot_bp_mean_flux"] - subset["phot_rp_mean_flux"])**-2,
+    initial = np.array([1e-2, 1e5, 1e10])# 1e-2, 1e-2])
+    return (dm, initial)
 
 
-mu_design_matrix = get_mu_design_matrix(subset[qc])
-sigma_design_matrix = get_sigma_design_matrix(subset[qc])
+mu_design_matrix, initial_mu_coefficients = get_mu_design_matrix(subset[qc])
+sigma_design_matrix, initial_sigma_coefficients = get_sigma_design_matrix(subset[qc])
 
 
 data.update(
@@ -155,9 +164,8 @@ data.update(
     M=mu_design_matrix.shape[0],
     S=sigma_design_matrix.shape[0])
 
-init = dict(outlier_fraction=0.1,
[…]

### 2018-05-08 model.py
stoopid stan model

@@ -16,7 +16,7 @@ plt.style.use(mpl_style)
 
 np.random.seed(42)
 
-def _savefig(fig, basename, dry_run=True, **kwargs):
+def _savefig(fig, basename, dry_run=False, **kwargs):
     kwds = dict(dpi=300)
     kwds.update(kwargs)
     if not dry_run:
@@ -164,7 +164,7 @@ data.update(
     M=mu_design_matrix.shape[0],
     S=sigma_design_matrix.shape[0])
 
-init = dict(outlier_fraction=0.1, mu_coefficients=initial_mu_coefficients,
+init = dict(theta=0.1, mu_coefficients=initial_mu_coefficients,
     sigma_coefficients=initial_sigma_coefficients)
 
 

### 2018-05-08 model.py
:boom:

@@ -329,9 +329,9 @@ for label_name, description, is_log in shorthand_parameters:
         right_edge = equidensity_bins[i + 1]
 
         in_bin = (right_edge > subset_x) * (subset_x >= left_edge)
-        equidensity_f_rv_variance[i] += np.sum(has_excess_rv_variance * in_bin) \
+        equidensity_f_rv_variance[i] += np.sum(subset_has_excess_rv_variance * in_bin) \
                                       / np.sum(in_bin)
-        equidensity_f_rv_variance_error[i] += np.sqrt(np.sum(has_excess_rv_variance * in_bin)) \
+        equidensity_f_rv_variance_error[i] += np.sqrt(np.sum(subset_has_excess_rv_variance * in_bin)) \
                                             / np.sum(in_bin)
 
 
@@ -378,8 +378,8 @@ def plot_model_residuals(samples, subset, label_name, semilogx=False,
     mean_sigma_coefficients = np.mean(
         samples.extract(("mu_coefficients", ))["mu_coefficients"], axis=0)
 
-    mu_design_matrix = get_mu_design_matrix(subset)
-    sigma_design_matrix = get_sigma_design_matrix(subset)
+    mu_design_matrix, _ = get_mu_design_matrix(subset)
+    sigma_design_matrix, _ = get_sigma_design_matrix(subset)
 
     subset_mu = np.dot(mean_mu_coefficients, mu_design_matrix)
     subset_sigma = np.dot(mean_sigma_coefficients, sigma_design_matrix)
@@ -423,8 +423,8 @@ def plot_excess_rv_variance(samples, subset, label_name, semilogx=False,
     mean_sigma_coefficients = np.mean(
         samples.extract(("mu_coefficients", ))["mu_coefficients"], axis=0)
 
-    mu_design_matrix = get_mu_design_matrix(subset)
-    sigma_design_matrix = get_sigma_design_matrix(subset)
+    mu_design_matrix, _ = get_mu_design_matrix(subset)
+    sigma_design_matrix, _ = get_sigma_design_matrix(subset)
 
     subset_mu = np.dot(mean_mu_coefficients, mu_design_matrix)
     subset_sigma = np.dot(mean_sigma_coefficients, sigma_design_matrix)
[…]

### 2018-05-08 model.py
dot points

@@ -537,10 +537,9 @@ H_bin, _, __ = np.histogram2d(
     bins=(xedges, yedges))
 
 H = H_bin/H_all
-#H[~np.isfinite(H)] = 0.0
-H[(H == 1) * (H_all < 2)] = np.nan
 
-image = axes[1].imshow(H.T, extent=(xedges[0], xedges[-1], yedges[-1], yedges[0]),
+image = axes[1].imshow(H.T, 
+    extent=(xedges[0], xedges[-1], yedges[-1], yedges[0]),
     aspect=np.ptp(xedges)/np.ptp(yedges), cmap="viridis")
 
 
@@ -560,3 +559,17 @@ axes[1].xaxis.set_major_locator(MaxNLocator(6))
 axes[1].yaxis.set_major_locator(MaxNLocator(6))
 
 _savefig(fig, "hrd-binary-fraction")
+
+
+# Nect steps:
+# [ ] decide what to plot as "residual"
+# [ ] plot "residual" vs other stellar parameters and decide if they need to be
+#     accounted for when fitting:
+#       [ ] absolute luminosity / stellar luminosity
+#       [ ] bp flux - rp flux
+#       [ ] which RV template was used
+#       [ ] RV template properties minus a guess at the stellar properties
+#       [ ] 
+    
+
+

### 2018-05-08 article/ms.tex
dot points
actual dot points

@@
 
-\shorttitle{Short title}
-\shortauthors{Casey et al.}
+\shorttitle{Stellar multiplicity}
+\shortauthors{Casey \& Hogg}
 
@@
 
-\title{Binarity and other things in Gaia}
+\title{Stellar multiplicity}
 
@@
 
-% What is provided by Gaia?
-% astrometric excess, number of measurements, brightness, etc
-% what is the error model that Gaia uses?
-% WE COULD USE A HIErARCHICAL MODEL FOR THE ERRORS
-% Thus for a single star we would expect, so we can calculate
-% p(\ra, \dec|data)
+\begin{itemize}
+\item The effect of binaries.
 
-% For stars with a companion, if the companion is large enough to induce an effect on the primary then this would produce astrometric noise that would  be observable in the
+\item RV noise vs astrometric excess noise.
 
-% Astrometric model:
+\item When would we expect to see astrometric noise and not RV noise, and vice-versa? in what situations?
 
-% Can we do this for DR1 already?
+\item Here we base our model using RV data only from Gaia DR2.
 
-% When would we expect to see astrometric noise and not RV noise, and vice-versa? in what situations?
+\item Some fraction of stars in Gaia are single star systems (or have such low semi-amplitudes that would not be detected by Gaia), and we will use those stars (within a mixture model) to calibrate the radial velocity error floor
 
-% we can calculate probability that it is consistent with a single  star,... p-value?
+\item We will do this on individual epoch measurements
 
-% From the Gaia expected error distributions: how many stars might we be able to discriminate between?
+\item ASSUMPTIONS
+
+\item MODEL
+
+\item CALIBRATION SAMPLE (including ADQL query)
+
+\item Fitting the model:
+
+\item Comparisons of residuals with other stellar properties.
+
+\item What is the RV error that Gaia can measure in a single epoch for a given magnitude vs colour, etc.
+
+\item What are the distributions of orbital parameters of binary systems that we would be able to detect?
+
+\item Calculating a probability of binarity for every star, taking the non-measurements into account!
+
+\item Completeness cuts,.... selection effects?
+
+\end{itemize}
 
 \section{Results}
+
+\begin{itemize}
+
+\item Binary fraction in the spaces that we were fitting: rp flux, colour, etc, just showing the transition of  probabilities
+
+\item Binary fraction as a function of fitting properties, and other stellar properties from Gaia
+
+\item Binarity across the H-R diagram
+
+\item Binarity with metallicity
+
+\item Binarity at height above the plane?
+
+\item Binarity in clusters vs the plane?
+\end{itemize}
 

### 2018-05-09 modeling.py
wip

@@ -1,0 +1,53 @@
+
+"""
+Modeling and validation of the radial velocity calibration model.
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+import velociraptor
+
+# Load the data and make some plots.
+sources = velociraptor.load_gaia_sources("data/rv-cal-subset.fits.gz", N=1000)
+
+model, data_dict, init_dict, used_in_fit = velociraptor.prepare_model(
+    sources["phot_rp_mean_flux"], sources["rv_single_epoch_variance"])
+
+samples = model.sampling(**velociraptor.stan.sampling_kwds(
+    data=data_dict, chains=2, iter=2000, init=init_dict))
+
+
+
+
+
+scatter_kwds = dict(s=1, facecolor="#000000", alpha=0.5, rasterized=True)
+
+
+fig, ax = plt.subplots()
+
+indices = np.argsort(sources["phot_rp_mean_flux"])
+x = sources["phot_rp_mean_flux"][indices]
+mu, sigma = velociraptor.predict_map_rv_single_epoch_variance(samples, x)
+
+ax.scatter(
+    sources["phot_rp_mean_flux"], sources["rv_single_epoch_variance"],
+    **scatter_kwds)
+ax.plot(x, mu, "r-")
+ax.fill_between(x, mu - sigma, mu + sigma, 
+    facecolor="r", alpha=0.3, edgecolor="none")
+ax.set_ylim(-0.5, 10)
+ax.semilogx()
+
+
+
+fig, ax = plt.subplots()
+
+ax.scatter(
+    x, sources["rv_single_epoch_variance"] - expectation,
+    **scatter_kwds)
+
+ax.semilogx()
+
+
+

### 2018-05-09 mpl_style.py

@@ -2,6 +2,8 @@
 # A matplotlib style based on the gala package by @adrn:
 # github.com/adrn/gala
 
+import numpy as np
+
 mpl_style = {
 
     # Lines
@@ -68,3 +70,21 @@ mpl_style = {
     # Other
     'savefig.dpi': 300,
 }
+
+
+
+
+def plot_histogram_steps(ax, x_bins, y, y_err):
+
+    xx = np.array(x_bins).repeat(2)[1:]
+    xstep = np.repeat((x_bins[1:] - x_bins[:-1]), 2)
+    xstep = np.concatenate(([xstep[0]], xstep, [xstep[-1]]))
+    # Now: add one step at end of row.
+    xx = np.append(xx, xx.max() + xstep[-1]) - xstep/2.0
+
+    yy = np.array(y).repeat(2)
+
+    _ = ax.plot(xx, yy, '-')
+    ax.errorbar(x_bins, y, y_err, fmt=None, capsize=0, ecolor=_[0].get_color())
+
+    return (xx, yy)

### 2018-05-09 velociraptor.py

@@ -1,0 +1,118 @@
+
+"""
+Code for the velociraptor project.
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+from astropy.table import Table
+
+import stan_utils as stan
+from mpl_style import mpl_style
+
+plt.style.use(mpl_style)
+
+np.random.seed(42)
+
+def load_gaia_sources(path, N=None, **kwargs):
+    """
+    Load a subset of Gaia data and calculate additional properties like
+    `absolute_g_mag`, `absolute_bp_mag`, `absolute_rp_mag`, and
+    `rv_single_epoch_variance`.
+
+    :param path:
+        The local path to load the sources from.
+
+    :param N: [optional]
+        If not `None`, draw a random `N` sources from the data.
+    """
+
+    sources = Table.read(path, **kwargs)
+
+    for band in ("g", "bp", "rp"):
+        sources["absolute_{}_mag".format(band)] = \
+              sources["phot_{}_mean_mag".format(band)] \
+            + 5 * np.log10(sources["parallax"]/100.0)
+
+    sources["rv_single_epoch_variance"] = sources["radial_velocity_error"]**2 \
+                                        * sources["rv_nb_transits"] * np.pi/2.0
+
+    if N is not None:
+        S = len(sources)
+        return sources[np.random.choice(S, size=min(N, S), replace=False)]
+    return sources
+
+
+def prepare_model(phot_rp_mean_flux, rv_single_epoch_variance, 
+    model_path="model.stan"):
+    """
+    Compile the Stan model, and prepare the data and initialisation dictionaries
+    for optimization and sampling.
+
+    :param phot_rp_mean_flux:
+        The mean RP flux.
+
+    :param rv_single_epoch_variance:
+        The variance in single epoch measurements of radial velocity.
+
+    :param model_path: [optional]
+        The local path of the Stan model.
+
+    :returns:
+        The compiled Stan model, the data dictionary, the initialsiation
+        dictionary, and a mask corresponding to which `sources` were used to
+        construct the data dictionary.
+    """
+
+    dm = _rvf_design_matrix(phot_rp_mean_flux)
+    finite = np.all(np.isfinite(dm), axis=0)
+
+    if not all(finite):
+        print("Design matrix contains non-finite values! Masking out.")
+
+    dm = dm[:, finite]
+    coeff = _rvf_initial_coefficients(dm)
[…]

### 2018-05-09 modeling.py
updates and refactor

@@ -9,45 +9,174 @@ import matplotlib.pyplot as plt
 import velociraptor
 
 # Load the data and make some plots.
-sources = velociraptor.load_gaia_sources("data/rv-cal-subset.fits.gz", N=1000)
+sources = velociraptor.load_gaia_sources("data/rv-cal-subset.fits.gz", N=1e3)
 
-model, data_dict, init_dict, used_in_fit = velociraptor.prepare_model(
-    sources["phot_rp_mean_flux"], sources["rv_single_epoch_variance"])
+model, data_dict, init_dict, used_in_fit = velociraptor.prepare_model(**sources)
 
 samples = model.sampling(**velociraptor.stan.sampling_kwds(
     data=data_dict, chains=2, iter=2000, init=init_dict))
 
 
+velociraptor.plot_model_predictions_corner(samples, sources,
+    log_parameters=("phot_rp_mean_flux", ),
+    labels=dict(phot_rp_mean_flux=r"\textrm{mean rp flux}"))
+
+raise a
 
 
 
 scatter_kwds = dict(s=1, facecolor="#000000", alpha=0.5, rasterized=True)
+model_scatter_kwds = dict(s=1, facecolor="r", alpha=0.5, rasterized=True)
 
+
+model_rv_sev_mu, model_rv_sev_sigma = \
+    velociraptor.predict_map_rv_single_epoch_variance(
+        samples, phot_rp_mean_flux=sources["phot_rp_mean_flux"])
+
+# Plot the model against various properties of interest.
+shorthand_parameters = [
+    ("phot_rp_mean_flux", "mean rp flux", True),
+    ("phot_bp_mean_flux", "mean bp flux", True),
+    ("phot_g_mean_flux", "mean g flux", True),
+    ("bp_rp", "bp-rp colour", False),
+    ("phot_g_mean_mag", "mean g magnitude", False),
+    ("phot_rp_mean_mag", "mean rp magnitude", False),
+    ("phot_bp_mean_mag", "mean bp magnitude", False),
+    ("teff_val", "inferred temperature", False),
+    ("ra", "right ascension", False),
+    ("dec", "declination", False),
+    ("radial_velocity", "radial velocity", False),
[…]

### 2018-05-09 velociraptor.py

@@ -3,9 +3,11 @@
 Code for the velociraptor project.
 """
 
+import logging
 import numpy as np
 import matplotlib.pyplot as plt
 from astropy.table import Table
+from matplotlib.ticker import MaxNLocator
 
 import stan_utils as stan
 from mpl_style import mpl_style
@@ -39,24 +41,32 @@ def load_gaia_sources(path, N=None, **kwargs):
 
     if N is not None:
         S = len(sources)
-        return sources[np.random.choice(S, size=min(N, S), replace=False)]
+        return sources[np.random.choice(S, size=min(int(N), S), replace=False)]
     return sources
 
 
-def prepare_model(phot_rp_mean_flux, rv_single_epoch_variance, 
-    model_path="model.stan"):
+def prepare_model(rv_single_epoch_variance, target=0.01, model_path="model.stan",
+    **source_params):
     """
     Compile the Stan model, and prepare the data and initialisation dictionaries
     for optimization and sampling.
 
-    :param phot_rp_mean_flux:
-        The mean RP flux.
-
     :param rv_single_epoch_variance:
         The variance in single epoch measurements of radial velocity.
 
+    :param target: [optional]
+        The target radial velocity variance to use when initialising the model
+        coefficients.
+
     :param model_path: [optional]
         The local path of the Stan model.
+
+    :Keyword Arguments:
+        * *source_params* (``dict``) These are passed directly to the
+        `_rvf_design_matrix`, so they should include all of the source labels
+        needed to construct the design matrix (e.g., `phot_rp_mean_flux`). The
+        array length for each source label should match that of
+        `rv_single_epoch_variance`.
 
     :returns:
         The compiled Stan model, the data dictionary, the initialsiation
@@ -64,17 +74,19 @@ def prepare_model(phot_rp_mean_flux, rv_single_epoch_variance,
         construct the data dictionary.
     """
 
-    dm = _rvf_design_matrix(phot_rp_mean_flux)
+    dm = _rvf_design_matrix(**source_params)
     finite = np.all(np.isfinite(dm), axis=0)
 
     if not all(finite):
-        print("Design matrix contains non-finite values! Masking out.")
-
+        logging.warn("Design matrix contains {0} (of {1}) non-finite values!"\
+                     "Exlcuding them from model fit.".format(
[…]

### 2018-05-09 modeling.py
correct, except for axes

@@ -18,8 +18,9 @@ samples = model.sampling(**velociraptor.stan.sampling_kwds(
 
 
 velociraptor.plot_model_predictions_corner(samples, sources,
-    log_parameters=("phot_rp_mean_flux", ),
-    labels=dict(phot_rp_mean_flux=r"\textrm{mean rp flux}"))
+    log_parameters=("phot_rp_mean_flux", ), labels=dict(
+        phot_rp_mean_flux=r"\textrm{mean rp flux}",
+        bp_rp=r"\textrm{bp-rp}"))
 
 raise a
 

### 2018-05-09 velociraptor.py

@@ -201,55 +201,85 @@ def plot_model_predictions_corner(samples, sources=None, parameter_limits=None,
     
     P = len(limits)
 
+    # Calculate the expected radial velocity variance for all combinations of
+    # parameters.
+    combinations = np.meshgrid(*[mesh(pn) for pn in parameter_names])
+    grid_combinations = np.vstack([comb.flatten() for comb in combinations])
+
+    expectation = np.dot(coefficients, _rvf_design_matrix(**dict(zip(
+        parameter_names, grid_combinations))))
+
     fig, axes = plt.subplots(P, P, figsize=(6 * P, 6 * P))
     axes = np.atleast_2d(axes)
+
 
     for i, x_param in enumerate(parameter_names):
         for j, y_param in enumerate(parameter_names):
 
-            ax = axes[i, j]
+            ax = axes[j, i]
 
             if i > j:
                 ax.set_visible(False)
                 continue
 
             elif i == j:
+                x = grid_combinations[i]
+                idx = np.argsort(x)
 
-                x = mesh(x_param)
+                # Get the mean at each unique x.
+                x_uniques = np.sort(np.unique(x))
+                y_percentiles = np.zeros((3, x_uniques.size), dtype=float)
 
-                map_value = np.dot(coefficients, _rvf_design_matrix(**dict([
-                    (x_param, x.flatten())])))
+                for k, x_unique in enumerate(x_uniques):
+                    match = (grid_combinations[i] == x_unique)
+                    y_percentiles[:, k] = np.percentile(
+                        expectation[match], [0, 50, 100])
 
-                ax.plot(x, map_value, 'r-')
-
+                ax.plot(x_uniques, y_percentiles[1], "r-")
+                ax.fill_between(
+                    x_uniques, y_percentiles[0], y_percentiles[2],
+                    facecolor="r", alpha=0.3, edgecolor="none")
+                
                 if x_param in log_parameters:
                     ax.semilogx()
 
-            else:
-
-                x, y = mesh(x_param), mesh(y_param)
-                x_mesh, y_mesh = np.meshgrid(x, y)
-
-                map_value = np.dot(coefficients, _rvf_design_matrix(**dict([
-                    (x_param, x_mesh.flatten()),
-                    (y_param, y_mesh.flatten())
-                ])))
-
-
-            ax.yaxis.set_major_locator(MaxNLocator(6))
-
-            if ax.is_first_col():
[…]

### 2018-05-09 velociraptor.py
wip

@@ -163,6 +163,11 @@ def plot_model_predictions_corner(samples, sources=None, parameter_limits=None,
     :param parameter_limits: [optional]
         A dictionary containing source parameters as keys and a two-length
         tuple containing the lower and upper bounds of the parameter.
+
+    :param log_parameters: [optional]
+        A tuple containing the parameter names that should be shown in log space
+
+    ""
     """
 
     parameter_names = tuple(
@@ -259,26 +264,29 @@ def plot_model_predictions_corner(samples, sources=None, parameter_limits=None,
                 _x = np.log10(x) if x_param in log_parameters else x
                 _y = np.log10(y) if y_param in log_parameters else y
 
+                # failed
+                #imshow_kwds = dict(cmap="Reds", aspect=np.ptp(x)/np.ptp(y),
+                #    extent=(np.min(x), np.max(x), np.max(y), np.min(y)))
 
-                imshow_kwds = dict(cmap="Reds", aspect=np.ptp(_x)/np.ptp(_y),
+                imshow_kwds = dict(cmap="Reds", aspect="auto",
+                    extent=(np.min(x), np.max(x), np.max(y), np.min(y)))
+
+                imshow_kwds = dict(cmap="Reds", aspect="auto",
                     extent=(np.min(_x), np.max(_x), np.max(_y), np.min(_y)))
 
-                ax.imshow(expectation.reshape((N, N)), **imshow_kwds)
-
-                """
                 if x_param in log_parameters:
                     ax.semilogx()
 
                 if y_param in log_parameters:
                     ax.semilogy()
-                """
+
+                ax.imshow(expectation.reshape((N, N)), **imshow_kwds)
 
                 #    extent=(np.min(x), np.max(x), np.max(y), np.min(y)),
                 #    aspect=np.ptp(x)/np.ptp(y), cmap="Reds")
 
                 ax.set_xlabel(labels.get(x_param, x_param))
                 ax.set_ylabel(labels.get(y_param, y_param))
-                raise a
 
 
     fig.tight_layout()

### 2018-05-09 modeling.py
figures

@@ -13,26 +13,24 @@ sources = velociraptor.load_gaia_sources("data/rv-cal-subset.fits.gz", N=1e3)
 
 model, data_dict, init_dict, used_in_fit = velociraptor.prepare_model(**sources)
 
+p_opt = model.optimizing(data=data_dict, init=init_dict)
+
 samples = model.sampling(**velociraptor.stan.sampling_kwds(
-    data=data_dict, chains=2, iter=2000, init=init_dict))
+    data=data_dict, chains=2, iter=2000, init=p_opt))
 
 
-velociraptor.plot_model_predictions_corner(samples, sources,
+fig = velociraptor.plot_model_predictions_corner(samples, sources,
     log_parameters=("phot_rp_mean_flux", ), labels=dict(
         phot_rp_mean_flux=r"\textrm{mean rp flux}",
         bp_rp=r"\textrm{bp-rp}"))
-
-raise a
-
-
+fig.savefig("figures/model-predictions.pdf", dpi=150)
 
 scatter_kwds = dict(s=1, facecolor="#000000", alpha=0.5, rasterized=True)
 model_scatter_kwds = dict(s=1, facecolor="r", alpha=0.5, rasterized=True)
 
 
 model_rv_sev_mu, model_rv_sev_sigma = \
-    velociraptor.predict_map_rv_single_epoch_variance(
-        samples, phot_rp_mean_flux=sources["phot_rp_mean_flux"])
+    velociraptor.predict_map_rv_single_epoch_variance(samples, **sources)
 
 # Plot the model against various properties of interest.
 shorthand_parameters = [
@@ -62,20 +60,6 @@ for label_name, description, is_semilogx in shorthand_parameters:
 
     fig, axes = plt.subplots(1, 2, figsize=(12, 6))
 
-    """
-    ax.plot(x[idx], model_rv_sev_mu[idx], "r-")
-    ax.fill_between(
-        x[idx],
-        (model_rv_sev_mu - model_rv_sev_sigma)[idx],
-        (model_rv_sev_mu + model_rv_sev_sigma)[idx],
-        facecolor="r", alpha=0.3, edgecolor="none")
-    """
-
-
-    """
-    ax.errorbar(x, model_rv_sev_mu, yerr=model_rv_sev_sigma, fmt='.',
-        c="r", capsize=0, alpha=0.05)
-    """
     for ax in axes:
         ax.scatter(x, y, **model_scatter_kwds)
 
[…]

### 2018-05-09 velociraptor.py

@@ -8,11 +8,12 @@ import numpy as np
 import matplotlib.pyplot as plt
 from astropy.table import Table
 from matplotlib.ticker import MaxNLocator
+from scipy.stats import norm
 
 import stan_utils as stan
-from mpl_style import mpl_style
+import mpl_utils
 
-plt.style.use(mpl_style)
+plt.style.use(mpl_utils.mpl_style)
 
 np.random.seed(42)
 
@@ -141,6 +142,12 @@ def binary_probability(samples, **source_params):
     from our MCMC.
     """
 
+    max_rv_variance = np.nanmax(source_params["rv_single_epoch_variance"])
+
+
+    log_ps1 = np.log(samples["theta"]) - np.log(max_rv_variance)
+    log_ps2 = np.log(1 - samples["theta"]) 
+
     raise NotImplementedError()
 
 
@@ -229,7 +236,6 @@ def plot_model_predictions_corner(samples, sources=None, parameter_limits=None,
 
             elif i == j:
                 x = grid_combinations[i]
-                idx = np.argsort(x)
 
                 # Get the mean at each unique x.
                 x_uniques = np.sort(np.unique(x))
@@ -255,25 +261,12 @@ def plot_model_predictions_corner(samples, sources=None, parameter_limits=None,
             else:
                 x, y = grid_combinations[[i, j]]
 
-                #if x_param in log_parameters:
-                #    ax.semilogx()
-
-                #if y_param in log_parameters:
-                #    ax.semilogy()
-
-                _x = np.log10(x) if x_param in log_parameters else x
-                _y = np.log10(y) if y_param in log_parameters else y
-
-                # failed
-                #imshow_kwds = dict(cmap="Reds", aspect=np.ptp(x)/np.ptp(y),
-                #    extent=(np.min(x), np.max(x), np.max(y), np.min(y)))
-
-                imshow_kwds = dict(cmap="Reds", aspect="auto",
+                #_x = np.log10(x) if x_param in log_parameters else x
+                #_y = np.log10(y) if y_param in log_parameters else y
+                
+                imshow_kwds = dict(cmap="Reds", aspect="equal",
                     extent=(np.min(x), np.max(x), np.max(y), np.min(y)))
-
-                imshow_kwds = dict(cmap="Reds", aspect="auto",
-                    extent=(np.min(_x), np.max(_x), np.max(_y), np.min(_y)))
-
+                
                 if x_param in log_parameters:
[…]

### 2018-05-09 modeling.py
run

@@ -49,6 +49,9 @@ shorthand_parameters = [
     ("absolute_g_mag", "absolute g magnitude", False),
     ("absolute_bp_mag", "absolute bp magnitude", False),
     ("absolute_rp_mag", "absolute rp magnitude", False),
+    ("rv_template_teff", "rv template teff", False),
+    ("rv_template_logg", "rv template logg", False),
+    ("rv_template_fe_h", "rv template [Fe/H]", False),
 ]
 
 for label_name, description, is_semilogx in shorthand_parameters:
@@ -228,6 +231,26 @@ for label_name, description, is_semilogx in shorthand_parameters:
         ax.set_xlabel(r"\textrm{{{0}}}".format(description))
 
     fig.tight_layout()
+
+
+# OK let's do this for a sample of giants and a sample of dwarfs and then
+# compare.
+
+qc = (sources["parallax"] > 0) \
+   * ((sources["parallax"]/sources["parallax_error"]) > 5)
+
+fig, ax = plt.subplots()
+ax.scatter(sources["bp_rp"], sources["absolute_g_mag"], **scatter_kwds)
+ax.set_ylim(ax.get_ylim()[::-1])
+ax.set_xlabel(r"\textrm{bp - rp}")
+ax.set_ylabel(r"\textrm{absolute g magnitude}")
+fig.tight_layout()
+
+
+qc_giants = qc * (3 < sources["bp_rp"]) * (sources["bp_rp"] < 1.0) \
+          * (sources["absolute_g_mag"] > 2.5)
+qc_dwarfs = qc * (3 < sources["bp_rp"]) * (sources["bp_rp"] < 1.0) \
+          * (sources["absolute_g_mag"] < 2.5)
 
 
 

### 2018-05-09 model.py
50% speedup in stan model

@@ -1,576 +1,0 @@
-
-"""
-Model the radial velocity floor in Gaia data.
-"""
-
-import numpy as np
-import matplotlib.pyplot as plt
-import scipy.optimize as op
-from mpl_style import mpl_style
-from matplotlib.ticker import MaxNLocator
-from astropy.table import Table
-
-import stan_utils as stan
-
-plt.style.use(mpl_style)
-
-np.random.seed(42)
-
-def _savefig(fig, basename, dry_run=False, **kwargs):
-    kwds = dict(dpi=300)
-    kwds.update(kwargs)
-    if not dry_run:
-        fig.savefig("figures/{}.pdf".format(basename), **kwds)
-        fig.savefig("figures/{}.png".format(basename), **kwds)
-    else:
-        print("Dry-run: not saving figure")
-
-    return None
-
-
-#gaia_sources = Table.read("data/rv_floor_cal-result.fits")
-
-subset = Table.read("data/rv_calibration_floor_subset-result.fits.gz")
-
-# Take a random subset of stars.
-subset_indices = np.random.choice(len(subset), size=10000, replace=False)
-subset = subset[subset_indices]
-
-# Take the rv_nb_transits into account
-subset["radial_velocity_scatter"] = subset["radial_velocity_error"] \
-                                  * np.sqrt(subset["rv_nb_transits"]) \
-                                  / np.sqrt(np.pi/2.0)
-
-
-rv_variance = np.array(subset["radial_velocity_scatter"])**2
-apparent_magnitude = np.array(subset["phot_g_mean_mag"])
-flux = np.array(subset["phot_g_mean_flux"])
-rp_flux = np.array(subset["phot_rp_mean_flux"])
-
-scatter_kwds = dict(s=1, facecolor="#000000", alpha=0.05, rasterized=True)
-
-
-
-fig, ax = plt.subplots()
-ax.scatter(flux, rv_variance, **scatter_kwds)
-ax.semilogx()
-ax.set_xlabel(r"\textrm{mean g flux}")
-ax.set_ylabel(r"\textrm{radial velocity variance} $(\textrm{km\,s}^{-1})$")
-ax.set_ylim(-0.5, 10)
-ax.yaxis.set_major_locator(MaxNLocator(6))
-fig.tight_layout()
[…]

### 2018-05-09 modeling.py

@@ -9,7 +9,7 @@ import matplotlib.pyplot as plt
 import velociraptor
 
 # Load the data and make some plots.
-sources = velociraptor.load_gaia_sources("data/rv-cal-subset.fits.gz", N=1e3)
+sources = velociraptor.load_gaia_sources("data/rv-cal-subset.fits.gz", N=1e4)
 
 model, data_dict, init_dict, used_in_fit = velociraptor.prepare_model(**sources)
 

### 2018-05-09 modeling.py
minor

@@ -9,7 +9,7 @@ import matplotlib.pyplot as plt
 import velociraptor
 
 # Load the data and make some plots.
-sources = velociraptor.load_gaia_sources("data/rv-cal-subset.fits.gz", N=1e3)
+sources = velociraptor.load_gaia_sources("data/rv-cal-subset.fits.gz", N=1e4)
 
 model, data_dict, init_dict, used_in_fit = velociraptor.prepare_model(**sources)
 
@@ -247,11 +247,59 @@ ax.set_ylabel(r"\textrm{absolute g magnitude}")
 fig.tight_layout()
 
 
-qc_giants = qc * (3 < sources["bp_rp"]) * (sources["bp_rp"] < 1.0) \
-          * (sources["absolute_g_mag"] > 2.5)
-qc_dwarfs = qc * (3 < sources["bp_rp"]) * (sources["bp_rp"] < 1.0) \
-          * (sources["absolute_g_mag"] < 2.5)
+qc_giants = qc * (3 > sources["bp_rp"]) * (sources["bp_rp"] > 1.0) \
+          * (sources["absolute_g_mag"] < 3)
+qc_dwarfs = qc * (3 > sources["bp_rp"]) * (sources["bp_rp"] > 1.0) \
+          * (sources["absolute_g_mag"] > 4)
 
 
+fig, axes = plt.subplots(1, 2, figsize=(8, 4))
+
+for ax in axes:
+    ax.scatter(
+        sources["bp_rp"][qc],
+        sources["absolute_g_mag"][qc],
+        **scatter_kwds)
+
+    ax.set_ylim(ax.get_ylim()[::-1])
+
+    ax.set_xlabel(r"\textrm{bp - rp}")
+    ax.set_ylabel(r"\textrm{absolute g magnitude}")
+
+axes[0].scatter(
+    sources["bp_rp"][qc_dwarfs],
+    sources["absolute_g_mag"][qc_dwarfs],
+    s=5, zorder=10)
+axes[1].scatter(
+    sources["bp_rp"][qc_giants],
+    sources["absolute_g_mag"][qc_giants],
+    s=5, zorder=10)
+
+axes[0].set_title(r"\textrm{main-sequence stars}")
+axes[1].set_title(r"\textrm{giant stars}")
+
+fig.tight_layout()
+
+
+print("Number of stars in dwarf model: {}".format(sum(qc_dwarfs)))
+print("Number of stars in giant model: {}".format(sum(qc_giants)))
+
+
[…]

### 2018-05-09 modeling.py
minor

@@ -247,11 +247,59 @@ ax.set_ylabel(r"\textrm{absolute g magnitude}")
 fig.tight_layout()
 
 
-qc_giants = qc * (3 < sources["bp_rp"]) * (sources["bp_rp"] < 1.0) \
-          * (sources["absolute_g_mag"] > 2.5)
-qc_dwarfs = qc * (3 < sources["bp_rp"]) * (sources["bp_rp"] < 1.0) \
-          * (sources["absolute_g_mag"] < 2.5)
+qc_giants = qc * (3 > sources["bp_rp"]) * (sources["bp_rp"] > 1.0) \
+          * (sources["absolute_g_mag"] < 3)
+qc_dwarfs = qc * (3 > sources["bp_rp"]) * (sources["bp_rp"] > 1.0) \
+          * (sources["absolute_g_mag"] > 4)
 
 
+fig, axes = plt.subplots(1, 2, figsize=(8, 4))
+
+for ax in axes:
+    ax.scatter(
+        sources["bp_rp"][qc],
+        sources["absolute_g_mag"][qc],
+        **scatter_kwds)
+
+    ax.set_ylim(ax.get_ylim()[::-1])
+
+    ax.set_xlabel(r"\textrm{bp - rp}")
+    ax.set_ylabel(r"\textrm{absolute g magnitude}")
+
+axes[0].scatter(
+    sources["bp_rp"][qc_dwarfs],
+    sources["absolute_g_mag"][qc_dwarfs],
+    s=5, zorder=10)
+axes[1].scatter(
+    sources["bp_rp"][qc_giants],
+    sources["absolute_g_mag"][qc_giants],
+    s=5, zorder=10)
+
+axes[0].set_title(r"\textrm{main-sequence stars}")
+axes[1].set_title(r"\textrm{giant stars}")
+
+fig.tight_layout()
+
+
+print("Number of stars in dwarf model: {}".format(sum(qc_dwarfs)))
+print("Number of stars in giant model: {}".format(sum(qc_giants)))
+
+
+# Run the model for each subset.
+dwarf_model, dwarf_data_dict, dwarf_init_dict, dwarf_used_in_fit \
+    = velociraptor.prepare_model(**sources[qc_dwarfs])
+
+dwarf_p_opt = dwarf_model.optimizing(data=dwarf_data_dict, init=dwarf_init_dict)
+dwarf_samples = dwarf_model.sampling(**velociraptor.stan.sampling_kwds(
[…]

### 2018-05-10 modeling.py
wip

@@ -296,10 +296,135 @@ dwarf_samples = dwarf_model.sampling(**velociraptor.stan.sampling_kwds(
 
 
 giant_model, giant_data_dict, giant_init_dict, giant_used_in_fit \
-    = velociraptor.prepare_model(**sources[qc_dwarfs])
+    = velociraptor.prepare_model(**sources[qc_giants])
 
 giant_p_opt = giant_model.optimizing(data=giant_data_dict, init=giant_init_dict)
 giant_samples = giant_model.sampling(**velociraptor.stan.sampling_kwds(
     data=giant_data_dict, chains=2, iter=2000, init=giant_p_opt))
+
+
+
+
+# Plot the performance of the two models on the same figure.
+giant_model_rv_sev_mu, giant_model_rv_sev_sigma = \
+    velociraptor.predict_map_rv_single_epoch_variance(giant_samples, **sources[qc_giants])
+
+dwarf_model_rv_sev_mu, dwarf_model_rv_sev_sigma = \
+    velociraptor.predict_map_rv_single_epoch_variance(dwarf_samples, **sources[qc_dwarfs])
+
+shorthand_parameters = [
+    ("phot_rp_mean_flux", "mean rp flux", True),
+    ("phot_bp_mean_flux", "mean bp flux", True),
+    ("phot_g_mean_flux", "mean g flux", True),
+    ("bp_rp", "bp-rp colour", False),
+    ("phot_g_mean_mag", "mean g magnitude", False),
+    ("phot_rp_mean_mag", "mean rp magnitude", False),
+    ("phot_bp_mean_mag", "mean bp magnitude", False),
+]
+
+def _show_number_of_data_points(ax, N):
+    return ax.text(0.95, 0.95, r"${0:.0f}$".format(N), transform=ax.transAxes,
+        horizontalalignment="right", verticalalignment="top")
+
+def _smooth_model(x, y, yerr, is_semilogx, N_smooth_points=100, equidensity=True,
+    average_function=np.median):
+    if not equidensity:
+        if is_semilogx:
+            xi = np.logspace(np.log10(x.min()), np.log10(x.max()), N_smooth_points)
+        else:
+            xi = np.linspace(x.min(), x.max(), N_smooth_points)
+
+    else:
+        if is_semilogx:
+            xi = 10**np.percentile(np.log10(x), np.linspace(0, 100, N_smooth_points))
[…]

### 2018-05-10 article/ms.tex
ack

@@
 \shorttitle{Stellar multiplicity}
-\shortauthors{Casey \& Hogg}
+\shortauthors{Casey et al.?}
 
@@
 
-\author[0000-0003-2866-9403]{David W. Hogg}
-\affiliation{Center for Cosmology and Particle Physics, Department of Physics,
-    New York University, 726 Broadway, New York, NY 10003, USA}
-\affiliation{Center for Data Science, New York University, 60 Fifth Ave, 
-    New York, NY 10011, USA}
-\affiliation{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 
-    Heidelberg}
-\affiliation{Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
+%\author[0000-0003-2866-9403]{David W. Hogg??}
+%\affiliation{Center for Cosmology and Particle Physics, Department of Physics,
+%    New York University, 726 Broadway, New York, NY 10003, USA}
+%\affiliation{Center for Data Science, New York University, 60 Fifth Ave, 
+%    New York, NY 10011, USA}
+%\affiliation{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 
+%    Heidelberg}
+%\affiliation{Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
 
@@
 \acknowledgements
+- If these peeps are not co-authors: Hogg, Spina, Schlaufman.
 - Gaia/DPAC
@@
 - Gaia DR2 hack week at Flatiron
+- ARC DP
+
+%This work has made use of CosmoHub.
+
+%CosmoHub has been developed by the Port d'Informació Científica (PIC), maintained through a collaboration of the Institut de Física d'Altes Energies (IFAE) and the Centro de Investigaciones Energéticas, Medioambientales y Tecnológicas (CIEMAT), and was partially funded by the "Plan Estatal de Investigación Científica y Técnica y de Innovación" program of the Spanish government.
 

### 2018-05-10 modeling.py
refactor stan model (~4x speedup)

@@ -320,30 +320,41 @@ shorthand_parameters = [
     ("phot_g_mean_mag", "mean g magnitude", False),
     ("phot_rp_mean_mag", "mean rp magnitude", False),
     ("phot_bp_mean_mag", "mean bp magnitude", False),
+    ("teff_val", "inferred temperature", False),
+    ("rv_template_teff", "rv template teff", False),
+    ("rv_template_logg", "rv template logg", False),
+    ("rv_template_fe_h", "rv template [Fe/H]", False),
 ]
 
-def _show_number_of_data_points(ax, N):
-    return ax.text(0.95, 0.95, r"${0:.0f}$".format(N), transform=ax.transAxes,
+def _show_text_upper_right(ax, text):
+    return ax.text(0.95, 0.95, text, transform=ax.transAxes,
         horizontalalignment="right", verticalalignment="top")
 
-def _smooth_model(x, y, yerr, is_semilogx, N_smooth_points=100, equidensity=True,
-    average_function=np.median):
-    if not equidensity:
+def _show_number_of_data_points(ax, N):
+    return _show_text_upper_right(ax, r"${0:.0f}$".format(N))
+
+
+N_smooth_points = 30
+equidensity = True
+
+def _smooth_model(x, y, yerr, is_semilogx, average_function=np.mean):
+    if equidensity:
         if is_semilogx:
-            xi = np.logspace(np.log10(x.min()), np.log10(x.max()), N_smooth_points)
+            xb = 10**np.percentile(np.log10(x), np.linspace(0, 100, N_smooth_points))
         else:
-            xi = np.linspace(x.min(), x.max(), N_smooth_points)
+            xb = np.percentile(x, np.linspace(0, 100, N_smooth_points))
 
     else:
         if is_semilogx:
-            xi = 10**np.percentile(np.log10(x), np.linspace(0, 100, N_smooth_points))
+            xb = np.logspace(np.log10(x.min()), np.log10(x.max()), N_smooth_points)
         else:
-            xi = np.percentile(x, np.linspace(0, 100, N_smooth_points))
+            xb = np.linspace(x.min(), x.max(), N_smooth_points)
 
+    xi = xb[:-1] + np.diff(xb)/2.
     yi = np.zeros_like(xi)
     yerri = np.zeros_like(xi)
-    for i, left_edge in enumerate(xi[:-1]):
[…]

### 2018-05-10 modeling.py
stupid stan

@@ -1,438 +1,25 @@
 
 """
-Modeling and validation of the radial velocity calibration model.
+Fitting and sampling of the radial velocity variance model.
 """
 
-import numpy as np
-import matplotlib.pyplot as plt
-
+import pickle
 import velociraptor
 
 # Load the data and make some plots.
-sources = velociraptor.load_gaia_sources("data/rv-cal-subset.fits.gz", N=1e4)
+sources = velociraptor.load_gaia_sources("data/rv-all.fits.gz")
 
 model, data_dict, init_dict, used_in_fit = velociraptor.prepare_model(**sources)
 
+print("Number of data points: {}".format(sum(used_in_fit)))
+
 p_opt = model.optimizing(data=data_dict, init=init_dict)
 
-samples = model.sampling(**velociraptor.stan.sampling_kwds(
+sampled_model = model.sampling(**velociraptor.stan.sampling_kwds(
     data=data_dict, chains=2, iter=2000, init=p_opt))
 
-
-fig = velociraptor.plot_model_predictions_corner(samples, sources,
-    log_parameters=("phot_rp_mean_flux", ), labels=dict(
-        phot_rp_mean_flux=r"\textrm{mean rp flux}",
-        bp_rp=r"\textrm{bp-rp}"))
-fig.savefig("figures/model-predictions.pdf", dpi=150)
-
-scatter_kwds = dict(s=1, facecolor="#000000", alpha=0.5, rasterized=True)
-model_scatter_kwds = dict(s=1, facecolor="r", alpha=0.5, rasterized=True)
-
-
-model_rv_sev_mu, model_rv_sev_sigma = \
-    velociraptor.predict_map_rv_single_epoch_variance(samples, **sources)
-
-# Plot the model against various properties of interest.
-shorthand_parameters = [
-    ("phot_rp_mean_flux", "mean rp flux", True),
-    ("phot_bp_mean_flux", "mean bp flux", True),
-    ("phot_g_mean_flux", "mean g flux", True),
-    ("bp_rp", "bp-rp colour", False),
-    ("phot_g_mean_mag", "mean g magnitude", False),
-    ("phot_rp_mean_mag", "mean rp magnitude", False),
-    ("phot_bp_mean_mag", "mean bp magnitude", False),
[…]

### 2018-05-10 validation.py

@@ -1,0 +1,439 @@
+
+"""
+Validation of the radial velocity calibration model.
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+import velociraptor
+
+# Load the data and make some plots.
+sources = velociraptor.load_gaia_sources("data/rv-cal-subset.fits.gz", N=1e4)
+
+model, data_dict, init_dict, used_in_fit = velociraptor.prepare_model(**sources)
+
+p_opt = model.optimizing(data=data_dict, init=init_dict)
+
+samples = model.sampling(**velociraptor.stan.sampling_kwds(
+    data=data_dict, chains=2, iter=2000, init=p_opt))
+
+
+fig = velociraptor.plot_model_predictions_corner(samples, sources,
+    log_parameters=("phot_rp_mean_flux", ), labels=dict(
+        phot_rp_mean_flux=r"\textrm{mean rp flux}",
+        bp_rp=r"\textrm{bp-rp}"))
+fig.savefig("figures/model-predictions.pdf", dpi=150)
+
+scatter_kwds = dict(s=1, facecolor="#000000", alpha=0.5, rasterized=True)
+model_scatter_kwds = dict(s=1, facecolor="r", alpha=0.5, rasterized=True)
+
+
+model_rv_sev_mu, model_rv_sev_sigma = \
+    velociraptor.predict_map_rv_single_epoch_variance(samples, **sources)
+
+# Plot the model against various properties of interest.
+shorthand_parameters = [
+    ("phot_rp_mean_flux", "mean rp flux", True),
+    ("phot_bp_mean_flux", "mean bp flux", True),
+    ("phot_g_mean_flux", "mean g flux", True),
+    ("bp_rp", "bp-rp colour", False),
+    ("phot_g_mean_mag", "mean g magnitude", False),
+    ("phot_rp_mean_mag", "mean rp magnitude", False),
+    ("phot_bp_mean_mag", "mean bp magnitude", False),
+    ("teff_val", "inferred temperature", False),
+    ("ra", "right ascension", False),
+    ("dec", "declination", False),
+    ("radial_velocity", "radial velocity", False),
+    ("rv_nb_transits", "number of radial velocity transits", False),
+    ("absolute_g_mag", "absolute g magnitude", False),
[…]

### 2018-05-10 modeling.py
checks

@@ -7,7 +7,7 @@ import pickle
 import velociraptor
 
 # Load the data and make some plots.
-sources = velociraptor.load_gaia_sources("data/rv-all.fits.gz")
+sources = velociraptor.load_gaia_sources("data/rv-all.fits")
 
 model, data_dict, init_dict, used_in_fit = velociraptor.prepare_model(**sources)
 
@@ -15,11 +15,16 @@ print("Number of data points: {}".format(sum(used_in_fit)))
 
 p_opt = model.optimizing(data=data_dict, init=init_dict)
 
+with open("model-optimized.pkl", "wb") as fp:
+    pickle.dump(p_opt, fp, -1)
+
+print(p_opt)
+
 sampled_model = model.sampling(**velociraptor.stan.sampling_kwds(
     data=data_dict, chains=2, iter=2000, init=p_opt))
 
 samples = sampled_model.extract(("theta", "mu_coefficients", 
     "sigma_coefficients", "log_membership_probability"))
 
-with open("model-samples.pkl", "wb") as fp:
+with open("model-sampled.pkl", "wb") as fp:
     pickle.dump(samples, fp, -1)

### 2018-05-10 velociraptor.py

@@ -76,12 +76,13 @@ def prepare_model(rv_single_epoch_variance, target=0.01, model_path="model.stan"
     """
 
     dm = _rvf_design_matrix(**source_params)
-    finite = np.all(np.isfinite(dm), axis=0)
+    finite = np.all(np.isfinite(dm), axis=0) \
+           * np.isfinite(rv_single_epoch_variance)
 
     if not all(finite):
-        logging.warn("Design matrix contains {0} (of {1}) non-finite values!"\
-                     "Exlcuding them from model fit.".format(
-                        sum(~finite), finite.size))
+        logging.warn("Design matrix contains {0} (of {1}; {2} are finite) "\
+                     "non-finite values! Exlcuding them from model fit."\
+                     .format(sum(~finite), finite.size, sum(finite)))
     
     dm = dm[:, finite]
     coeff = _rvf_initial_coefficients(dm, target=target)

### 2018-05-20 article/ms.tex
words

@@
 \usepackage{bm}
+\usepackage{color}
 
@@
 \newcommand{\project}[1]{\textsl{#1}}
+\newcommand{\package}[1]{\texttt{#1}}
 \newcommand{\acronym}[1]{{\small{#1}}}
@@
 \newcommand{\gaia}{\project{gaia}}
+\newcommand{\todo}[1]{\textcolor{red}{#1}}
+
+\newcommand{\GaiaRVE}{\sigma_{\mathrm{V}_\mathrm{R}}^\mathrm{MTA}}
+\newcommand{\RVJitter}{\sigma(\mathrm{V}_\mathrm{R}^{t})}
 
@@
 \accepted{2018 XX XX}
+
+%\graphicspath{../figures}
 
@@
 
-\title{Stellar multiplicity}
+\title{Stellar multiplicity in \Gaia}
 
@@
     Victoria, Australia}
+
+%\author{Daniel Foreman-Mackey}
+%\affiliation{Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
 
@@
 \begin{abstract}
-Velociraptors can't see you if you don't move. (\Gaia can see you if you can.)
+Most stars are in binary systems. 
+The occurrence rate of binary star systems is critically important to many 
+fields of astrophysics, but the effects of binaries are commonly ignored in 
+astronomical data analysis applications.
+Here we construct a model to probabilistically identify binary stars based on
+excess radial velocity scatter, photometry, and astrometric excess noise.
+
+We identify binaries \todo{within some orbital and SP limits} and show that 
+the orbital semi-amplitude is well-approximated by the excess radial velocity
+scatter for systems within these criteria.
+
+\todo{We make a few interesting tests about binarity...}
 \end{abstract}
@@
 
-\section{Model}
+\section{Method}
+
+There are at least three ways that binary systems can be reliably
+identified using from \Gaia data: through radial velocity variations, 
+anomalous photometric colours, or excess astrometric noise. In this 
+work we do not make use of astrometric noise because of the 
+degree-of-freedom astrometry bug present in the second \Gaia\ data 
+release \citep{Lindegren:2018}. We will describe the radial velocity
+information that forms part of our model for identifying binary stars,
+and then discuss the photometric information used, before explicitly
+outlining the assumptions of our model.
+
+The second \Gaia\ data release provides median radial velocity 
+estimates for 7,224,631 stars. The radial velocity error provided
+($\GaiaRVE$; \texttt{radial\_velocity\_error}) is a function of the 
+number of transits $N$ (i.e., observations; column 
+\texttt{rv\_nb\_transits}) and the scatter among those $N$ measurements. 
+This allows us to calculate the scatter among those measurements, or
+rather the \emph{radial velocity jitter}, for each source
+\begin{eqnarray}
+\RVJitter = \sqrt{\frac{2N}{\pi}}\GaiaRVE \quad .
+\end{eqnarray}
+
+For a single star system observed $N$ times, the source radial velocity
+jitter represents either the minimum uncertainty with which
+\Gaia\ can measure radial velocity for a star of that apparent 
+brightness and colour. However for stars with stellar companions,
+the radial velocity scatter is a function of orbital parameters 
+and observing epochs. Thus we can reasonably assume that the
+radial velocity jitter will be a mixture of
+two populations: single stars, where the radial velocity jitter
+represents the minimum radial velocity noise that \Gaia\ can
+measure for a star of that type, and a binary population that
+could have radial velocity variances well above the noise floor.
+
+\begin{figure*}
+	\includegraphics[width=1.0\textwidth]{../figures/sb9-comparison.pdf}
+    \caption{Source radial velocity variance shown with respect to flux in
+    the \Gaia\ BP, G, and RP bands. The source radial velocity variance 
+    most closely follows the RP flux as it is closer to the \Gaia\ RVS
+    band.}
+\end{figure*}
+
+\begin{figure}
+	\includegraphics[width=0.5\textwidth]{../figures/sb9-comparison.pdf}
+    \caption{Source mean \Gaia\ RP flux and radial velocity variance.
+    Points are represented by their mean $BP-RP$ color, confirming that
+    there is some color dependency on the source radial velocity variance.}
+    \label{fig:rp-flux-color}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.5\textwidth]{../figures/sb9-comparison.pdf}
+    \caption{Mixture model fitted.}
+    \label{fig:mixture-model}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.5\textwidth]{../figures/sb9-comparison.pdf}
+    \caption{Detected binary fraction as a function of the mean RP flux
+    (top panels), the $BP-RP$ color (second panels), the number of 
+    radial velocity transits (third panels), and the inferred temperature
+    (bottom panels). No trend is observed with mean RP flux or the number 
+    of radial velocity transits. However, our model finds bluer (hotter)
+    stars are more likely to be binaries, despite including color as a
+    dependency in our model.}
+    \label{fig:binary-fraction-wrt-params}
+\end{figure}
+
+% we assume what?
+%
+
+% given these assumptions, the model we adopt is XYZ
+% we chose to model in RP flux as it more closely represents the flux counts that would be received by the Gaia RVS, rather than G or BP flux.
+
+% the model we use is
+
+%
 
@@
 
-\software{numpy, scipy, astropy, Stan}
+\acknowledgements
 
-\acknowledgements
-- If these peeps are not co-authors: Hogg, Spina, Schlaufman.
-- Gaia/DPAC
-- Gaia archive?
-- Gaia DR2 hack week at Flatiron
-- ARC DP
+It is a pleasure to thank
+	David W. Hogg,
+	Lorenzo Spina, 
+		and 
+	Kevin C. Schlaufman.
+This work was partly supported through the Australian Research Council 
+through Discovery Grant DP160100637.
+This work has made use of data from the European Space Agency (ESA) mission {\it
+Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia} Data
+Processing and Analysis Consortium (DPAC,
+\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC
+has been provided by national institutions, in particular the institutions
+participating in the {\it Gaia} Multilateral Agreement.  This research was
+started at the NYC Gaia DR2 Workshop at the Center for Computational
+Astrophysics of the Flatiron Institute in 2018 April.
 
-%This work has made use of CosmoHub.
+This work has made use of CosmoHub.
+CosmoHub has been developed by the Port d'Informaci\'o Cient\'ifica (PIC), maintained through a collaboration of the Institut de F\'isica d'Altes Energies (IFAE) and the Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), and was partially funded by the "Plan Estatal de Investigaci\'on Científica y T\'ecnica y de Innovaci\'on" program of the Spanish government.
 
-%CosmoHub has been developed by the Port d'Informació Científica (PIC), maintained through a collaboration of the Institut de Física d'Altes Energies (IFAE) and the Centro de Investigaciones Energéticas, Medioambientales y Tecnológicas (CIEMAT), and was partially funded by the "Plan Estatal de Investigación Científica y Técnica y de Innovación" program of the Spanish government.
+\software{
+	\package{Astropy} \citep{astropy},
+    \package{IPython} \citep{ipython},
+    \package{matplotlib} \citep{mpl},
+    \package{numpy} \citep{numpy},
+    \package{scipy} \citep{scipy},
+    \package{Stan} \citep{stan},
+    \package{CosmoHub} \citep{cosmohub}
+    %tensorflow
+}
+
+\bibliographystyle{aasjournal}
+\bibliography{velociraptor}
 

### 2018-05-20 data/download.py
wip

@@ -49,3 +49,23 @@ job = Gaia.launch_job_async("""
     """.format(", ".join(columns)))
 sources = job.get_results()
 sources.write("rv_calibration_floor-result.fits")
+
+
+
+all_possible_sources = Gaia.launch_job_async("""
+    SELECT  *
+    FROM    gaiadr2.gaia_source
+    WHERE   phot_rp_mean_mag < 13
+    """)
+
+
+"""
+SELECT  *
+FROM    gaiadr2.gaia_source
+WHERE   radial_velocity IS NOT NULL
+AND     duplicated_source = 'false'
+AND     rv_nb_transits > 10
+AND     visibility_periods_used > 10
+AND     radial_velocity_error < 20
+AND     mod(random_index, 10) = 0
+"""

### 2018-05-20 validation.py

@@ -9,17 +9,122 @@ import matplotlib.pyplot as plt
 import velociraptor
 
 # Load the data and make some plots.
-sources = velociraptor.load_gaia_sources("data/rv-cal-subset.fits.gz", N=1e4)
+sources = velociraptor.load_gaia_sources("data/rv-cal-subset.fits.gz")
 
-model, data_dict, init_dict, used_in_fit = velociraptor.prepare_model(**sources)
+model, data_dict, init_dict, used_in_fit = velociraptor.prepare_model(
+    S=1e4, **sources)
+
+init_dict = dict([
+    ('theta', 0.15167079461165178),
+    #('mu_coefficient', 0.3),
+    #('mu_coefficients', np.array([1e12, 1])),
+    ('mu_coeff_0', 0.3),
+    ('mu_coeff_1', 1e-4)
+    ('mu_coeff_2', 1e12),
+    ('mu_coeff_3', 1),
+    ('mu_coeff_4', 1)
+    #('sigma_coefficients', np.array([0.3, 4e11, 1]))
+    ('sigma_coeff_0', 0.3),
+    ('sigma_coeff_1', 1e-4),
+    ('sigma_coeff_2', 4e11),
+    ('sigma_coeff_3', 1),
+    ('sigma_coeff_4', 1),
+])
+
 
 p_opt = model.optimizing(data=data_dict, init=init_dict)
 
-samples = model.sampling(**velociraptor.stan.sampling_kwds(
-    data=data_dict, chains=2, iter=2000, init=p_opt))
+print(p_opt)
+
+fig, ax = plt.subplots()
+ax.scatter(sources["phot_rp_mean_flux"][used_in_fit], data_dict["rv_variance"],
+    s=1, alpha=0.5, facecolor="k")
+
+x = np.logspace(
+    np.log10(np.nanmin(sources["phot_rp_mean_flux"])),
+    np.log10(np.nanmax(sources["phot_rp_mean_flux"])),
+    1000)
+bp_rp = np.nanmean(sources["bp_rp"]) * np.ones(x.size)
+mu_opt = np.dot(
+    np.hstack([p_opt["mu_coeff_0"], p_opt["mu_coeff_1"], p_opt["mu_coeff_2"], p_opt["mu_coeff_3"], p_opt["mu_coeff_4"]]),
+    velociraptor._rvf_design_matrix(x, bp_rp=bp_rp))
+sigma_opt = np.dot(
+    p_opt["sigma_coefficients"],
[…]

### 2018-05-20 velociraptor.py

@@ -40,7 +40,7 @@ def load_gaia_sources(path, **kwargs):
 
 
 def prepare_model(rv_single_epoch_variance, target=0.01, model_path="model.stan",
-    S=None, **source_params):
+    S=None, max_rv_variance=None, **source_params):
     """
     Compile the Stan model, and prepare the data and initialisation dictionaries
     for optimization and sampling.
@@ -59,6 +59,10 @@ def prepare_model(rv_single_epoch_variance, target=0.01, model_path="model.stan"
         If not `None`, draw a random `S` valid sources from the data rather than
         using the full data set.
 
+    :param max_rv_variance: [optional]
+        Optionally set a maximum radial velocity variance. Sources with variance
+        larger than this will be excluded.
+
     :Keyword Arguments:
         * *source_params* (``dict``) These are passed directly to the
         `_rvf_design_matrix`, so they should include all of the source labels
@@ -75,6 +79,9 @@ def prepare_model(rv_single_epoch_variance, target=0.01, model_path="model.stan"
     dm = _rvf_design_matrix(**source_params)
     finite = np.all(np.isfinite(dm), axis=0) \
            * np.isfinite(rv_single_epoch_variance)
+    if max_rv_variance is not None:
+        finite *= (rv_single_epoch_variance <= max_rv_variance)
+
     indices = np.where(finite)[0]
     N = sum(finite)
 
@@ -84,8 +91,14 @@ def prepare_model(rv_single_epoch_variance, target=0.01, model_path="model.stan"
     if (S is not None and S > N) or not all(finite):
         logger.warn("Excluding non-finite entries in design matrix! "\
                     "Number of data points: {0}".format(indices.size))
-    
+
+
     dm = dm[:, indices]
+
+    #offsets = np.min()
+    #scales = np.ptp(dm, axis=1)
+
+
     coeff = _rvf_initial_coefficients(dm, target=target)
 
     init = dict(theta=0.1, mu_coefficients=coeff, sigma_coefficients=coeff)
@@ -124,7 +137,7 @@ def predict_map_rv_single_epoch_variance(samples, **source_params):
     """
     Predict the maximum a-posteriori estimate of the radial velocity variance
     from a single epoch.
-    
+
     :param samples:
         The Stan chains from the model.
     """
@@ -134,7 +147,7 @@ def predict_map_rv_single_epoch_variance(samples, **source_params):
     dm = _rvf_design_matrix(**source_params)
[…]

### 2018-05-21 validation.py
finally works -- fucking arrays

@@ -14,23 +14,12 @@ sources = velociraptor.load_gaia_sources("data/rv-cal-subset.fits.gz")
 model, data_dict, init_dict, used_in_fit = velociraptor.prepare_model(
     S=1e4, **sources)
 
+# This *works*, but you could start from any random position.
 init_dict = dict([
-    ('theta', 0.15167079461165178),
-    #('mu_coefficient', 0.3),
-    #('mu_coefficients', np.array([1e12, 1])),
-    ('mu_coeff_0', 0.3),
-    ('mu_coeff_1', 1e-4)
-    ('mu_coeff_2', 1e12),
-    ('mu_coeff_3', 1),
-    ('mu_coeff_4', 1)
-    #('sigma_coefficients', np.array([0.3, 4e11, 1]))
-    ('sigma_coeff_0', 0.3),
-    ('sigma_coeff_1', 1e-4),
-    ('sigma_coeff_2', 4e11),
-    ('sigma_coeff_3', 1),
-    ('sigma_coeff_4', 1),
+    ('theta', 0.20),
+    ('mu_coefficients', np.array([0.3, 1e-4, 1e12, 1, 1])),
+    ('sigma_coefficients', np.array([0.3, 1e-4, 4e11, 1, 1])),
 ])
-
 
 p_opt = model.optimizing(data=data_dict, init=init_dict)
 
@@ -46,7 +35,7 @@ x = np.logspace(
     1000)
 bp_rp = np.nanmean(sources["bp_rp"]) * np.ones(x.size)
 mu_opt = np.dot(
-    np.hstack([p_opt["mu_coeff_0"], p_opt["mu_coeff_1"], p_opt["mu_coeff_2"], p_opt["mu_coeff_3"], p_opt["mu_coeff_4"]]),
+    p_opt["mu_coefficients"],
     velociraptor._rvf_design_matrix(x, bp_rp=bp_rp))
 sigma_opt = np.dot(
     p_opt["sigma_coefficients"],
@@ -54,12 +43,12 @@ sigma_opt = np.dot(
 
 
 mu_init = np.dot(
-    np.hstack([init_dict["mu_coeff_0"], init_dict["mu_coeff_1"], init_dict["mu_coeff_2"], p_opt["mu_coeff_3"], p_opt["mu_coeff_4"]]),
+    init_dict["mu_coefficients"],
     velociraptor._rvf_design_matrix(x, bp_rp=bp_rp))
 
 
 sigma_init = np.dot(
[…]

### 2018-05-28 article/ms.tex
words

@@
 \newcommand\mh{[\rm{M}/\rm{H}]}
+\newcommand{\luminosity}{L_\circ}
+\newcommand{\radius}{R_\circ}
 
@@
 \newcommand{\todo}[1]{\textcolor{red}{#1}}
+\newcommand{\rp}{\textsl{rp}}
+\newcommand{\bp}{\textsl{bp}}
+
+\newcommand{\NumberOfHigherOrderSystems}{XXX}
+\newcommand{\NumberOfSingleLinedSystems}{XXX}
 
@@
 
-%\submitjournal{AAS Journals}
+\submitjournal{AAS Journals}
 
 \shorttitle{Stellar multiplicity}
-\shortauthors{Casey et al.?}
+\shortauthors{Casey et al.}
 
@@
 
-%\author{Daniel Foreman-Mackey}
-%\affiliation{Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
+\author{Daniel Foreman-Mackey}
+\affiliation{Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
 
@@
 \begin{abstract}
-Most stars are in binary systems. 
-The occurrence rate of binary star systems is critically important to many 
-fields of astrophysics, but the effects of binaries are commonly ignored in 
-astronomical data analysis applications.
-Here we construct a model to probabilistically identify binary stars based on
-excess radial velocity scatter, photometry, and astrometric excess noise.
-
-We identify binaries \todo{within some orbital and SP limits} and show that 
-the orbital semi-amplitude is well-approximated by the excess radial velocity
-scatter for systems within these criteria.
-
-\todo{We make a few interesting tests about binarity...}
+Most stars are in binary systems. The occurrence rate of multiple order systems 
+(e.g., binaries, triples) is relevant to many fields of astrophysics, but the 
+effects of stellar multiplicity are commonly ignored in the analysis of most 
+stellar and galactic surveys. Here we use \Gaia\ data to identify 
+\todo{\NumberOfHigherOrderSystems} higher-order (e.g., binary) systems based on
+radial velocity or astrometric jitter, photometry, and for a subset of systems, 
+we show that the \emph{lack} of reported radial velocities can be a reliable 
+indicator of double-lined spectroscopic binary systems. Our sample includes
+\todo{\NumberOfSingleLinedSystems} unresolved binary systems detected solely
+from radial velocity or astrometric jitter, which have orbital periods up to 
+$\approx$3.5\,yrs. We show that the ratio of jitter in radial velocity and
+astrometry can be calibrated to estimate inclination angles (and companion masses)
+for thousands of systems. \todo{We comment on stellar multiplicity with stellar
+metallicity}. \todo{We comment on systems with stellar remnants (e.g. NS/BHs).}
 \end{abstract}
@@
 
-There are at least three ways that binary systems can be reliably
-identified using from \Gaia data: through radial velocity variations, 
-anomalous photometric colours, or excess astrometric noise. In this 
-work we do not make use of astrometric noise because of the 
-degree-of-freedom astrometry bug present in the second \Gaia\ data 
-release \citep{Lindegren:2018}. We will describe the radial velocity
-information that forms part of our model for identifying binary stars,
-and then discuss the photometric information used, before explicitly
-outlining the assumptions of our model.
+There are many ways that binaries can be reliably identified using \Gaia\
+data, including radial velocity measurements, photometry, and astrometry.
+Here we describe the radial velocity information that forms part of our 
+model before describing how photometry and astrometry are incorporated into our
+model.
 
-The second \Gaia\ data release provides median radial velocity 
-estimates for 7,224,631 stars. The radial velocity error provided
-($\GaiaRVE$; \texttt{radial\_velocity\_error}) is a function of the 
-number of transits $N$ (i.e., observations; column 
-\texttt{rv\_nb\_transits}) and the scatter among those $N$ measurements. 
-This allows us to calculate the scatter among those measurements, or
-rather the \emph{radial velocity jitter}, for each source
+\subsection{SB1 Systems: Single-lined spectroscopic binaries}
+
+Radial velocity measurements are not available for individual epochs in the 
+second \Gaia\ data release, but the median and error of radial velocity
+measurements are available for 7,224,631 point sources \cite{gaia_rvs}. 
+The radial velocity error provided ($\GaiaRVE$; column \texttt{radial\_velocity\_error}) 
+is a function of the number of transits $N$ (i.e., the number of observations; 
+represented by column \texttt{rv\_nb\_transits}) and the scatter among those $N$ measurements. 
+This allows us to calculate the scatter among those measurements, which we will
+refer to as the \emph{radial velocity jitter}, for each source
 \begin{eqnarray}
@@
 For a single star system observed $N$ times, the source radial velocity
-jitter represents either the minimum uncertainty with which
-\Gaia\ can measure radial velocity for a star of that apparent 
-brightness and colour. However for stars with stellar companions,
-the radial velocity scatter is a function of orbital parameters 
-and observing epochs. Thus we can reasonably assume that the
-radial velocity jitter will be a mixture of
-two populations: single stars, where the radial velocity jitter
-represents the minimum radial velocity noise that \Gaia\ can
-measure for a star of that type, and a binary population that
-could have radial velocity variances well above the noise floor.
+jitter represents either the minimum uncertainty with which \Gaia\ can measure 
+radial velocity for a star of that apparent brightness and colour. However for 
+stars with stellar companions and no evidence of double lines or binarity in the
+cross-correlation-function, so-called SB1 type systems \citep{someone}, the
+radial velocity jitter is a function of the orbital parameters and observing
+epochs. We assume that the distribution of radial velocity jitter in the second
+\Gaia\ data release is a mixture of two populations:
+
+\begin{enumerate}
+\item \emph{single-star systems}, where the radial velocity jitter represents the
+      minimum radial velocity noise that \Gaia\ can measure for a star of that
+      apparent magnitude and colour; and
+
+\item \emph{higher-order star systems}, where the radial velocity jitter is
+      significantly above that for single-star systems of a similar apparent
+      magnitude and colour.
+\end{enumerate}
+
+We make a number of other explicit assumptions about the radial velocity
+measurements reported by \Gaia:
+
+\begin{enumerate}
+    \item the RV error is or is not a function of colour
+    \item the template used for RV determination does or does not affect the 
+          error floor
+    \item the jitter in radial velocity measurements for a single star will
+          increase as a function of apparent flux, etc.
+    \item the spread in the jitter will also increase as a function of these
+          properties
+\end{enumerate}
+
+Given these assumptions, the model we adopt for the radial velocity jitter is
+\begin{eqnarray}
+    MODEL
+\end{eqnarray}
+
+\noindent{}where \todo{X is Y} abd .
+
+In principle for well-sampled circular orbits this model would allow us to
+identify binary star systems with orbital periods up to twice the observing
+span in the second \Gaia\ data release ($\approx$21 months observing span; 
+$\approx$44 month orbital periods). For higher-order star systems with long 
+periods, or with mass ratios that result in a small velocity semi-amplitude $K$, 
+under our assumptions these systems may be classified as single-star systems 
+because the intrinsic radial velocity variation is within the expected 
+variations for single-stars of a similar apparent magnitude and colour. We 
+revisit this issue in Section \ref{sec:sb_limits}, where we explicitly define 
+the probability that a \Gaia\ source is a higher-order star system $p(y|\textrm{SB1})$ 
+within defined limits of orbital period (and other orbital parameters). Outside of this
+parameter range, we are insensitive to distinguishing single-star systems from
+higher-order SB1-type systems.
+
+\begin{itemize}
+    \item show jitter compared to RP flux, etc
+    \item how the model was fit
+    \item show results of the fit
+    \item what things did we \emph{not} include in the model (eg difference between temperature and the temperature of the template used), and why
+    \item results
+\end{itemize}
+
+\subsection{SB2 Systems: Double-lined spectroscopic binaries}
+
+Many sources in the second \Gaia\ data release do not have any radial velocity
+measurements reported, despite being bright enough ($G \lesssim 13$) and in a
+suitable temperature range for radial velocities to be measured. The principle
+reason why radial velocity measurements are \emph{not} reported for these stars
+is because the \Gaia/DPAC team have identified the source to be a double-lined
+spectroscopic binary (a SB2-type system), in which case it is not sensible to 
+report a single radial velocity measurement. For fainter or bluer sources, 
+however, the radial velocity may not be reported simply because it is too faint, 
+blue, or red.
+
+We calculated the completeness of radial velocity measurements (e.g., the
+fraction of sources with reported radial velocities) as a function of all 
+available source properties that might affect whether the radial velocity may
+be reported or not. This included position ($\alpha$, $\delta$, $l$, $b$),
+parallax, proper motions, apparent magnitudes, colours, properties of the radial
+velocity templates, stellar parameters ($\teff$, $\radius$, $\luminosity$),
+and other properties. The full list of properties investigated is described in
+Appendix \todo{X}. In Figure \ref{fig:radial_velocty_completeness} we show the
+completeness as a function of just some of these properties, which demonstrate 
+that the radial velocity completeness is relatively flat until a star becomes 
+too faint (low \rp\ flux), or is either too blue or too red. We adopt 
+conservative limits for when the radial velocity completeness starts to drop 
+with these properties, and we assume that any source within the following
+parameter range
+
+\begin{eqnarray}
+    \todo{X} =& \textrm{phot rp mean flux}  = &\todo{X} \\
+    \todo{X} =& \textrm{bp}-\textrm{rp}    = &\todo{X}
+\end{eqnarray}
+\noindent{}is likely to be a double-lined spectroscopic binary (SB2) if no 
+radial velocity is reported.
+
+\begin{itemize}
+    \item completeness
+    \item cross-check against SB catalogues -- does it match expectations?
+\end{itemize}
+
+\subsection{Unresolved equal mass binaries: photometry}
+
+\subsection{Astrometric binaries}
+
+Many higher-order star systems are unresolved point sources in \Gaia, but the
+stellar companion(s) have sufficient mass to measurably affect the tangential 
+velocity of the primary and produce a jitter in astrometric positions. There are
+many other sources of astrometric jitter, but here we show that the principle
+effect of astrometric excess noise is due to stellar companions. Note that due
+to the astrometry `degree of freedom bug' in the processing of the second \Gaia\
+data release \citep[e.g., see Appendix A1 of ][]{Lindegren:2018}, we only make
+use of astrometric excess noise from the \todo{Tycho-Gaia Astrometric Solution}
+\citep{Michalik, TGAS, etc}.
+
+\begin{itemize}
+    \item when would we expect AEN? what stellar or orbital properties does it 
+          correlate or grow with?
+    \item calibration model for AEN in DR1
+\end{itemize}
 

### 2018-05-28 article/ms.tex
words

@@
 
-\title{Stellar multiplicity in \Gaia}
+\title{Detection of millions of stellar binaries using \Gaia\ data}
 
@@
 
-%\author[0000-0003-2866-9403]{David W. Hogg??}
+%\author[0000-0003-2866-9403]{David W. Hogg}
 %\affiliation{Center for Cosmology and Particle Physics, Department of Physics,
@@
 \begin{itemize}
-\item the binary fraciton is important
+\item the binary fraction is important
 \item Gaia provides exquisite astrometry and RV measurements over many years
@@
 \item the excess RV noise (and in DR1 or DR3, excess astrometric noise) can be used to infer the presence of a companion
+\item here we do that for lots of sources
 \end{itemize}
@@
 data, including radial velocity measurements, photometry, and astrometry.
+These methods are sensitive to binary systems with different properties.
 Here we describe the radial velocity information that forms part of our 
@@
 
+We caution that this deductive inference is potentially far more uncertain (or
+contaminated) than the binary probabilities we calculate from the radial
+velocity jitter, astrometric excess noise, and photometric colours. Here we are
+explicitly assuming that if a source is bright enough and within the right 
+colour range, then a radial velocity would be provided \emph{unless} the source
+is a spectroscopic binary. In Section \ref{sec:discussion} we detail some of the
+immediate limitations or interpretations of this assumption, but here we simply
+note that the probability of a source being a double-lined spectroscopic binary
+is calculated (or rather, deduced) separately to our calculations of binarity
+probability given \emph{measured} quantities like radial velocity, astrometry,
+and photometric colours.
+
 \begin{itemize}
@@
     \item cross-check against SB catalogues -- does it match expectations?
+    \item what about close to the plane? 
+    \item when and why is this bad?
 \end{itemize}
@@
 
-% we assume what?
-%
-
-% given these assumptions, the model we adopt is XYZ
-% we chose to model in RP flux as it more closely represents the flux counts that would be received by the Gaia RVS, rather than G or BP flux.
-
-% the model we use is
-
-%
-
 \begin{itemize}
-\item The effect of binaries.
-
-\item RV noise vs astrometric excess noise.
-
-\item When would we expect to see astrometric noise and not RV noise, and vice-versa? in what situations?
-
-\item Here we base our model using RV data only from Gaia DR2.
-
-\item Some fraction of stars in Gaia are single star systems (or have such low semi-amplitudes that would not be detected by Gaia), and we will use those stars (within a mixture model) to calibrate the radial velocity error floor
-
-\item We will do this on individual epoch measurements
-
-\item ASSUMPTIONS
-
-\item MODEL
 
@@
 \item Binarity with metallicity 
-
-\item Binarity at height above the plane?
-
-\item Binarity in clusters vs the plane?
 \end{itemize}
@@
 
+\subsection{Comparison with \citet{El_Badry:2017,El_Badry:2018}}
+
+\subsection{Comparison with \citet{Badenes:2017}}
+
+\subsection{Comparison with \citet{Raghavan:2010}}
+
+\subsection{Comparison with \citet{Troup:2010}}
+
+\subsection{Comparison with \citet{Price-Whelan:2010}}
+
+\begin{itemize}
+    \item \todo{binarity with metallicity}
+    \item \todo{binarity across the H-R diagram}
+    \item \todo{binarity in clusters vs the field?}
+    \item \todo{binarity among extremely metal-poor stars?}
+\end{itemize}
+
 \acknowledgements
@@
 It is a pleasure to thank
+    Adrian Price-Whelan,
 	David W. Hogg,
@@
 
-This work has made use of CosmoHub.
-CosmoHub has been developed by the Port d'Informaci\'o Cient\'ifica (PIC), maintained through a collaboration of the Institut de F\'isica d'Altes Energies (IFAE) and the Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), and was partially funded by the "Plan Estatal de Investigaci\'on Científica y T\'ecnica y de Innovaci\'on" program of the Spanish government.
+This work has made use of CosmoHub. CosmoHub has been developed by the Port 
+d'Informaci\'o Cient\'ifica (PIC), maintained through a collaboration of the 
+Institut de F\'isica d'Altes Energies (IFAE) and the Centro de Investigaciones 
+Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), and was partially 
+funded by the ``Plan Estatal de Investigaci\'on Científica y T\'ecnica y de 
+Innovaci\'on'' program of the Spanish government.
 

### 2018-05-28 tensorflow_model.py
add tensorflow model

@@ -1,0 +1,67 @@
+
+import numpy as np
+import tensorflow as tf
+import matplotlib.pyplot as plt
+import velociraptor
+
+
+sources = velociraptor.load_gaia_sources("data/rv-all.fits")
+model, data_dict, init_dict, idx = velociraptor.prepare_model(S=1e4, **sources)
+
+N, M = (data_dict["N"], data_dict["M"])
+init_dict.update(
+    mu_coefficients=[2.1465e-05, 1.4585e+02, 2.0827e+11, 7.7332e-08, 5.8626e+00],
+    sigma_coefficients=[7.4405e-04, 3.9762e-01, 1.7293e+11, 4.1103e-04, 5.9489e+00]
+)
+
+theta = tf.clip_by_value(tf.Variable(0.5, dtype=tf.float64), 0, 1)
+mu_coefficients = tf.clip_by_value(
+    tf.Variable(np.atleast_2d(init_dict["mu_coefficients"]), dtype=tf.float64),
+    0, np.inf)
+sigma_coefficients = tf.clip_by_value(
+    tf.Variable(np.atleast_2d(init_dict["sigma_coefficients"]), dtype=tf.float64),
+    0, np.inf)
+
+mu = tf.matmul(mu_coefficients, data_dict["design_matrix"].T)
+ivar = tf.matmul(sigma_coefficients, data_dict["design_matrix"].T)**-2
+
+log_p_s = tf.log(1.0 - theta) \
+        - 0.5 * np.log(2 * np.pi) + 0.5 * tf.log(ivar) \
+        - 0.5 * (data_dict["rv_variance"] - mu)**2 * ivar
+
+log_p_b = tf.log(theta) \
+        - np.log(np.max(data_dict["rv_variance"])) * np.ones((1, data_dict["N"]))
+
+log_prob = tf.reduce_sum(tf.reduce_logsumexp(tf.concat([log_p_s, log_p_b], 0), 0))
+
+tolerance = 1e-6
+learning_rate = 1e-3
+max_iterations = 100000
+
+training_step = tf.train.AdamOptimizer(learning_rate).minimize(-log_prob)
+
+cost_history = np.empty(shape=[1],dtype=float)
+
+with tf.Session() as session:
+    session = tf.Session()
+    session.run(tf.global_variables_initializer())
+
+    print("Initial log prob: {0}".format(session.run(log_prob)))
+    print("Initial gradient: {0}".format(
+        session.run(tf.gradients(log_prob, 
[…]

### 2018-05-28 calculate_rv_completeness.py
calculate RV completeness for labels, and plot  separately

@@ -1,0 +1,142 @@
+
+"""
+Calculate the completeness of radial velocity measurements as a function of
+various properties so that we can gauge when we should use the *lack* of radial
+velocity as an indicator of binarity.
+"""
+
+import pickle
+import numpy as np
+from astropy.io import fits
+from collections import Counter
+
+N_bins = 50
+
+semilogx = [True, False]
+equidensity = [True, False]
+
+completeness = {
+    "N_bins": N_bins,
+    "semilogx": np.repeat(semilogx, 2),
+    "equidensity": np.tile(equidensity, 2),
+    "completeness": dict()
+}
+
+image = fits.open("data/rv-completeness.fits")
+data = image[1].data
+
+has_rv = np.isfinite(data["radial_velocity"])
+
+label_names = (
+ 'ra',
+ 'dec',
+ 'parallax',
+ 'pmra',
+ 'pmdec',
+ 'astrometric_gof_al',
+ 'astrometric_chi2_al',
+ 'astrometric_weight_al',
+ 'astrometric_pseudo_colour',
+ 'mean_varpi_factor_al',
+ 'astrometric_matched_observations',
+ 'visibility_periods_used',
+ 'astrometric_sigma5d_max',
+ 'phot_g_n_obs',
+ 'phot_g_mean_flux',
+ 'phot_g_mean_mag',
+ 'phot_bp_n_obs',
+ 'phot_bp_mean_flux',
+ 'phot_bp_mean_mag',
+ 'phot_rp_n_obs',
+ 'phot_rp_mean_flux',
+ 'phot_rp_mean_mag',
+ 'phot_bp_rp_excess_factor',
+ 'bp_rp',
+ 'bp_g',
+ 'g_rp',
+ 'radial_velocity',
+ 'radial_velocity_error',
+ 'rv_nb_transits',
+ 'rv_template_teff',
+ 'rv_template_logg',
+ 'rv_template_fe_h',
+ 'phot_variable_flag',
+ 'l',
+ 'b',
+ 'ecl_lon',
+ 'ecl_lat',
+ 'teff_val',
+ 'a_g_val',
+ 'e_bp_min_rp_val',
+ 'radius_val',
+ 'lum_val'
+)
+
+for label_name in label_names:
+
+    print(label_name)
+    if label_name in completeness["completeness"]: continue
+
+    x = image[1].data[label_name]
+    x_min, x_max = (np.nanmin(x), np.nanmax(x))
+
[…]

### 2018-05-28 tensorflow_model.py
move TF to notebook

@@ -1,67 +1,0 @@
-
-import numpy as np
-import tensorflow as tf
-import matplotlib.pyplot as plt
-import velociraptor
-
-
-sources = velociraptor.load_gaia_sources("data/rv-all.fits")
-model, data_dict, init_dict, idx = velociraptor.prepare_model(S=1e4, **sources)
-
-N, M = (data_dict["N"], data_dict["M"])
-init_dict.update(
-    mu_coefficients=[2.1465e-05, 1.4585e+02, 2.0827e+11, 7.7332e-08, 5.8626e+00],
-    sigma_coefficients=[7.4405e-04, 3.9762e-01, 1.7293e+11, 4.1103e-04, 5.9489e+00]
-)
-
-theta = tf.clip_by_value(tf.Variable(0.5, dtype=tf.float64), 0, 1)
-mu_coefficients = tf.clip_by_value(
-    tf.Variable(np.atleast_2d(init_dict["mu_coefficients"]), dtype=tf.float64),
-    0, np.inf)
-sigma_coefficients = tf.clip_by_value(
-    tf.Variable(np.atleast_2d(init_dict["sigma_coefficients"]), dtype=tf.float64),
-    0, np.inf)
-
-mu = tf.matmul(mu_coefficients, data_dict["design_matrix"].T)
-ivar = tf.matmul(sigma_coefficients, data_dict["design_matrix"].T)**-2
-
-log_p_s = tf.log(1.0 - theta) \
-        - 0.5 * np.log(2 * np.pi) + 0.5 * tf.log(ivar) \
-        - 0.5 * (data_dict["rv_variance"] - mu)**2 * ivar
-
-log_p_b = tf.log(theta) \
-        - np.log(np.max(data_dict["rv_variance"])) * np.ones((1, data_dict["N"]))
-
-log_prob = tf.reduce_sum(tf.reduce_logsumexp(tf.concat([log_p_s, log_p_b], 0), 0))
-
-tolerance = 1e-6
-learning_rate = 1e-3
-max_iterations = 100000
-
-training_step = tf.train.AdamOptimizer(learning_rate).minimize(-log_prob)
-
-cost_history = np.empty(shape=[1],dtype=float)
-
-with tf.Session() as session:
-    session = tf.Session()
-    session.run(tf.global_variables_initializer())
-
-    print("Initial log prob: {0}".format(session.run(log_prob)))
-    print("Initial gradient: {0}".format(
-        session.run(tf.gradients(log_prob, 
[…]

### 2018-06-03 article/ms.tex
words

@@
 
-%\graphicspath{../figures}
-
 \submitjournal{AAS Journals}
@@
 
-\title{Detection of millions of stellar binaries using \Gaia\ data}
+\title{Detection and characterisation of stellar multiplicity from \Gaia\ data}
 
@@
 \author[0000-0003-0174-0564]{Andrew R. Casey}
-\affiliation{School of Physics \& Astronomy, Monash University, Clayton 3800, 
-    Victoria, Australia}
-\affiliation{Faculty of Information Technology, Monash University, Clayton 3800, 
-    Victoria, Australia}
+\affiliation{School of Physics \& Astronomy, 
+			 Monash University,
+			 Wellington Rd, Clayton 3800, Victoria, Australia}
+\affiliation{Faculty of Information Technology, 
+			 Monash University, 
+			 Wellington Rd, Clayton 3800, Victoria, Australia}
 
-\author{Daniel Foreman-Mackey}
-\affiliation{Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
+\author[0000-0002-9328-5652]{Daniel Foreman-Mackey}
+\affiliation{Flatiron Institute, 
+			 162 Fifth Ave, New York, NY 10010, USA}
 
-%\author[0000-0003-2866-9403]{David W. Hogg}
-%\affiliation{Center for Cosmology and Particle Physics, Department of Physics,
-%    New York University, 726 Broadway, New York, NY 10003, USA}
-%\affiliation{Center for Data Science, New York University, 60 Fifth Ave, 
-%    New York, NY 10011, USA}
-%\affiliation{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 
-%    Heidelberg}
-%\affiliation{Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
+\author[0000-0003-2866-9403]{David W. Hogg}
+\affiliation{Center for Cosmology and Particle Physics, Department of Physics,
+    		 New York University, 
+		  	 726 Broadway, New York, NY 10003, USA}
+\affiliation{Center for Data Science, 
+			 New York University, 60 Fifth Ave, New York, NY 10011, USA}
+\affiliation{Max-Planck-Institut f\"ur Astronomie, 
+			 K\"onigstuhl 17, D-69117 Heidelberg}
+\affiliation{Flatiron Institute,
+			 162 Fifth Ave, New York, NY 10010, USA}
 
 \begin{abstract}
-Most stars are in binary systems. The occurrence rate of multiple order systems 
-(e.g., binaries, triples) is relevant to many fields of astrophysics, but the 
-effects of stellar multiplicity are commonly ignored in the analysis of most 
-stellar and galactic surveys. Here we use \Gaia\ data to identify 
-\todo{\NumberOfHigherOrderSystems} higher-order (e.g., binary) systems based on
-radial velocity or astrometric jitter, photometry, and for a subset of systems, 
-we show that the \emph{lack} of reported radial velocities can be a reliable 
-indicator of double-lined spectroscopic binary systems. Our sample includes
+Most stars are in binary systems. The occurrence rate and properties of stellar 
+multiples (e.g., binaries, triples) is relevant to many fields of astrophysics, 
+but the effects of stellar multiplicity are commonly ignored in the analysis of 
+most stellar and galactic surveys. Here we use \Gaia\ data to identify 
+\todo{\NumberOfHigherOrderSystems} stellar multiple systems from radial velocity
+and astrometric jitter, photometry, and intrinsic luminosity. Our sample includes
 \todo{\NumberOfSingleLinedSystems} unresolved binary systems detected solely
@@
 
-\begin{itemize}
-\item the binary fraction is important
-\item Gaia provides exquisite astrometry and RV measurements over many years
-\item The errors in these quantities are well-characterised
-\item When a star is in a binary system (or triary, etc), there will be excess scatter relative to a single star of the primary's properties
-\item the excess RV noise (and in DR1 or DR3, excess astrometric noise) can be used to infer the presence of a companion
-\item here we do that for lots of sources
-\end{itemize}
+Higher order stellar systems describes any stars that have at least one stellar 
+companion (e.g., binaries, trinaries). The presence of stellar companions
+introduces a number of complications to the inference of stellar and galactic
+properties, but these factors are often ignored.
+
+The \Gaia\ space telescope provides exquisite astrometry, photometry, and radial
+velocity measurements over many years for million of point stars in our galaxy.
+%These data provides an excellent opportunity to self-calibrate the errors in those
+%measurements. 
+When a point source has an unresolved stellar companion within some range of orbital
+and stellar parameters, it is expected that there will be an anomalous excess in 
+these measured properties relative to a single star of a similar type. This is
+nothing new: astronomers have used radial velocity \citep{firstplanet} and 
+astrometric variations \citep{someone} and photometry \citep{someone} to infer the 
+presence of stellar and exoplanet companions forever. However, the opportunity
+presented by \Gaia\ is to infer the presence of stellar companions for millions
+of point sources, across both hemispheres. No ground- or space-based instrument
+has ever provided such an opportunity.
+
+In this work we make use of photometry and astrometry (including radial motion)
+measurements in the second \Gaia\ data release to infer the presence of stellar
+companions for millions of point sources, and provide some basic characterisations
+of these systems. In Section \ref{sec:method} we describe our methods, and in
+Section \ref{sec:results} we outline our results in context of other stellar
+multiplicity surveys. We include comparisons of binary properties between
+clusters and the field, as a function of stellar properties (e.g., stellar
+metallicity), and highlight some particularly noteworthy systems that would
+benefit from immediate follow-up spectroscopic observations.
 
@@
 data, including radial velocity measurements, photometry, and astrometry.
-These methods are sensitive to binary systems with different properties.
-Here we describe the radial velocity information that forms part of our 
-model before describing how photometry and astrometry are incorporated into our
-model.
+These detection methods are sensitive to binary systems with different 
+properties. Here we describe the radial velocity information that forms 
+part of our model before describing how photometry and astrometry are 
+incorporated into the detection and characterisation of stellar multiples.
 
 \subsection{SB1 Systems: Single-lined spectroscopic binaries}
+\label{sec:sb1_methods}
 
 Radial velocity measurements are not available for individual epochs in the 
-second \Gaia\ data release, but the median and error of radial velocity
-measurements are available for 7,224,631 point sources \cite{gaia_rvs}. 
-The radial velocity error provided ($\GaiaRVE$; column \texttt{radial\_velocity\_error}) 
+second \Gaia\ data release \citep{Lindegren:2018,Gaia_RVS}, but the a radial velocity point estimate and error
+is available for 7,224,631 sources \cite{gaia_rvs}. 
+The radial velocity error provided by \Gaia\ ($\GaiaRVE$; column name \texttt{radial\_velocity\_error}) 
 is a function of the number of transits $N$ (i.e., the number of observations; 
@@
 
-For a single star system observed $N$ times, the source radial velocity
-jitter represents either the minimum uncertainty with which \Gaia\ can measure 
+For a single star (without a significant mass companion) the source radial velocity
+jitter represents the minimum uncertainty with which \Gaia\ can measure 
 radial velocity for a star of that apparent brightness and colour. However for 
@@
 cross-correlation-function, so-called SB1 type systems \citep{someone}, the
-radial velocity jitter is a function of the orbital parameters and observing
+radial velocity jitter is a function of the orbital parameters and observation
 epochs. We assume that the distribution of radial velocity jitter in the second
@@
 
-\begin{enumerate}
-    \item the RV error is or is not a function of colour
-    \item the template used for RV determination does or does not affect the 
-          error floor
-    \item the jitter in radial velocity measurements for a single star will
-          increase as a function of apparent flux, etc.
-    \item the spread in the jitter will also increase as a function of these
-          properties
-\end{enumerate}
+\begin{itemize}
+	\item the radial velocity jitter is a function of the apparent brightness 
+		  of the point source
+	\item the radial velocity jitter is a function of the apparent $bp-rp$
+		  colour of the point source
+	\item there is some intrinsic spread to the radial velocity jitter of
+		  single stars, and that intrinsic spread also increases with the 
+		  apparent brightness and colour of the source
+	\item the difference between the true surface gravity $\logg$ and the
+		  surface gravity of the radial velocity template used
+		  (column \texttt{rv\_template\_logg}) has no impact on the radial
+		  velocity jitter
+	\item the difference between the true stellar metallicity [Fe/H] and
+	 	  the metallicity of the radial velocity template used
+		  (column \texttt{rv\_template\_feh}) has no impact on the radial
+		  velocity jitter
+	\item \todo{state something about teff template}		  
+\end{itemize}
 
-Given these assumptions, the model we adopt for the radial velocity jitter is
+Conditioned on these assumptions, we adopt a two-component mixture model for 
+the radial velocity jitter: a normal distribution where the mean and variance
+are a function of stellar properties, and a uniform background component to
+account for stellar multiples. Specifically, the likelihood is
+
 \begin{eqnarray}
+%	\mathcal{L} & \equiv & p\left(\{y
     MODEL
@@
 \subsection{SB2 Systems: Double-lined spectroscopic binaries}
+\label{sec:sb2_methods}
 
-Many sources in the second \Gaia\ data release do not have any radial velocity
-measurements reported, despite being bright enough ($G \lesssim 13$) and in a
-suitable temperature range for radial velocities to be measured. The principle
-reason why radial velocity measurements are \emph{not} reported for these stars
-is because the \Gaia/DPAC team have identified the source to be a double-lined
-spectroscopic binary (a SB2-type system), in which case it is not sensible to 
-report a single radial velocity measurement. For fainter or bluer sources, 
-however, the radial velocity may not be reported simply because it is too faint, 
-blue, or red.
+Many sources in the second \Gaia\ data release do not have any reported radial 
+velocity measurement, despite being bright enough ($G \lesssim 13$) and in a
+suitable temperature range (e.g., between $X$ and $Y$) for radial velocities to 
+be measured \citep{someone}. The principle reason why radial velocity measurements 
+are \emph{not} reported for these stars is because the \Gaia/DPAC team have 
+identified the source to be a double-lined spectroscopic binary (a SB2-type system), 
+either through a composition of two stellar sources present in the spectra, or
+from multiple (significant) modes in the cross-correlation function. In these
+situations it is not sensible to report a point estimate of the radial velocity
+of the point source. For fainter or bluer sources, however, the radial velocity 
+may not be reported simply because it is too faint, blue, or red.
 
@@
 and other properties. The full list of properties investigated is described in
-Appendix \todo{X}. In Figure \ref{fig:radial_velocty_completeness} we show the
-completeness as a function of just some of these properties, which demonstrate 
+Appendix \ref{appendix:sb2}. In Figure \ref{fig:radial_velocity_completeness} 
+we show the completeness as a function of some pertinent properties, which demonstrate 
 that the radial velocity completeness is relatively flat until a star becomes 
@@
 \begin{eqnarray}
-    \todo{X} =& \textrm{phot rp mean flux}  = &\todo{X} \\
-    \todo{X} =& \textrm{bp}-\textrm{rp}    = &\todo{X}
+    \mathrm{phot\,\,rp\,\,mean\,\,flux} & \in & [10^{4}, 10^8] \nonumber \\
+    \mathrm{bp\,\,-rp} & \in & [X, Y] 
+    \label{eq:sb2_criteria}
 \end{eqnarray}
 \noindent{}is likely to be a double-lined spectroscopic binary (SB2) if no 
-radial velocity is reported.
+radial velocity is reported. That is to say that we are explicitly assuming
+that if a point source meets the criteria in Equation \ref{eq:sb2_criteria} 
+and does not have a radial velocity measurement, then the point source is
+an unresolved double-lined spectroscopic binary. In principle we could
+make more realistic attempts to model the radial velocity completeness as
+a function of stellar properties rather than simply stating ``sources within
+this parameter range should have radial velocities unless they are SB2 systems'',
+but the radial velocity completeness within our specified range is reasonably
+flat.
 
-We caution that this deductive inference is potentially far more uncertain (or
-contaminated) than the binary probabilities we calculate from the radial
-velocity jitter, astrometric excess noise, and photometric colours. Here we are
-explicitly assuming that if a source is bright enough and within the right 
-colour range, then a radial velocity would be provided \emph{unless} the source
-is a spectroscopic binary. In Section \ref{sec:discussion} we detail some of the
-immediate limitations or interpretations of this assumption, but here we simply
-note that the probability of a source being a double-lined spectroscopic binary
-is calculated (or rather, deduced) separately to our calculations of binarity
-probability given \emph{measured} quantities like radial velocity, astrometry,
-and photometric colours.
+In Section \ref{sec:discussion} we detail some of the immediate limitations or 
+interpretations of this assumption, but here we  note that the probability of a 
+source being a double-lined spectroscopic binary is calculated (or rather, 
+deduced) separately to our calculations of binarity probability given 
+\emph{measured} quantities like radial velocity, astrometry, and photometric 
+colours. And as we discuss in Section \ref{sec:discussion}, the main conclusions
+of this work do not depend on SB2 systems. Nevertheless, we \emph{strongly} 
+caution that our deductive inference of SB2 systems will be far more uncertain 
+than the binary probabilities we derive from other information. Our indicator 
+whether a star is likely a SB2 system likely suffers from more contamination
+and lower completeness compared to the probabilities we derive from other information
+(e.g., the radial velocity jitter, astrometric excess noise, and photometric 
+colours), and the contamination and completeness likely vary over some
+complex (unknown) function of parameter space.
 
-\begin{itemize}
-    \item completeness
-    \item cross-check against SB catalogues -- does it match expectations?
-    \item what about close to the plane? 
-    \item when and why is this bad?
-\end{itemize}
+% Figure: radial velocity completeness as a function of pertinent properties
+% 		  with the range that we have said is "fully complete"
+
+% Figure: h-r diagram of SB2 binary fraction
+
+% Figure: cross-match against a catalog of SB2 systems,... do we find all of
+% 		  what they find from SB2 deductive inference alone? or do we miss
+% 		  some SB2s in some parameter range?
+
+% Figure: radial velocity fraction as a function of galactic latitude and
+% 		  longitude
 

### 2018-06-05 article/ms.tex
whoa

@@
 
-\title{Detection and characterisation of stellar multiplicity from \Gaia\ data}
+\title{Detection and characterisation of stellar multiplicity with \Gaia}
 
@@
 \begin{abstract}
-Most stars are in binary systems. The occurrence rate and properties of stellar 
+Most stars have stellar companions. The occurrence rate and properties of stellar 
 multiples (e.g., binaries, triples) is relevant to many fields of astrophysics, 
@@
 In principle for well-sampled circular orbits this model would allow us to
-identify binary star systems with orbital periods up to twice the observing
+identify binary star systems with orbital periods at least up to twice the observing
 span in the second \Gaia\ data release ($\approx$21 months observing span; 
-$\approx$44 month orbital periods). For higher-order star systems with long 
-periods, or with mass ratios that result in a small velocity semi-amplitude $K$, 
-under our assumptions these systems may be classified as single-star systems 
-because the intrinsic radial velocity variation is within the expected 
-variations for single-stars of a similar apparent magnitude and colour. We 
+allowing for well-characterised detections for systems with orbital periods up to $\approx$44 months).
+
+For higher-order star systems with longer periods, we can still find that a
+stellar multiple system is more likely than a single star scenario, but our
+estimate of the orbital properties (e.g., the semi-amplitude $K$) will be biased
+to lower values because we have not fully sampled half of the orbital period.
+At even longer orbital periods, or for stellar multiples with mass ratios that
+produce a small velocity semi-amplitude $K$, under our assumptions these systems 
+may be classified as single-star systems because the intrinsic radial velocity variation is within the expected 
+variations for single stars of a similar apparent magnitude and colour. We 
 revisit this issue in Section \ref{sec:sb_limits}, where we explicitly define 
-the probability that a \Gaia\ source is a higher-order star system $p(y|\textrm{SB1})$ 
-within defined limits of orbital period (and other orbital parameters). Outside of this
-parameter range, we are insensitive to distinguishing single-star systems from
-higher-order SB1-type systems.
+the probability that a \Gaia\ source is a SB1-type system $p(\textrm{SB1}|y)$ 
+within defined limits of orbital period (and other orbital and source parameters). 
+Outside of this parameter range, we are insensitive to distinguishing single-star 
+systems from higher-order SB1-type systems.
 
-\begin{itemize}
-    \item show jitter compared to RP flux, etc
-    \item how the model was fit
-    \item show results of the fit
-    \item what things did we \emph{not} include in the model (eg difference between temperature and the temperature of the template used), and why
-    \item results
-\end{itemize}
+% Model fitting.
+
+% Figure: fitting the radial velocity variance as a function of stellar properties.
+
+% Figures: Binarity as a function of source properties.
+
+% Figure: h-r diagram of SB1 binary fraction
+
+% Figure: histogram of binaries as a fucntion of the Rv excess
 
@@
 parameter range
-
 \begin{eqnarray}
     \mathrm{phot\,\,rp\,\,mean\,\,flux} & \in & [10^{4}, 10^8] \nonumber \\
-    \mathrm{bp\,\,-rp} & \in & [X, Y] 
+    \mathrm{bp-rp} & \in & [X, Y] 
     \label{eq:sb2_criteria}
@@
 
-\subsection{Unresolved equal mass binaries: photometry}
+\subsection{Unresolved near-equal mass binaries: photometry}
+\label{sec:pb_methods}
+
+If a binary system is observed nearly face-on (at an inclination angle 
+$i \approx 0^\circ$) then there is will be no detectable excess radial 
+velocity variations. In principle there may be detectable astrometric
+variations, but most near-equal mass binaries would be detected more 
+reliably through intrinsic magnitudes that are anomalously lower (brighter)
+and bluer colours than what would be expected for a single star system.
+
+\todo{How the fuck am I going to model this, exactly?}
 
@@
 It is a pleasure to thank
-    Adrian Price-Whelan,
-	David W. Hogg,
-	Lorenzo Spina, 
+    Adrian Price-Whelan (Princeton University),
+	Lorenzo Spina (Monash University), 
 		and 
-	Kevin C. Schlaufman.
+	Kevin C. Schlaufman (Johns Hopkins University).
 This work was partly supported through the Australian Research Council 

### 2018-06-05 calculate_rv_completeness.py

@@ -41,13 +41,10 @@ label_names = (
  'astrometric_matched_observations',
  'visibility_periods_used',
  'astrometric_sigma5d_max',
- 'phot_g_n_obs',
  'phot_g_mean_flux',
  'phot_g_mean_mag',
- 'phot_bp_n_obs',
  'phot_bp_mean_flux',
  'phot_bp_mean_mag',
- 'phot_rp_n_obs',
  'phot_rp_mean_flux',
  'phot_rp_mean_mag',
  'phot_bp_rp_excess_factor',
@@ -56,18 +53,15 @@ label_names = (
  'g_rp',
  'radial_velocity',
  'radial_velocity_error',
- 'rv_nb_transits',
  'rv_template_teff',
  'rv_template_logg',
  'rv_template_fe_h',
- 'phot_variable_flag',
  'l',
  'b',
  'ecl_lon',
  'ecl_lat',
  'teff_val',
  'a_g_val',
- 'e_bp_min_rp_val',
  'radius_val',
  'lum_val'
 )
@@ -139,4 +133,7 @@ for label_name in label_names:
     completeness["completeness"][label_name]["x"] = plot_x
     completeness["completeness"][label_name]["y"] = plot_y
     completeness["completeness"][label_name]["yerr"] = plot_y_err
+
+    with open("rv_completeness.pkl", "wb") as fp:
+        pickle.dump(completeness, fp, -1)
 

### 2018-06-05 validation.py

@@ -15,8 +15,103 @@ import velociraptor
 data_path = "data/rv-all.fits"
 sources = velociraptor.load_gaia_sources(data_path)
 
+# Calculate temperature from bp-rp colour and use the distance from the
+# template as a predictor for the error in radial velocity (under the assumption
+# that the wrong template was used, giving a higher velocity error)
+use_in_fit = np.isfinite(sources["radial_velocity"]) \
+           * (sources["phot_bp_rp_excess_factor"] < 1.5) \
+           * np.isfinite(sources["bp_rp"]) \
+           * np.isfinite(sources["teff_val"]) \
+           * (sources["bp_rp"] < 2.5) \
+           * (sources["bp_rp"] > 0.5)
+
+x = sources["bp_rp"][use_in_fit]
+y = sources["teff_val"][use_in_fit]
+
+coeff = np.polyfit(1.0/x, y, 2)
+
+xi = np.linspace(x.min(), x.max(), 100)
+
+fig, ax = plt.subplots()
+ax.scatter(x, y, s=1, alpha=0.05, facecolor="k")
+ax.plot(xi, np.polyval(coeff, 1.0/xi), c='r', lw=2)
+
+
+#sources["approx_teff"] = np.polyval(coeff, 1.0/sources["bp_rp"])
+sources["teff_from_rv_template"] = np.abs(
+    np.polyval(coeff, 1.0/sources["bp_rp"]) - sources["rv_template_teff"])
+
+N_bins = 100
+H_all, xedges, yedges = np.histogram2d(
+    sources["teff_from_rv_template"][use_in_fit],
+    sources["rv_excess_variance"][use_in_fit],
+    bins=(
+        np.linspace(0, 3000, N_bins), 
+        np.linspace(0, 10000, N_bins)
+    ))
+
+
+kwds = dict(
+    aspect=np.ptp(xedges)/np.ptp(yedges), 
+    extent=(xedges[0], xedges[-1], yedges[-1], yedges[0]),
+)
+
+from matplotlib.colors import LogNorm
+fig, ax = plt.subplots(figsize=(7.0, 5.5))
+image = ax.imshow(H_all.T, norm=LogNorm(), cmap="Blues", **kwds)
+cbar = plt.colorbar(image, ax=ax)
+cbar.set_label(r"\textrm{count}")
+
+ax.set_ylim(ax.get_ylim()[::-1])
+
[…]

### 2018-06-05 velociraptor.py

@@ -9,6 +9,7 @@ import matplotlib.pyplot as plt
 from astropy.table import Table
 from matplotlib.ticker import MaxNLocator
 from scipy.stats import norm
+from astropy.io import fits
 
 import stan_utils as stan
 import mpl_utils
@@ -27,6 +28,10 @@ def load_gaia_sources(path, **kwargs):
         The local path to load the sources from.
     """
 
+    #try:
+    #    sources = Table(fits.open(path)[1].data)
+
+    #except:
     sources = Table.read(path, **kwargs)
 
     for band in ("g", "bp", "rp"):
@@ -36,11 +41,31 @@ def load_gaia_sources(path, **kwargs):
 
     sources["rv_single_epoch_variance"] = sources["radial_velocity_error"]**2 \
                                         * sources["rv_nb_transits"] * np.pi/2.0
+
+    # Approximate temperature from bp-rp colour
+    
+    use_in_fit = np.isfinite(sources["radial_velocity"]) \
+               * (sources["phot_bp_rp_excess_factor"] < 1.5) \
+               * np.isfinite(sources["bp_rp"]) \
+               * np.isfinite(sources["teff_val"]) \
+               * (sources["bp_rp"] < 2.5) \
+               * (sources["bp_rp"] > 0.5)
+
+    x = sources["bp_rp"][use_in_fit]
+    y = sources["teff_val"][use_in_fit]
+
+    coeff = np.polyfit(1.0/x, y, 2)
+
+    sources["approx_teff_from_bp_rp"] = np.clip(
+        np.polyval(coeff, 1.0/sources["bp_rp"]),
+        3500, 8000)
+
     return sources
 
 
 def prepare_model(rv_single_epoch_variance, target=0.01, model_path="model.stan",
-    S=None, max_rv_variance=None, **source_params):
+    S=None, design_matrix_function=None, mask=None,
+    **source_params):
     """
     Compile the Stan model, and prepare the data and initialisation dictionaries
     for optimization and sampling.
@@ -59,9 +84,6 @@ def prepare_model(rv_single_epoch_variance, target=0.01, model_path="model.stan"
         If not `None`, draw a random `S` valid sources from the data rather than
         using the full data set.
 
-    :param max_rv_variance: [optional]
-        Optionally set a maximum radial velocity variance. Sources with variance
-        larger than this will be excluded.
 
     :Keyword Arguments:
         * *source_params* (``dict``) These are passed directly to the
[…]

### 2018-06-28 npm_mr_postprocess.py
things

@@ -1,0 +1,82 @@
+"""
+Script to aggregate results together from the non-parametric model.
+"""
+
+import numpy as np
+import os
+import logging
+import pickle
+import yaml
+from astropy.io import fits
+from glob import glob
+
+import npm_utils as npm
+
+with open(npm.CONFIG_PATH, "r") as fp:
+    config = yaml.load(fp)
+
+# Load the data frame so that we can get the right index.
+image = fits.open(config["data_path"])
+data = image[1].data
+
+N = len(data)
+    
+# Figure out what columns we are going to add to the data.
+common_suffix = ""
+common_label_names = ["theta"]
+
+output_label_names = ["{0}{1}".format(cln, common_suffix) for cln in common_label_names]
+
+for kind in ("single", "multiple"):
+    for prefix in ("mu", "sigma"):
+        for predictor_label_name in config["predictor_label_names"]:
+            output_label_names.append("_".join([prefix, kind, predictor_label_name]))
+
+
+K = len(output_label_names)
+L = len(config["predictor_label_names"])
+
+outputs = np.nan * np.ones((N, K), dtype=float)
+
+for index in range(N):
+
+    output_path = npm.get_output_path(data["source_id"][index], config,
+                                      check_path_exists=False)
+    
+    logging.info("({}/{}): Aggregating results from {}".format(
+        index, N, output_path))
+
+    if os.path.lexists(output_path) and not os.path.exists(output_path):
+        raise a
+
+    if not os.path.exists(os.path.realpath(os.path.abspath(output_path))):
+        realpath = os.path.realpath(output_path)
+        logging.info("Skipping because {} does not exist ({} -> {})".format(
+            output_path, output_path, realpath))
+        continue
+
+    with open(output_path, "rb") as fp:
+        result = pickle.load(fp)
+
+    # Parse the output from the result file.
+    p_opt = result["p_opt"]
+    theta = p_opt["theta"]
+
+    offset = 1
+    outputs[index, 0] = theta
+    
+    for k, key in enumerate(("mu_single", "sigma_single", "mu_multiple", "sigma_multiple")):
+        outputs[index, offset + k * L:offset + (k + 1) * L] = p_opt[key]
+    
+
+# Add these columns to the existing data file?
+columns = fits.ColDefs(
+    [c for c in image[1].data.columns] + \
+    [fits.Column(name=pln, format="E", array=outputs[:, i]) for i, pln in enumerate(output_label_names)]
+)
[…]

### 2018-06-28 npm_mr_run.py

@@ -43,7 +43,7 @@ def initialize_from_nearby_point(indices, data, config):
 
 
 
-def optimize_npm_at_point(indices_path, output_path, data, config):
+def optimize_npm_at_point(data_index, indices_path, output_path, data, config):
 
     # Get indices and load data.
     with open(indices_path, "rb") as fp:
@@ -81,11 +81,21 @@ def optimize_npm_at_point(indices_path, output_path, data, config):
 
     model = stan.load_stan_model(config["model_path"])
 
+    S = config.get("share_optimised_result_with_nearest", 0)
+    
     t_init = time()
-    p_opt = model.optimizing(**opt_kwds)
+    try:
+        p_opt = model.optimizing(**opt_kwds)
+    
+    except:
+        logging.exception("Failed to optimize from {}".frmat(indices_path))
+        return (indices, S, None)
+
     t_complete = time() - t_init
 
-    result = dict(p_opt=p_opt, init_from_source_id=init_from_source_id)
+    result = dict(p_opt=p_opt, 
+                  data_index=data_index, 
+                  init_from_source_id=init_from_source_id)
 
     logging.info("p_opt ({0:.0f}s): {1}".format(t_complete, p_opt))
     logging.info("Writing to {}".format(output_path))
@@ -94,16 +104,19 @@ def optimize_npm_at_point(indices_path, output_path, data, config):
         pickle.dump(result, fp, -1)
 
     # Assign this result to nearby stars?
-    S = config.get("share_optimised_result_with_nearest", 0)
     logging.info("Sharing result with nearest {} neighbours".format(S))
-    for nearby_index in indices[1:1 + S]:
 
+
+    for nearby_index in indices[:1 + S]:
+        if nearby_index == data_index: continue
+        
         nearby_output_path = npm.get_output_path(data["source_id"][nearby_index],
                                                  config)
         if os.path.lexists(nearby_output_path):
             os.unlink(nearby_output_path)
 
-        os.symlink(output_path, nearby_output_path)
+        os.symlink(os.path.abspath(output_path), 
+                   os.path.abspath(nearby_output_path))
 
         logging.info("Shared result with neighbour {} -> {}".format(
             output_path, nearby_output_path))
@@ -171,7 +184,8 @@ if __name__ == "__main__":
                         index, output_path))
                 break
 
-            indices, S, p_opt = optimize_npm_at_point(indices_path,
+            indices, S, p_opt = optimize_npm_at_point(index,
+                                                      indices_path,
                                                       output_path,
[…]

### 2018-06-28 npm_mr_setup.py

@@ -5,9 +5,11 @@ Setup and run the non-parametric model as a map-reduce job.
 import numpy as np
 import os
 import logging
+import multiprocessing as mp
 import pickle
 import yaml
 from astropy.io import fits
+from time import time
 
 import npm_utils as npm
 
@@ -20,9 +22,12 @@ if not os.path.exists(config["results_path"]):
 # Load in the data.
 data = fits.open(config["data_path"])[1].data
 
+all_label_names = list(config["kdtree_label_names"]) \
+                + list(config["predictor_label_names"])
+
 # Set up a KD-tree.
 X = np.vstack([data[ln] for ln in config["kdtree_label_names"]]).T
-finite = np.where(np.all(np.isfinite(X), axis=1))[0]
+finite = np.where(np.all([np.isfinite(data[ln]) for ln in all_label_names], axis=0))[0]
 
 X = X[finite]
 N, D = X.shape
@@ -30,29 +35,43 @@ N, D = X.shape
 kdt, scale, offset = npm.build_kdtree(
     X, relative_scales=config.get("kdtree_relative_scales", None))
 
-query_kwds = dict(offset=offset, scale=scale, full_output=False)
-query_kwds.update(
-    minimum_points=config.get("kdtree_minimum_points", 1024), # DEFAULT
+kwds = dict(offset=offset, scale=scale, full_output=False)
+kwds.update(
     minimum_radius=config.get("kdtree_minimum_radius", None), # DEFAULT
+    minimum_points=config.get("kdtree_minimum_points", 1024), # DEFAULT
+    maximum_points=config.get("kdtree_maximum_points", 8192), # DEFAULT
 )
+overwrite = config.get("overwrite_indices", False)
 
-# For each point, get the indices needed for that star, ensuring that the KD-tree
-# ball follows certain properties (e.g., has enough stars).
-for i in range(N):
+def setup_point(index):
+
+    path = npm.get_indices_path(data["source_id"][index], config)
+    
+    if os.path.exists(path) and not overwrite: 
+        logging.info("Skipping {}/{} because {} exists".format(index, N, path))
+        return None
 
     # Get indices relative to the finite X
-    finite_indices = npm.query_around_point(kdt, X[i], **query_kwds)
+    finite_indices = npm.query_around_point(kdt, X[index], **kwds)
 
     # Translate those finite indices to the original data file, since that is
     # what will be used by the map-jobs to run the optimisation.
[…]

### 2018-06-28 npm_utils.py

@@ -4,29 +4,32 @@ import os
 from scipy import (optimize as op, stats)
 from sklearn import neighbors as neighbours
 from scipy.special import logsumexp
+from time import time
 
 import stan_utils as stan
 
 CONFIG_PATH = "npm-config.yaml"
 
+def split_source_id(source_id, chunk=4):
+    return "/".join(map("".join, zip(*[iter(str(source_id))] * chunk)))
+
+
 
 def get_indices_path(source_id, config):
-    path = os.path.join(config["results_path"], "{0}/indices".format(source_id))
-    dirname = os.path.dirname(path)
-    if not os.path.exists(dirname):
-        os.mkdir(dirname)
-
+    path = os.path.join(config["results_path"], 
+        "{0}/indices".format(split_source_id(source_id)))
+    os.makedirs(os.path.dirname(path), exist_ok=True)
     return path
 
-def get_output_path(source_id, config):
+
+def get_output_path(source_id, config, check_path_exists=True):
     results_suffix = config.get("results_suffix", None)
     path = os.path.join(config["results_path"], "{0}/{1}{2}".format(
-        source_id, 
+        split_source_id(source_id), 
         "+".join(config["predictor_label_names"]),
         ".{}".format(results_suffix) if results_suffix is not None else ""))
-    dirname = os.path.dirname(path)
-    if not os.path.exists(dirname):
-        os.mkdir(dirname)
+    if check_path_exists:
+        os.makedirs(os.path.dirname(path), exist_ok=True)
 
     return path
 
@@ -53,8 +56,8 @@ def build_kdtree(X, relative_scales=None,**kwargs):
     return (kdt, relative_scales, offset)
 
 
-def query_around_point(kdtree, point, offset=0, scale=1, minimum_points=1,
-    minimum_radius=None, dualtree=False, full_output=False):
+def query_around_point(kdtree, point, offset=0, scale=1, minimum_radius=None, 
+    minimum_points=1, maximum_points=None, dualtree=False, full_output=False):
     """
     Query around a point in the KD-Tree until certain conditions are met (e.g.,
     the number of points in the ball, and the minimum radius that the ball
@@ -131,9 +134,22 @@ def query_around_point(kdtree, point, offset=0, scale=1, minimum_points=1,
                 # appropriate radius.
                 left, right = (min_radius, max_radius)
 
-                ri = np.linspace(left, right, N - K)
-                counts = kdtree.two_point_correlation(point, ri)
[…]

### 2018-06-29 article/ms.tex
minor

@@
 
-\newcommand{\NumberOfHigherOrderSystems}{XXX}
-\newcommand{\NumberOfSingleLinedSystems}{XXX}
+\newcommand{\NumberOfStellarMultiples}{XXX}
+\newcommand{\NumberOfStellarSingles}{XXX}
 
@@
 
-\title{Detection and characterisation of stellar multiplicity with \Gaia}
+\title{Detection and characterisation of stellar multiplicity with Gaia}
 
@@
 \email{andrew.casey@monash.edu}
+
+% Some order of the following authors (order is TBD)
 
@@
 
+\author[0000-0003-3494-343X]{Carles Badenes}
+\affiliation{Department of Physics and Astronomy, 
+			 University of Pittsburgh, 
+			 3941 O'Hara Street, Pittsburgh, PA 15260, USA}
+
+\author[0000-0003-0872-7098]{Adrian M. Price-Whelan}
+\affiliation{Department of Astrophysical Sciences, 
+			 Princeton University, 
+			 Princeton, NJ 08544, USA}
+
 \begin{abstract}
-Most stars have stellar companions. The occurrence rate and properties of stellar 
-multiples (e.g., binaries, triples) is relevant to many fields of astrophysics, 
-but the effects of stellar multiplicity are commonly ignored in the analysis of 
-most stellar and galactic surveys. Here we use \Gaia\ data to identify 
-\todo{\NumberOfHigherOrderSystems} stellar multiple systems from radial velocity
-and astrometric jitter, photometry, and intrinsic luminosity. Our sample includes
-\todo{\NumberOfSingleLinedSystems} unresolved binary systems detected solely
-from radial velocity or astrometric jitter, which have orbital periods up to 
-$\approx$3.5\,yrs. We show that the ratio of jitter in radial velocity and
-astrometry can be calibrated to estimate inclination angles (and companion masses)
-for thousands of systems. \todo{We comment on stellar multiplicity with stellar
-metallicity}. \todo{We comment on systems with stellar remnants (e.g. NS/BHs).}
+The properties and occurrence rate of stellar multiples (e.g., binaries, triples) 
+underpins much of astrophysics. Precisely measuring these quantities is challenging
+and observationally demanding. Here we use \Gaia\ data to detect and partially 
+characterise \NumberOfStellarMultiples\ stellar multiple systems based on excess 
+astrometric jitter, excess radial velocity jitter, the absence of reported radial
+velocities, photometry, and intrinsic luminosity.
+\todo{We show that astrometric and radial velocity jitter can be calibrated to
+estimate inclination angles (and directly constrain companion masses) for thousands 
+of systems.}
+Comparisons with literature surveys on stellar multiplicity demonstrate that we 
+reliably detect binary systems with orbital periods up to $\approx$3.5\,yr. 
+\todo{A comment on stellar multiplicity with stellar metallicity.}
+\todo{A comment on stellar multiplicity in the field relative to clusters.}
+\todo{A comment on systems with stellar remnants (e.g., NS/BHs).}
 \end{abstract}
 
-\keywords{stars: dynamics}
+\keywords{(stars:) binaries: general, (stars:) binaries: spectroscopic, (stars:) binaries (including multiple): close, astrometry, techniques: radial velocities, methods: statistical}
 
@@
 these measured properties relative to a single star of a similar type. This is
-nothing new: astronomers have used radial velocity \citep{firstplanet} and 
-astrometric variations \citep{someone} and photometry \citep{someone} to infer the 
+nothing new: astronomers have used radial velocity \citep{Butler:1996} and 
+astrometric variations \citep{Muterspaugh:2010} and photometry \citep{someone} to infer the 
 presence of stellar and exoplanet companions forever. However, the opportunity
@@
 
-\section{Method}
+\section{Method} \label{sec:method}
 
@@
 Radial velocity measurements are not available for individual epochs in the 
-second \Gaia\ data release \citep{Lindegren:2018,Gaia_RVS}, but the a radial velocity point estimate and error
-is available for 7,224,631 sources \cite{gaia_rvs}. 
+second \Gaia\ data release \citep{Lindegren:2018,Cropper:2018}, but the a radial velocity point estimate and error
+is available for 7,224,631 sources \cite{Cropper:2018}. 
 The radial velocity error provided by \Gaia\ ($\GaiaRVE$; column name \texttt{radial\_velocity\_error}) 
@@
 use of astrometric excess noise from the \todo{Tycho-Gaia Astrometric Solution}
-\citep{Michalik, TGAS, etc}.
+\citep{Michalik:2015, Lindegren:2016}.
 
@@
 \begin{figure*}
-	\includegraphics[width=1.0\textwidth]{../figures/sb9-comparison.pdf}
+	\includegraphics[width=1.0\textwidth]{../figures/todo.png}
     \caption{Source radial velocity variance shown with respect to flux in
@@
 \begin{figure}
-	\includegraphics[width=0.5\textwidth]{../figures/sb9-comparison.pdf}
+	\includegraphics[width=0.5\textwidth]{../figures/todo.png}
     \caption{Source mean \Gaia\ RP flux and radial velocity variance.
@@
 \begin{figure}
-	\includegraphics[width=0.5\textwidth]{../figures/sb9-comparison.pdf}
+	\includegraphics[width=0.5\textwidth]{../figures/todo.png}
     \caption{Mixture model fitted.}
@@
 \begin{figure}
-	\includegraphics[width=0.5\textwidth]{../figures/sb9-comparison.pdf}
+	\includegraphics[width=0.5\textwidth]{../figures/todo.png}
     \caption{Detected binary fraction as a function of the mean RP flux
@@
 
-\subsection{Comparison with \citet{El_Badry:2017,El_Badry:2018}}
+\subsection{Comparison with \citet{El-Badry:2018}}
 
-\subsection{Comparison with \citet{Badenes:2017}}
+\subsection{Comparison with \citet{Badenes:2018}}
 
@@
 
-\subsection{Comparison with \citet{Troup:2010}}
+\subsection{Comparison with \citet{Troup:2016}}
 
-\subsection{Comparison with \citet{Price-Whelan:2010}}
+\subsection{Comparison with \citet{Price-Whelan:2018}}
 
@@
 
+% At least some of these people will be promoted to the author list.
 It is a pleasure to thank
-    Adrian Price-Whelan (Princeton University),
-	Lorenzo Spina (Monash University), 
+	Berry Holl (Geneva),
+	Jose Hernandez (ESAC),
+	Daniel Michalik (ESA/ESTEC),
+	Kevin C. Schlaufman (Johns Hopkins University),
 		and
-	Kevin C. Schlaufman (Johns Hopkins University).
-This work was partly supported through the Australian Research Council 
+	Lorenzo Spina (Monash University).
+This work was supported in part by the Australian Research Council 
 through Discovery Grant DP160100637.
@@
 participating in the {\it Gaia} Multilateral Agreement.  This research was
-started at the NYC Gaia DR2 Workshop at the Center for Computational
+developed in part at the NYC Gaia DR2 Workshop at the Center for Computational
 Astrophysics of the Flatiron Institute in 2018 April.
@@
     \package{Stan} \citep{stan},
-    \package{CosmoHub} \citep{cosmohub}
-    %tensorflow
+    \package{CosmoHub} \citep{cosmohub},
+    \package{TensorFlow} \citep{tensorflow}
+    \package{Jupyter Notebooks} \citep{jupyter-notebooks}
 }

### 2018-06-29 article/ms.tex
words

@@
 \usepackage{bm}
+\usepackage{amsmath}
 \usepackage{color}
@@
 These detection methods are sensitive to binary systems with different 
-properties. Here we describe the radial velocity information that forms 
-part of our model before describing how photometry and astrometry are 
-incorporated into the detection and characterisation of stellar multiples.
+properties. Here we describe the absence of reported radial velocity in
+\Gaia\ data can be used to deduce stellar multiples, before we describe
+our model that incorporates measurements of radial velocity, astrometry, 
+and photometry, to detect and characterise stellar multiples.
+
+\subsection{SB2 Systems: Double-lined spectroscopic binaries}
+\label{sec:sb2_methods}
+
+Many sources in the second \Gaia\ data release do not have a reported radial 
+velocity, despite being bright enough ($G \lesssim 13$) and in a suitable 
+temperature range (e.g., between $\approx4000\,\textrm{K}$ and $\approx6500\,\textrm{K}$) 
+for radial velocities to be measured \citep{Cropper:2018}. The principle reason 
+why radial velocity measurements are \emph{not} reported for these stars is 
+because the \Gaia/DPAC team have identified the source to be a double-lined 
+spectroscopic binary (a so-called SB2-type system), either through a 
+composition of two stellar sources present in the spectra, or from multiple 
+(significant) modes in the cross-correlation function. In these situations it
+is not sensible to report a point estimate of the radial velocity of the point 
+source. For fainter or bluer sources, however, the radial velocity 
+may not be reported simply because it is too faint, blue, or red.
+
+We calculated the completeness of radial velocity measurements (e.g., the
+fraction of sources with reported radial velocities) as a function of all 
+available source properties that might affect whether the radial velocity may
+be reported or not. This included position ($\alpha$, $\delta$, $l$, $b$),
+parallax, proper motions, apparent magnitudes, \bp\ - \rp\ colour, 
+properties of the radial velocity templates, stellar parameters ($\teff$, $\radius$, $\luminosity$),
+and other properties. The full list of properties investigated is described in
+Appendix \ref{appendix:sb2}, with corresponding figures.
+
+In Figure \ref{fig:radial_velocity_completeness} we show the completeness as 
+a function of some pertinent properties, which demonstrate that the radial 
+velocity completeness is relatively flat until a source becomes too faint 
+(low \rp\ flux), or is either too blue or too red. We adopt conservative 
+limits for when the radial velocity completeness starts to drop with these 
+properties, and we assume that any source within the following range of 
+source parameters
+\begin{eqnarray}
+    \mathrm{phot\,\,rp\,\,mean\,\,flux} & \in & [10^{4}, 10^8] \nonumber \\
+    \mathrm{bp-rp} & \in & [X, Y] 
+    \label{eq:sb2_criteria}
+\end{eqnarray}
+\noindent{}is likely to be a double-lined spectroscopic binary (SB2) if no 
+radial velocity is reported. That is to say that we are explicitly assuming
+that if a point source meets the criteria in Equation \ref{eq:sb2_criteria} 
+and does not have a reported radial velocity measurement, then the point 
+source is an unresolved double-lined spectroscopic binary. In principle we 
+could make more realistic attempts to model the radial velocity completeness as
+a function of stellar properties rather than simply stating ``sources within
+this parameter range should have radial velocities unless they are SB2 systems'',
+but the radial velocity completeness within our specified range is reasonably
+flat.
+
+\todo{Some confirmation of this? The properties of SB2 candidate systems relative to others. Figure \ref{fig:sb2_histograms}.}
+
+We stress that our sample of SB2 type systems is calculated (or rather, deduced)
+separately to our calculations of multiplicity probability given the 
+\emph{measured} quantities from \Gaia\ like radial velocity, astrometry, and
+photometry. In Section \ref{sec:discussion} we detail some of the immediate
+limitations or interpretations of this assumption, and we emphasise that the 
+main conclusions of this work do not depend on SB2 systems. Nevertheless, we 
+\emph{strongly} 
+caution that our deductive inference of SB2 systems will be far more uncertain 
+than the binary probabilities we derive from other information. Our indicator 
+whether a star is likely a SB2 system likely suffers from more contamination
+and lower completeness compared to the probabilities we derive from other information
+(e.g., the radial velocity jitter, astrometric excess noise, and photometric 
+colours), and the contamination and completeness likely vary over some
+complex (unknown) function of parameter space.
+
+\begin{figure*}
+	\includegraphics[width=1.0\textwidth]{../figures/todo.png}
+    \caption{Fraction of \Gaia\ sources with reported radial velocities
+		     as a function of source properties. Within the adopted source
+		     parameter ranges (gray; see Section \ref{sec:sb2_methods})
+		     the radial velocity completeness is approximately flat.
+		     We flag sources within this range that do not have reported 
+		     radial velocities to be candidate SB2 systems.}
+    \label{fig:radial_velocity_completeness}
+\end{figure*}
+
+\begin{figure*}
+	\includegraphics[width=1.0\textwidth,height=3.0in]{../figures/todo.png}
+    \caption{Distribution of 
+    			(a) photometric \bp\ - \rp\ excess factor, 
+				(b) astrometric unit weight error, 
+					and 
+				(c) photometric variability
+			of candidate SB2 type systems relative to a control sample.}
+    \label{fig:sb2_histograms}
+\end{figure*}
+
+% Figure: radial velocity completeness as a function of pertinent properties
+% 		  with the range that we have said is "fully complete"
+
+% Figure: h-r diagram of SB2 binary fraction
+
+% Figure: cross-match against a catalog of SB2 systems,... do we find all of
+% 		  what they find from SB2 deductive inference alone? or do we miss
+% 		  some SB2s in some parameter range?
+
+% Figure: radial veloci
 
@@
 Radial velocity measurements are not available for individual epochs in the 
-second \Gaia\ data release \citep{Lindegren:2018,Cropper:2018}, but the a radial velocity point estimate and error
-is available for 7,224,631 sources \cite{Cropper:2018}. 
-The radial velocity error provided by \Gaia\ ($\GaiaRVE$; column name \texttt{radial\_velocity\_error}) 
+second \Gaia\ data release \citep{Lindegren:2018,Cropper:2018}, but a point estimate
+and error in radial velocity is available for 7,224,631 sources \cite{Cropper:2018}.
+% TODO: That's the value from the paper, but I think the true value is lower. Check.
+The reported radial velocity error ($\GaiaRVE$; column name \texttt{radial\_velocity\_error}) 
 is a function of the number of transits $N$ (i.e., the number of observations; 
-represented by column \texttt{rv\_nb\_transits}) and the scatter among those $N$ measurements. 
-This allows us to calculate the scatter among those measurements, which we will
-refer to as the \emph{radial velocity jitter}, for each source
+represented by column \texttt{rv\_nb\_transits}) and the standard deviation among 
+those $N$ measurements of radial velocity. This allows us to calculate the standard
+deviation in radial velocity for each source, independent of the number of measurements,
+which we will refer to as the \emph{radial velocity jitter} $\RVJitter$,
 \begin{eqnarray}
@@
 jitter represents the minimum uncertainty with which \Gaia\ can measure 
-radial velocity for a star of that apparent brightness and colour. However for 
-stars with stellar companions and no evidence of double lines or binarity in the
+radial velocity for a star of that colour, apparent brightness, and absolute brightness.
+However for stars with stellar companions and no evidence of double lines or binarity in the
 cross-correlation-function, so-called SB1 type systems \citep{someone}, the
@@
       minimum radial velocity noise that \Gaia\ can measure for a star of that
-      apparent magnitude and colour; and
+      \bp\ - \rp\ colour, apparent \rp\ magnitude, and absolute \rp\ magnitude;
 
 \item \emph{higher-order star systems}, where the radial velocity jitter is
-      significantly above that for single-star systems of a similar apparent
-      magnitude and colour.
+      significantly above that for single-star systems of a similar colour, 
+      apparent magnitude, and absolute magnitude.
 \end{enumerate}
 
-We make a number of other explicit assumptions about the radial velocity
-measurements reported by \Gaia:
+The mean radial velocity jitter of a single star will vary depending on the 
+\bp\ - \rp\ colour of the source and the apparent magnitude. Similarly, the mean
+radial velocity jitter of stellar \emph{multiples} will depend on the source 
+colour and its position in the Hertzsprung-Russell diagram. For these reasons,
+we must consider the radial velocity jitter of a source in context with other
+sources that have a similar \bp\ - \rp\ colour, a similar apparent \rp\ magnitude,
+and a similar absolute \rp\ magnitude. In principle if we knew the underlying
+distribution and properties of stellar multiples then we could model how different
+binary (and higher order) systems would be observable across the Hertszprung-Russell
+diagram.
 
-\begin{itemize}
-	\item the radial velocity jitter is a function of the apparent brightness 
-		  of the point source
-	\item the radial velocity jitter is a function of the apparent $bp-rp$
-		  colour of the point source
-	\item there is some intrinsic spread to the radial velocity jitter of
-		  single stars, and that intrinsic spread also increases with the 
-		  apparent brightness and colour of the source
-	\item the difference between the true surface gravity $\logg$ and the
-		  surface gravity of the radial velocity template used
-		  (column \texttt{rv\_template\_logg}) has no impact on the radial
-		  velocity jitter
-	\item the difference between the true stellar metallicity [Fe/H] and
-	 	  the metallicity of the radial velocity template used
-		  (column \texttt{rv\_template\_feh}) has no impact on the radial
-		  velocity jitter
-	\item \todo{state something about teff template}		  
-\end{itemize}
+We chose to adopt a non-parametric model for stellar multiplicity across the
+Hertsprung-Russell diagram because we do not know the distribution of properties
+of stellar multiples. Specifically, for each source we take the following steps:
 
-Conditioned on these assumptions, we adopt a two-component mixture model for 
-the radial velocity jitter: a normal distribution where the mean and variance
-are a function of stellar properties, and a uniform background component to
-account for stellar multiples. Specifically, the likelihood is
+\begin{enumerate}
 
+	\item Select sources with similar \bp\ - \rp\ colours, a similar apparent \rp\
+		  magnitude, and a similar absolute \rp\ magnitude.
+
+	\item We fit a two-component mixture model to the distribution of radial
+		  velocity jitter for the subset of similar sources selected in Step 1.
+
+	\item Evaluate whether the source of interest is more consistent with being
+		  drawn from the single star component (the normal distribution) or the
+		  stellar multiple component (the log-normal distribution).
+
+	\item For sources that are more consistent with being drawn from the stellar
+		  multiple component, provide a partial characterisation of the system by
+		  calculating the excess radial velocity jitter (above what is expected 
+		  for a single star with similar observed properties), among other 
+		  quantities.
+
+\end{enumerate}
+
+We illustrate this procedure in Figure \ref{fig:npm_schematic}, where we demonstrate
+how the distribution of source properties (e.g., radial velocity jitter) varies across
+the Hertsprung-Russell diagram. While this model is flexible enough to describe varying
+properties of single stars and stellar multiples, there are a number of explicit 
+assumptions that go into this model:
+
+\todo{assumptions}
+
+\begin{figure*}
+	\includegraphics[width=1.0\textwidth]{../figures/todo.png}
+    \caption{Illustration of the method we adopt for identifying and characterising
+    		 stellar multiplicity across the Hertsprung-Russell diagram. For each
+		 	 \Gaia\ source we select between 128 and 1024 sources that have
+			 similar \bp\ - \rp\ colours, apparent \rp\ magnitudes, and absolute \rp\
+			 magnitudes (Step 1). With these similar sources we fit a two-component
+			 model to distinguish single star sources from stellar multiples (Step 2).
+			 With the model parameters optimised from Step 2, we evaluate probability
+			 of stellar multiplicity and partially characterise the higher order
+			 systems by assuming binarity and estimating the system semi-amplitude (Step 3).}
+    \label{fig:npm_schematic}
+\end{figure*}
+
+\subsubsection*{Step 1: Selecting similar sources}
+
+We construct a $k$-dimensional tree \citep[k-d tree;][]{} for all sources in the
+dimensions of \bp\ - \rp\ colour, apparent \rp\ magnitude, and absolute \rp\ magnitude.
+When doing so we scale the dimensions such that 0.1\,mags in \bp\ - \rp\ colour is
+approximately equal to 1 magnitude in apparent \rp\ magnitude and 1 magnitude in 
+absolute \rp\ magnitude. We use this k-d tree for constructing the `ball' of similar
+stars for every source.
+
+We enforce a number of constraints when querying for similar stars. For each source
+we require the `ball' in multi-dimensional space to include at least 128 sources.
+We further require that the radius of the ball extends to at least 0.05\,mags in 
+\bp\ - \rp\ colour, 0.5\,mags in apparent \rp\ magnitude, and 0.5\,mags in absolute
+\rp\ magnitude. If the initial query of 128 points does not extend wide enough to
+meet our constraints on minimum radius, we extend the radius (linearly in each scaled)
+dimension until the minimum constraints on radii are met. Our minimum radii constraints
+imply that there will be many (e.g., $>>128$) points returned in dense regions of the
+Hertsprung-Russell diagram. In these situations we then subsample the selected points
+to return a random 1024 sources in any ball, while enforcing our constraints on
+minimum radii. As a result, the number of `similar' sources selected will range from
+a minimum of 128 to a maximum of 1024, and the minimum radii constraints on various
+dimensions are always maintained.
+
+\subsubsection*{Step 2: Optimizing the two-component mixture}
+
+We fit a two-component mixture model to the distribution of radial velocity jitter for
+all similar sources selected through Step 1. This mixture model includes a normal
+distribution to represent the intrinsic jitter for single stars, and a log-normal
+distribution to represent stellar multiples. Specifically the model parameters include
+a mixing weight parameter $\theta$ that is constrained between $(0, 1)$, the mean 
+$\mu_{s}$ and standard deviation $\sigma_{s}$ of the normal distribution,
+and the mean $\mu_{m}$ and standard deviation $\sigma_{m}$ of the 
+log-normal distribution. For brevity we
+
+We assume a beta prior on $\theta$  such that $\theta \sim \mathcal{B}\left(5, 5\right)$
+and we require $\sigma_{s}$ and $\sigma_{m}$ to be larger than zero.
+Depending on the distribution of the radial velocity jitter, we often found this mixture
+challenging to optimize because the log-normal distribution could easily account for 
+most of the observed distribution, which tended the mixing weight to zero and implied
+that all selected sources were stellar multiples. To avoid this situation we placed a
+constraint on the mean of the log-normal distribution to ensure that the \emph{mode}
+of the log-normal distribution is between $\mu_{s} + 1\sigma_{s}$ and
+$\mu_{s} + 5\sigma_{s}$. The mode of the log-normal distribution is
+$\exp\left(\mu_{m} - \sigma_{m}^2\right)$ such that
 \begin{eqnarray}
-%	\mathcal{L} & \equiv & p\left(\{y
-    MODEL
+	\mu_{s} + 5\sigma_{s} & > & \exp\left(\mu_{m} - \sigma_{m}^2\right) \label{eq:rvm_lower_bound} \\
+	\mu_{s} + 1\sigma_{s} & < & \exp\left(\mu_{m} - \sigma_{m}^2\right) \label{eq:rvm_upper_bound} \quad .
 \end{eqnarray}
 
-\noindent{}where \todo{X is Y} abd .
+However, we found that this constraint alone was insufficient. Often the mode would
+be very close to the $\mu_{s} + 1\sigma_{s}$ lower bound, which implies that
+the log-normal distribution is providing a considerable fraction of probability support
+to sources with radial velocity jitter near the mean of the \emph{normal} distribution.
+For this reason we chose to alter our lower bound constraint in that we also required
+that at the mean of the normal distribution, the cumulative probability distribution
+of the log-normal must be less than 10\% (let $q = 0.10$). The cumulative distribution 
+function for the log-normal distribution is
+\begin{eqnarray}
+\textrm{CDF}(x|\mu_{m},\sigma_{m}) = \frac{1}{2} + \frac{1}{2}\textrm{erf}\left(\frac{\log{x} - \mu_{m}}{\sqrt{2}\sigma_{m}}\right) 
+\end{eqnarray}
+
+\noindent{}where \textrm{erf} is the standard error function. We use the approximation 
+to the error function 
+\begin{eqnarray}
+	\textrm{erf}\left(x\right) \approx \tanh\left(\sqrt{\pi}\log{2}x\right)
+\end{eqnarray}
+
+\noindent{}and use the identity
+\begin{eqnarray}
+	\tanh^{-1}\left(x\right) = \frac{1}{2}\left[\log{\left(1 + x\right)} - \log{\left(1 - x\right)}\right]
+\end{eqnarray}
+
+\noindent{}to write our constraint on $\mu_{m}$ as
+\begin{eqnarray}
+	\mu_{m} & > & \log\mu_{s} + \frac{\sigma_{m}}{2\log{2}}\sqrt{\frac{2}{\pi}}\log{\left(\frac{1}{q} - 1\right)} \nonumber \\
+	\mu_{m} & > & \log\mu_{s} + \frac{\sigma_{m}}{2\log{2}}\sqrt{\frac{2}{\pi}}\log{9} \label{eq:rvm_lower_bound2} \quad .
+\end{eqnarray}
+
+Incorporating the bounds from Equations \ref{eq:rvm_lower_bound}, \ref{eq:rvm_upper_bound},
+and \ref{eq:rvm_lower_bound2} leads to the prior
+\begin{eqnarray}
+	\mu_{m} \sim \mathcal{U}\left(\alpha, \beta\right)
+\end{eqnarray}
+
+\noindent{}where:
+\begin{eqnarray}
+  \alpha & = & \max\left(\!
+	\begin{aligned}
+		\log{\left(\mu_{s} + 1\sigma_{s}\right)} + \sigma_{m}^2\\
+		\log\mu_{s} + \frac{\sigma_{m}}{2\log{2}}\sqrt{\frac{2}{\pi}}\log{9}
+	\end{aligned}\right) \\
+  \beta & = & \log{\left(\mu_{s} + 5\sigma_{s}\right)} + \sigma_{m}^2 \quad .
+\end{eqnarray}
+
+We found these priors were sufficient to ensure numerical stability and efficiency.
+
+Wherever possible we initialised the model parameters from nearby points (within the
+ball selected by the k-d tree) where an optimised result already exists. This was primarily to minimise
+computational cost; the initialisation did not have a noticeable impact on the
+optimised result. We required the log probability and the gradient of the log
+probability to reach machine precision before considering the optimisation
+converged. After convergence, we queried to see if any points within the ball of
+similar stars had not been optimized. If no result existed, we would next move to that
+point. We refer to this process as \emph{swarm optimisation}, which we conducted in 
+parallel, and illustrate in Figure \ref{fig:swarm_optimisation}.
+
+\todo{To minimise computational cost we assigned the closest 8 stars with the result
+from others}
+
+\subsubsection*{Step 3: Evaluating probability and system properties}
+
+\todo{selecting similar sources}
+
+\todo{fitting the two-component model}
+
+\todo{evaluating probability}
+
+\todo{calculating source properties}
 
@@
 
-\subsection{SB2 Systems: Double-lined spectroscopic binaries}
-\label{sec:sb2_methods}
-
-Many sources in the second \Gaia\ data release do not have any reported radial 
-velocity measurement, despite being bright enough ($G \lesssim 13$) and in a
-suitable temperature range (e.g., between $X$ and $Y$) for radial velocities to 
-be measured \citep{someone}. The principle reason why radial velocity measurements 
-are \emph{not} reported for these stars is because the \Gaia/DPAC team have 
-identified the source to be a double-lined spectroscopic binary (a SB2-type system), 
-either through a composition of two stellar sources present in the spectra, or
-from multiple (significant) modes in the cross-correlation function. In these
-situations it is not sensible to report a point estimate of the radial velocity
-of the point source. For fainter or bluer sources, however, the radial velocity 
-may not be reported simply because it is too faint, blue, or red.
-
-We calculated the completeness of radial velocity measurements (e.g., the
-fraction of sources with reported radial velocities) as a function of all 
-available source properties that might affect whether the radial velocity may
-be reported or not. This included position ($\alpha$, $\delta$, $l$, $b$),
-parallax, proper motions, apparent magnitudes, colours, properties of the radial
-velocity templates, stellar parameters ($\teff$, $\radius$, $\luminosity$),
-and other properties. The full list of properties investigated is described in
-Appendix \ref{appendix:sb2}. In Figure \ref{fig:radial_velocity_completeness} 
-we show the completeness as a function of some pertinent properties, which demonstrate 
-that the radial velocity completeness is relatively flat until a star becomes 
-too faint (low \rp\ flux), or is either too blue or too red. We adopt 
-conservative limits for when the radial velocity completeness starts to drop 
-with these properties, and we assume that any source within the following
-parameter range
-\begin{eqnarray}
-    \mathrm{phot\,\,rp\,\,mean\,\,flux} & \in & [10^{4}, 10^8] \nonumber \\
-    \mathrm{bp-rp} & \in & [X, Y] 
-    \label{eq:sb2_criteria}
-\end{eqnarray}
-\noindent{}is likely to be a double-lined spectroscopic binary (SB2) if no 
-radial velocity is reported. That is to say that we are explicitly assuming
-that if a point source meets the criteria in Equation \ref{eq:sb2_criteria} 
-and does not have a radial velocity measurement, then the point source is
-an unresolved double-lined spectroscopic binary. In principle we could
-make more realistic attempts to model the radial velocity completeness as
-a function of stellar properties rather than simply stating ``sources within
-this parameter range should have radial velocities unless they are SB2 systems'',
-but the radial velocity completeness within our specified range is reasonably
-flat.
-
-In Section \ref{sec:discussion} we detail some of the immediate limitations or 
-interpretations of this assumption, but here we  note that the probability of a 
-source being a double-lined spectroscopic binary is calculated (or rather, 
-deduced) separately to our calculations of binarity probability given 
-\emph{measured} quantities like radial velocity, astrometry, and photometric 
-colours. And as we discuss in Section \ref{sec:discussion}, the main conclusions
-of this work do not depend on SB2 systems. Nevertheless, we \emph{strongly} 
-caution that our deductive inference of SB2 systems will be far more uncertain 
-than the binary probabilities we derive from other information. Our indicator 
-whether a star is likely a SB2 system likely suffers from more contamination
-and lower completeness compared to the probabilities we derive from other information
-(e.g., the radial velocity jitter, astrometric excess noise, and photometric 
-colours), and the contamination and completeness likely vary over some
-complex (unknown) function of parameter space.
-
-% Figure: radial velocity completeness as a function of pertinent properties
-% 		  with the range that we have said is "fully complete"
-
-% Figure: h-r diagram of SB2 binary fraction
-
-% Figure: cross-match against a catalog of SB2 systems,... do we find all of
-% 		  what they find from SB2 deductive inference alone? or do we miss
-% 		  some SB2s in some parameter range?
-
-% Figure: radial velocity fraction as a function of galactic latitude and
+ty fraction as a function of galactic latitude and
 % 		  longitude
@@
 
-\subsection{Astrometric binaries}
+\subsection{Astrometric binaries} \label{sec:method_astrometric_binaries}
 
 Many higher-order star systems are unresolved point sources in \Gaia, but the
-stellar companion(s) have sufficient mass to measurably affect the tangential 
-velocity of the primary and produce a jitter in astrometric positions. There are
-many other sources of astrometric jitter, but here we show that the principle
-effect of astrometric excess noise is due to stellar companions. Note that due
-to the astrometry `degree of freedom bug' in the processing of the second \Gaia\
-data release \citep[e.g., see Appendix A1 of ][]{Lindegren:2018}, we only make
-use of astrometric excess noise from the \todo{Tycho-Gaia Astrometric Solution}
-\citep{Michalik:2015, Lindegren:2016}.
+stellar companion(s) may have sufficient mass to measurably affect the tangential 
+velocity of the primary and produce a detectable jitter in the astrometric 
+positions. There are many other sources of astrometric jitter, but here we show 
+that the principle effect of astrometric excess noise is due to stellar 
+companions. Note that due to the astrometry `degree of freedom bug' in the 
+processing of the second \Gaia\ data release \citep[e.g., see Appendix A1 of ][]{Lindegren:2018}, 
+that resulted in 80\% of sources having zero astrometric excess noise,
+we instead make use of the astrometric unit weight error $u$
+\begin{eqnarray}
+	u = \sqrt{\frac{\chi^2_{al}}{N_{obs,al} - 5}}
+\end{eqnarray}
+\noindent{}where $\chi^2_{al}$ is the astrometric $\chi^2$ value
+in the along-scan direction (column \texttt{astrometric\_chi2\_al}) and 
+$N_{obs,al}$ is the number of astrometric observations in the along-scan
+direction (column \texttt{astrometric\_n\_obs\_al}). The values of this
+quantity can be interpreted similarly to how the astrometric excess noise
+would be interpreted (e.g., large values are less consistent with a
+single star model), although like the radial velocity and astrometric
+jitter, it must be considered in context with stars of similar colour,
+apparent magnitude (and preferably, absolute magnitude).
 
-\begin{itemize}
-    \item when would we expect AEN? what stellar or orbital properties does it 
-          correlate or grow with?
-    \item calibration model for AEN in DR1
-\end{itemize}
+\subsection{Photometric binaries} \label{sec:method_photometric_binaries}
 
@@
 
-\section{Discussion}
+\section{Discussion} \label{sec:discussion}
 
@@
 It is a pleasure to thank
-	Berry Holl (Geneva),
+	Berry Holl (Observatoire de Gen\'eve),
 	Jose Hernandez (ESAC),
@@
 	Kevin C. Schlaufman (Johns Hopkins University),
+	Lorenzo Spina (Monash University),
 		and
-	Lorenzo Spina (Monash University).
+	Sergey Koposov (Carnegie Mellon University).
 This work was supported in part by the Australian Research Council 

### 2018-06-29 npm_utils.py
minor

@@ -135,7 +135,7 @@ def query_around_point(kdtree, point, offset=0, scale=1, minimum_radius=None,
                 left, right = (min_radius, max_radius)
 
                 max_scale = int(np.log10(N - K))
-                Q = None # 10 * max_scale
+                Q = 10 * max_scale
                 for scale in np.logspace(-max_scale, 1, Q):
     
                     P = int(scale * (N - K))

### 2018-06-29 npm_mr_setup.py
use memory mapped file for indices instead of the other bullshit

@@ -10,6 +10,7 @@ import pickle
 import yaml
 from astropy.io import fits
 from time import time
+import tqdm
 
 import npm_utils as npm
 
@@ -41,15 +42,18 @@ kwds.update(
     minimum_points=config.get("kdtree_minimum_points", 1024), # DEFAULT
     maximum_points=config.get("kdtree_maximum_points", 8192), # DEFAULT
 )
-overwrite = config.get("overwrite_indices", False)
+
+
+# Prepare the memory mapped array.
+memmap_kwds = dict(filename=config["indices_path"], dtype=np.int32)
+fp = np.memmap(mode="w+", shape=(N, kwds["maximum_points"]), **memmap_kwds)
+del fp
+
+# Calculate the number of bytes per row.
+bpr = kwds["maximum_points"] * np.dtype(memmap_kwds["dtype"]).itemsize
+
 
 def setup_point(index):
-
-    path = npm.get_indices_path(data["source_id"][index], config)
-    
-    if os.path.exists(path) and not overwrite: 
-        logging.info("Skipping {}/{} because {} exists".format(index, N, path))
-        return None
 
     # Get indices relative to the finite X
     finite_indices = npm.query_around_point(kdt, X[index], **kwds)
@@ -57,21 +61,23 @@ def setup_point(index):
     # Translate those finite indices to the original data file, since that is
     # what will be used by the map-jobs to run the optimisation.
     indices = finite[finite_indices]
+    
+    # Write the result to the right place in memory.
+    fp = np.memmap(mode="r+", shape=(1, kwds["maximum_points"]),
+                   offset=index * bpr, **memmap_kwds)
 
-    # Save the indices to disk.
-    with open(path, "wb") as fp:
-        pickle.dump(indices, fp, -1)
-
-    logging.info("{}/{}: Saved {} nearby points to {}".format(
-        index, N, indices.size, path))
+    P = indices.size
+    fp[0, :P] = indices
+    fp[0, P:] = -1
+    fp.flush()
+    del fp
 
     return None
 
 
-t_init = time()
+pool = mp.Pool(processes=100) # mp.cou_count()
+for _ in tqdm.tqdm(pool.imap_unordered(setup_point, range(N)), total=N):
+    pass
 
-processes = mp.cpu_count()
-with mp.Pool(processes) as pool:
-    pool.map(setup_point, range(N))
-
-logging.info("Setup completed in {:.0f}s".format(time() - t_init))
+pool.join()
+pool.close()

### 2018-06-29 npm_utils.py

@@ -10,18 +10,6 @@ import stan_utils as stan
 
 CONFIG_PATH = "npm-config.yaml"
 
-def split_source_id(source_id, chunk=4):
-    return "/".join(map("".join, zip(*[iter(str(source_id))] * chunk)))
-
-
-
-def get_indices_path(source_id, config):
-    path = os.path.join(config["results_path"], 
-        "{0}/indices".format(split_source_id(source_id)))
-    os.makedirs(os.path.dirname(path), exist_ok=True)
-    return path
-
-
 def get_output_path(source_id, config, check_path_exists=True):
     results_suffix = config.get("results_suffix", None)
     path = os.path.join(config["results_path"], "{0}/{1}{2}".format(

### 2018-06-29 npm_mr_run.py
wip

@@ -14,12 +14,12 @@ import stan_utils as stan
 
 
 
-
-
 def initialize_from_nearby_point(indices, data, config):
     """
     Look for an initialization point from nearby results.
     """
+    raise NotImplementedError
+    return None
 
     for index in indices:
         source_id = data["source_id"][index]
@@ -43,16 +43,14 @@ def initialize_from_nearby_point(indices, data, config):
 
 
 
-def optimize_npm_at_point(data_index, indices_path, output_path, data, config):
+def optimize_npm_at_point(index, indices, data, config):
 
     # Get indices and load data.
-    with open(indices_path, "rb") as fp:
-        indices = pickle.load(fp)
-
     y = np.array([data[ln][indices] for ln in config["predictor_label_names"]]).T
     N, D = y.shape
 
     init_dict = initialize_from_nearby_point(indices, data, config)
+
     if init_dict is None:
         logging.info("Doing pre-initialization based on the data.")
         init_from_source_id = -1
@@ -145,59 +143,53 @@ if __name__ == "__main__":
     with open(npm.CONFIG_PATH, "r") as fp:
         config = yaml.load(fp)
 
-    # Construct function to get result path.
 
     # Load the data.
     data = fits.open(config["data_path"])[1].data
     N = len(data)
 
+
+    indices_kwds = dict(filename=config["indices_path"], dtype=np.int32)
+    indices_bpr = config["kdtree_maximum_points"] \
+                * np.dtype(indices_kwds["dtype"]).itemsize
+
+
+
     # Generator for a random index to run when the swarm gets bored.
     def indices():
         yield from np.random.choice(N, N, replace=False)
 
-    count = 0
-    for jumps, index in enumerate(indices()):
+    def run_swarm():
 
-        # Swarm strategy.
-        while True:
-            source_id = data["source_id"][index]
+        count = 0
+        for jumps, index in enumerate(indices()):
 
-            logging.info("At source {}/{}: {} (Gaia DR2 {}) {} jumps".format(
-                count, N, index, source_id, jumps))
+            # Swarm strategy.
+            while True:
+                
+                # Get indices for this index.
+                indices = np.memmap(mode="r", 
+                                    shape=(config["kdtree_maximum_points"], ),
+                                    offset=index * indices_bpr,
+                                    **indices_kwds)
 
-            indices_path = npm.get_indices_path(source_id, config)
+                if not any(fp > 0):
+                    # No indices. Skip.
+                    # TODO: update output file with nans at this row?
+                    break
+                
+                
[…]

### 2018-06-29 npm_mr_run.py
make npm run script use memory mapping

@@ -2,45 +2,66 @@
 import pickle
 import logging
 import numpy as np
+import multiprocessing as mp
+import tqdm
 import os
 import yaml
-from glob import glob
 from time import time
 from astropy.io import fits
-from random import shuffle
 
 import npm_utils as npm
 import stan_utils as stan
 
 
+with open(npm.CONFIG_PATH, "r") as fp:
+    config = yaml.load(fp)
 
-def initialize_from_nearby_point(indices, data, config):
+
+# Load the data.
+data = fits.open(config["data_path"])[1].data
+
+
+indices_kwds = dict(filename=config["indices_path"], dtype=np.int32)
+indices_bpr = config["kdtree_maximum_points"] \
+            * np.dtype(indices_kwds["dtype"]).itemsize
+
+
+N, L = (len(data), 4 * len(config["predictor_label_names"]) + 1)        
+results_kwds = dict(filename=config["results_path"], dtype=np.float32)
+results_bpr = N * np.dtype(results_kwds["dtype"]).itemsize
+
+# Create a results file.
+if not os.path.exists(config["results_path"]):
+    logging.info("Creating results file: {}".format(config["results_path"]))
+
+    fp = np.memmap(mode="w+", shape=(N, L), **results_kwds)
+    fp[:] = np.nan
+    fp.flush()
+    del fp
+
+
+def initialize_from_nearby_point(indices):
     """
     Look for an initialization point from nearby results.
     """
-    raise NotImplementedError
-    return None
 
     for index in indices:
-        source_id = data["source_id"][index]
-        output_path = npm.get_output_path(source_id, config)
-        if os.path.exists(output_path):
+        fp = np.memmap(mode="c", shape=(L, ), offset=results_bpr * index, 
+                       **results_kwds)
 
-            with open(output_path, "rb") as fp:
-                result = pickle.load(fp)
+        if np.all(np.isfinite(fp)):
+            values = npm._unpack_params(np.copy(fp))
+            keys = ("theta",
+                    "mu_single", "sigma_single",
+                    "mu_multiple", "sigma_multiple")
+            del fp
+            return (index, dict(zip(keys, values)))
 
-            # Ensure that we have the right dimensions.
-            p_opt = result["p_opt"]
-            for k in ("mu_single", "sigma_single", "mu_multiple", 
-                      "sigma_multiple", "mu_multiple_uv"):
-                p_opt[k] = np.atleast_1d(p_opt[k])
+        else:
[…]

### 2018-06-29 npm_mr_setup.py

@@ -4,12 +4,9 @@ Setup and run the non-parametric model as a map-reduce job.
 
 import numpy as np
 import os
-import logging
 import multiprocessing as mp
-import pickle
 import yaml
 from astropy.io import fits
-from time import time
 import tqdm
 
 import npm_utils as npm

### 2018-06-30 npm_mr_run.py
good stuff

@@ -5,13 +5,20 @@ import numpy as np
 import multiprocessing as mp
 import tqdm
 import os
+import contextlib
+import io
+import sys
 import yaml
+import warnings
 from time import time
 from astropy.io import fits
 
 import npm_utils as npm
 import stan_utils as stan
 
+np.random.seed(42)
+
+warnings.filterwarnings("ignore")
 
 with open(npm.CONFIG_PATH, "r") as fp:
     config = yaml.load(fp)
@@ -20,24 +27,58 @@ with open(npm.CONFIG_PATH, "r") as fp:
 # Load the data.
 data = fits.open(config["data_path"])[1].data
 
+all_label_names = list(config["kdtree_label_names"]) \
+                + list(config["predictor_label_names"])
+
+# Set up a KD-tree.
+X = np.vstack([data[ln] for ln in config["kdtree_label_names"]]).T
+finite = np.where(np.all([np.isfinite(data[ln]) for ln in all_label_names], axis=0))[0]
+
+N, _ = X.shape
+F = finite.size
+L = 4 * len(config["predictor_label_names"]) + 1
 
 indices_kwds = dict(filename=config["indices_path"], dtype=np.int32)
 indices_bpr = config["kdtree_maximum_points"] \
             * np.dtype(indices_kwds["dtype"]).itemsize
 
-
-N, L = (len(data), 4 * len(config["predictor_label_names"]) + 1)        
 results_kwds = dict(filename=config["results_path"], dtype=np.float32)
-results_bpr = N * np.dtype(results_kwds["dtype"]).itemsize
+results_bpr = L * np.dtype(results_kwds["dtype"]).itemsize
 
 # Create a results file.
-if not os.path.exists(config["results_path"]):
-    logging.info("Creating results file: {}".format(config["results_path"]))
-
+if not os.path.exists(results_kwds["filename"]):
+    logging.info("Creating results file: {}".format(results_kwds["filename"]))
     fp = np.memmap(mode="w+", shape=(N, L), **results_kwds)
-    fp[:] = np.nan
-    fp.flush()
     del fp
+
+class suppress_stdout_stderr(object):
+    '''
+    A context manager for doing a "deep suppression" of stdout and stderr in
+    Python, i.e. will suppress all print, even if the print originates in a
+    compiled C/Fortran sub-function.
+       This will not suppress raised exceptions, since exceptions are printed
+    to stderr just before a script exits, and after the context manager has
+    exited (at least, I think that is why it lets exceptions through).
+
+    '''
+    def __init__(self):
[…]

### 2018-06-30 npm_mr_setup.py

@@ -14,9 +14,6 @@ import npm_utils as npm
 with open(npm.CONFIG_PATH, "r") as fp:
     config = yaml.load(fp)
 
-if not os.path.exists(config["results_path"]):
-    os.mkdir(config["results_path"])
-
 # Load in the data.
 data = fits.open(config["data_path"])[1].data
 
@@ -26,12 +23,11 @@ all_label_names = list(config["kdtree_label_names"]) \
 # Set up a KD-tree.
 X = np.vstack([data[ln] for ln in config["kdtree_label_names"]]).T
 finite = np.where(np.all([np.isfinite(data[ln]) for ln in all_label_names], axis=0))[0]
-
-X = X[finite]
 N, D = X.shape
+F = finite.size
 
 kdt, scale, offset = npm.build_kdtree(
-    X, relative_scales=config.get("kdtree_relative_scales", None))
+    X[finite], relative_scales=config.get("kdtree_relative_scales", None))
 
 kwds = dict(offset=offset, scale=scale, full_output=False)
 kwds.update(
@@ -40,41 +36,37 @@ kwds.update(
     maximum_points=config.get("kdtree_maximum_points", 8192), # DEFAULT
 )
 
-
 # Prepare the memory mapped array.
 memmap_kwds = dict(filename=config["indices_path"], dtype=np.int32)
-fp = np.memmap(mode="w+", shape=(N, kwds["maximum_points"]), **memmap_kwds)
-del fp
+if not os.path.exists(memmap_kwds["filename"]):
+    fp = np.memmap(mode="w+", shape=(N, kwds["maximum_points"]), **memmap_kwds)
+    del fp
 
 # Calculate the number of bytes per row.
 bpr = kwds["maximum_points"] * np.dtype(memmap_kwds["dtype"]).itemsize
 
 
 def setup_point(index):
-
-    # Get indices relative to the finite X
-    finite_indices = npm.query_around_point(kdt, X[index], **kwds)
-
-    # Translate those finite indices to the original data file, since that is
-    # what will be used by the map-jobs to run the optimisation.
-    indices = finite[finite_indices]
-    
-    # Write the result to the right place in memory.
-    fp = np.memmap(mode="r+", shape=(1, kwds["maximum_points"]),
+    fp = np.memmap(mode="r+", shape=(kwds["maximum_points"], ),
                    offset=index * bpr, **memmap_kwds)
 
+    if any(fp > 0):
+        del fp
+        return None
+
[…]

### 2018-06-30 stan_utils.py

@@ -9,7 +9,8 @@ import pystan.plots as plots
 __all__ = ["load_stan_model", "sampling_kwds", "plots"]
 
 
-def load_stan_model(path, cached_path=None, recompile=False, overwrite=True):
+def load_stan_model(path, cached_path=None, recompile=False, overwrite=True,
+    verbose=True):
     r"""
     Load a Stan model from a file. If a cached file exists, use it by default.
 
@@ -36,13 +37,15 @@ def load_stan_model(path, cached_path=None, recompile=False, overwrite=True):
             model = pickle.load(fp)
 
         if model.model_code != model_code:
-            logging.warn("Cached model at {} differs from the code in {}; "\
-                         "recompiling model".format(cached_path, path))
+            if verbose:
+                logging.warn("Cached model at {} differs from the code in {}; "\
+                             "recompiling model".format(cached_path, path))
             recompile = True
             continue
 
         else:
-            logging.info("Using pre-compiled model from {}".format(cached_path)) 
+            if verbose:
+                logging.info("Using pre-compiled model from {}".format(cached_path)) 
             break
 
     else:

### 2018-06-30 article/ms.tex
things

@@
 
-\title{Detection and characterisation of stellar multiplicity with Gaia}
+\title{Detection and partial characterisation of stellar multiplicity with Gaia}
 
@@
 		     radial velocities to be candidate SB2 systems.}
-    \label{fig:radial_velocity_completeness}
+    \label{fig:sb2_rvs_completeness}
 \end{figure*}
@@
 
-% Figure: radial velocity completeness as a function of pertinent properties
-% 		  with the range that we have said is "fully complete"
-
-% Figure: h-r diagram of SB2 binary fraction
+\begin{figure*}
+	\includegraphics[width=1.0\textwidth,height=3.0in]{../figures/todo.png}
+    \caption{Hertzsprung-Russell diagram of X \Gaia\ sources that are in the
+    		 ranges specified in Equation \ref{eq:rvs_parameter_ranges} (gray).
+		 	 Overlaid in colour we show the SB2 candidates we deduce by the
+			 absence of reported radial velocity.}
+    \label{fig:sb2_hrd}
+\end{figure*}
 
@@
 % 		  some SB2s in some parameter range?
-
-% Figure: radial veloci
 
@@
 
-\todo{selecting similar sources}
+\todo{Bayes factors for each dimension?}
 
-\todo{fitting the two-component model}
+\todo{source properties: semi-amplitude (easy)}
 
-\todo{evaluating probability}
+\todo{Can we recover period and eccentricity using The Joker?}
 
-\todo{calculating source properties}
-
-In principle for well-sampled circular orbits this model would allow us to
-identify binary star systems with orbital periods at least up to twice the observing
-span in the second \Gaia\ data release ($\approx$21 months observing span; 
-allowing for well-characterised detections for systems with orbital periods up to $\approx$44 months).
-
-For higher-order star systems with longer periods, we can still find that a
-stellar multiple system is more likely than a single star scenario, but our
-estimate of the orbital properties (e.g., the semi-amplitude $K$) will be biased
-to lower values because we have not fully sampled half of the orbital period.
-At even longer orbital periods, or for stellar multiples with mass ratios that
-produce a small velocity semi-amplitude $K$, under our assumptions these systems 
-may be classified as single-star systems because the intrinsic radial velocity variation is within the expected 
-variations for single stars of a similar apparent magnitude and colour. We 
-revisit this issue in Section \ref{sec:sb_limits}, where we explicitly define 
-the probability that a \Gaia\ source is a SB1-type system $p(\textrm{SB1}|y)$ 
-within defined limits of orbital period (and other orbital and source parameters). 
-Outside of this parameter range, we are insensitive to distinguishing single-star 
-systems from higher-order SB1-type systems.
-
-% Model fitting.
-
-% Figure: fitting the radial velocity variance as a function of stellar properties.
-
-% Figures: Binarity as a function of source properties.
-
-% Figure: h-r diagram of SB1 binary fraction
-
-% Figure: histogram of binaries as a fucntion of the Rv excess
-
-ty fraction as a function of galactic latitude and
-% 		  longitude
+%For higher-order star systems with longer periods, we can still find that a
+%stellar multiple system is more likely than a single star scenario, but our
+%estimate of the orbital properties (e.g., the semi-amplitude $K$) will be biased
+%to lower values because we have not fully sampled half of the orbital period.
+%At even longer orbital periods, or for stellar multiples with mass ratios that
+%produce a small velocity semi-amplitude $K$, under our assumptions these systems 
+%may be classified as single-star systems because the intrinsic radial velocity variation is within the expected 
+%variations for single stars of a similar apparent magnitude and colour. We 
+%revisit this issue in Section \ref{sec:sb_limits}, where we explicitly define 
+%the probability that a \Gaia\ source is a SB1-type system $p(\textrm{SB1}|y)$ 
+%within defined limits of orbital period (and other orbital and source parameters). 
+%Outside of this parameter range, we are insensitive to distinguishing single-star 
+%systems from higher-order SB1-type systems.
 
@@
 
-\begin{figure*}
-	\includegraphics[width=1.0\textwidth]{../figures/todo.png}
-    \caption{Source radial velocity variance shown with respect to flux in
-    the \Gaia\ BP, G, and RP bands. The source radial velocity variance 
-    most closely follows the RP flux as it is closer to the \Gaia\ RVS
-    band.}
-\end{figure*}
+\section{Results}
 
-\begin{figure}
-	\includegraphics[width=0.5\textwidth]{../figures/todo.png}
-    \caption{Source mean \Gaia\ RP flux and radial velocity variance.
-    Points are represented by their mean $BP-RP$ color, confirming that
-    there is some color dependency on the source radial velocity variance.}
-    \label{fig:rp-flux-color}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=0.5\textwidth]{../figures/todo.png}
-    \caption{Mixture model fitted.}
-    \label{fig:mixture-model}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=0.5\textwidth]{../figures/todo.png}
-    \caption{Detected binary fraction as a function of the mean RP flux
-    (top panels), the $BP-RP$ color (second panels), the number of 
-    radial velocity transits (third panels), and the inferred temperature
-    (bottom panels). No trend is observed with mean RP flux or the number 
-    of radial velocity transits. However, our model finds bluer (hotter)
-    stars are more likely to be binaries, despite including color as a
-    dependency in our model.}
-    \label{fig:binary-fraction-wrt-params}
-\end{figure}
+\section{Discussion} \label{sec:discussion}
 
 \begin{itemize}
-
-\item CALIBRATION SAMPLE (including ADQL query)
-
-\item Fitting the model:
-
-\item Comparisons of residuals with other stellar properties.
-
-\item What is the RV error that Gaia can measure in a single epoch for a given magnitude vs colour, etc.
-
-\item What are the distributions of orbital parameters of binary systems that we would be able to detect?
-
-\item Calculating a probability of binarity for every star, taking the non-measurements into account!
-
-\item Completeness cuts,.... selection effects?
-
+	\item \todo{Binary fraction in the spaces that we were fitting: rp flux, colour, etc, just showing the transition of  probabilities}
+	\item \todo{Binary fraction as a function of fitting properties (e.g., colour, absolute RP mag, apparent RP mag)}
+	\item \todo{Binarity across the H-R diagram}
+	\item \todo{What are the distributions of orbital parameters of binary systems that we would be able to detect?}
 \end{itemize}
-
-\section{Results}
-
-\begin{itemize}
-
-\item Binary fraction in the spaces that we were fitting: rp flux, colour, etc, just showing the transition of  probabilities
-
-\item Binary fraction as a function of fitting properties, and other stellar properties from Gaia
-
-\item Binarity across the H-R diagram
-
-\item Binarity with metallicity 
-\end{itemize}
-
-\section{Discussion} \label{sec:discussion}
 
@@
 \begin{itemize}
-    \item \todo{binarity with metallicity}
-    \item \todo{binarity across the H-R diagram}
+	\item \todo{Binarity with metallicity }
     \item \todo{binarity in clusters vs the field?}
@@
 
+\appendix
+\section{Reproducibility}
+This project was developed in a \texttt{git} repository that is publicly
+accessible at \url{https://github.com/andycasey/velociraptor}. The repository
+includes notebooks that detail our work, \LaTeX\ to compile this manuscript, 
+and scripts to reproduce our analysis in its entirety. Reproducing this work in 
+full will require at least \todo{X}\,Gb of disk space and \todo{Y}\,compute hours.
+To reproduce this work (including data retrieval, analysis, production of figures,
+and manuscript compilation) use the \todo{following commands on a modern terminal}:
+
+\section{Relevant Astronomical Data Query Language (ADQL) queries}
+\todo{Include relevant ADQL queries}
+
+\section{Radial velocity completeness as a function of \Gaia\ source properties}
+\todo{Figures}
+
 \software{

### 2018-07-01 npm_run.py
finally

@@ -1,0 +1,148 @@
+
+""" Set up and run the non-parametric model. """
+
+import numpy as np
+import os
+import multiprocessing as mp
+import yaml
+from astropy.io import fits
+import tqdm
+
+import npm_utils as npm
+import stan_utils as stan
+
+with open(npm.CONFIG_PATH, "r") as fp:
+    config = yaml.load(fp)
+
+# Load in the data.
+data = fits.open(config["data_path"])[1].data
+
+all_label_names = list(config["kdtree_label_names"]) \
+                + list(config["predictor_label_names"])
+
+# Set up a KD-tree.
+X = np.vstack([data[ln] for ln in config["kdtree_label_names"]]).T
+finite = np.all([np.isfinite(data[ln]) for ln in all_label_names], axis=0)
+finite_indices = np.where(finite)[0]
+
+N, D = X.shape
+F = finite_indices.size
+L = 4 * len(config["predictor_label_names"]) + 1
+C = config.get("share_optimised_result_with_nearest", 0)
+
+kdt, scale, offset = npm.build_kdtree(X[finite], 
+    relative_scales=config.get("kdtree_relative_scales", None))
+
+kdt_kwds = dict(offset=offset, scale=scale, full_output=False)
+kdt_kwds.update(
+    minimum_radius=config.get("kdtree_minimum_radius", None), # DEFAULT
+    minimum_points=config.get("kdtree_minimum_points", 1024), # DEFAULT
+    maximum_points=config.get("kdtree_maximum_points", 8192)) # DEFAULT
+
+model = stan.load_stan_model(config["model_path"], verbose=False)
+
+default_opt_kwds = dict(
+    verbose=False,
+    tol_obj=7./3 - 4./3 - 1, # machine precision
+    tol_grad=7./3 - 4./3 - 1, # machine precision
+    tol_rel_grad=1e3,
+    tol_rel_obj=1e4,
+    iter=10000)
+default_opt_kwds.update(config.get("optimisation_kwds", {}))
+
+# Make sure that some entries have the right units.
+for key in ("tol_obj", "tol_grad", "tol_rel_grad", "tol_rel_obj"):
+    if key in default_opt_kwds:
+        default_opt_kwds[key] = float(default_opt_kwds[key])
+
+
+done = np.zeros(N, dtype=bool)
+results = np.nan * np.ones((N, L), dtype=float)
+
+
+
+def optimize_mixture_model(index, init=None):
+
+    # Select indices and get data.
[…]

### 2018-07-01 npm_utils.py

@@ -261,7 +261,7 @@ def ln_likelihood(y, theta, s_mu, s_sigma, b_mu, b_sigma):
     ll = np.sum(s_lpdf) + np.sum(b_lpdf)
     #print(lpdf)
     
-    assert np.isfinite(ll)
+    #assert np.isfinite(ll)
     return ll
 
 
@@ -308,6 +308,10 @@ def _pack_params(theta, mu_single, sigma_single, mu_multiple, sigma_multiple, **
     return np.hstack([theta, mu_single, sigma_single, mu_multiple, sigma_multiple])
 
 
+def _check_params_dict(d):
+    for k in ("mu_single", "sigma_single", "mu_multiple", "sigma_multiple", "mu_multiple_uv"):
+        d[k] = np.atleast_1d(d[k])
+    return d
 
 
 def nlp(params, y, L):

### 2018-07-01 stan_utils.py

@@ -79,20 +79,15 @@ def sampling_kwds(**kwargs):
 
 
 class suppress_output(object):
-    '''
-    A context manager for doing a "deep suppression" of stdout and stderr in
-    Python, i.e. will suppress all print, even if the print originates in a
-    compiled C/Fortran sub-function.
-       This will not suppress raised exceptions, since exceptions are printed
-    to stderr just before a script exits, and after the context manager has
-    exited (at least, I think that is why it lets exceptions through).
+    """ Suppress all stdout and stderr. """
 
-    '''
     def __init__(self):
-        # Open a pair of null files
-        self.null_fds = [os.open(os.devnull, os.O_RDWR) for x in range(2)]
+        self.null_fds = [
+            os.open(os.devnull, os.O_RDWR),
+            os.open(os.devnull, os.O_RDWR)
+        ]
         # Save the actual stdout (1) and stderr (2) file descriptors.
-        self.save_fds = (os.dup(1), os.dup(2))
+        self.save_fds = [os.dup(1), os.dup(2)]
 
     def __enter__(self):
         # Assign the null pointers to stdout and stderr.
@@ -103,6 +98,7 @@ class suppress_output(object):
         # Re-assign the real stdout/stderr back to (1) and (2)
         os.dup2(self.save_fds[0], 1)
         os.dup2(self.save_fds[1], 2)
-        # Close the null files
-        os.close(self.null_fds[0])
-        os.close(self.null_fds[1])
+
+        # Close the null files and descriptors.
+        for fd in self.save_fds + self.null_fds:
+            os.close(fd)

### 2018-07-01 npm_run.py
things

@@ -129,7 +129,7 @@ def swarm(*indices, queue=None):
 
 
 
-P = 8 # mp.cpu_count()
+P = 4 # mp.cpu_count()
 
 with mp.Pool(processes=P) as pool:
 

### 2018-07-01 npm_mr_run.py
clean up old files

@@ -1,283 +1,0 @@
-
-import pickle
-import logging
-import numpy as np
-import multiprocessing as mp
-import tqdm
-import os
-import io
-import gc
-import sys
-import yaml
-import warnings
-import resource
-from time import time, sleep
-from astropy.io import fits
-
-import npm_utils as npm
-import stan_utils as stan
-
-
-VERBOSE = False
-
-logging.info("RLIMIT_NOFILE = {}".format(resource.getrlimit(resource.RLIMIT_NOFILE)))
-
-try:
-    resource.setrlimit(resource.RLIMIT_NOFILE, (2048, -1))
-
-except ValueError:
-    logging.exception("Could not update RLIMIT_NOFILE")
-
-warnings.filterwarnings("ignore")
-
-with open(npm.CONFIG_PATH, "r") as config_fp:
-    config = yaml.load(config_fp)
-
-
-# Load the data.
-with fits.open(config["data_path"]) as image:
-    data = image[1].data
-
-    all_label_names = list(config["kdtree_label_names"]) \
-                    + list(config["predictor_label_names"])
-
-    # Set up a KD-tree.
-    X = np.vstack([data[ln] for ln in config["kdtree_label_names"]]).T
-    finite = np.where(np.all([np.isfinite(data[ln]) for ln in all_label_names], axis=0))[0]
-
-N, _ = X.shape
-F = finite.size
-L = 4 * len(config["predictor_label_names"]) + 1
-
-indices_kwds = dict(filename=config["indices_path"], dtype=np.int32)
-indices_bpr = config["kdtree_maximum_points"] \
-            * np.dtype(indices_kwds["dtype"]).itemsize
-
-results_kwds = dict(filename=config["results_path"], dtype=np.float32)
-results_bpr = L * np.dtype(results_kwds["dtype"]).itemsize
-
-# Create a results file.
-if not os.path.exists(results_kwds["filename"]):
-    logging.info("Creating results file: {}".format(results_kwds["filename"]))
-    fp = np.memmap(mode="w+", shape=(N, L), **results_kwds)
-    del fp
-    
-            
-
-def initialize_from_nearby_point(indices, ignore_index):
-    """
-    Look for an initialization point from nearby results.
-    """
-
-    for index in indices:
-        if index == ignore_index: continue
-        print("Opening at index {}".format(index))
-
-        #fp = np.memmap(results_fd, mode="r", shape=(L, ), 
[…]

### 2018-07-01 npm_run.py

@@ -7,6 +7,8 @@ import multiprocessing as mp
 import yaml
 from astropy.io import fits
 import tqdm
+import logging
+from time import sleep
 
 import npm_utils as npm
 import stan_utils as stan
@@ -57,14 +59,23 @@ for key in ("tol_obj", "tol_grad", "tol_rel_grad", "tol_rel_obj"):
 
 
 done = np.zeros(N, dtype=bool)
+queued = np.zeros(N, dtype=bool)
 results = np.nan * np.ones((N, L), dtype=float)
 
+indices_bpr = config["kdtree_maximum_points"] * np.dtype(np.int32).itemsize
 
 
 def optimize_mixture_model(index, init=None):
 
     # Select indices and get data.
     indices = finite_indices[npm.query_around_point(kdt, X[index], **kdt_kwds)]
+    """
+    with open(config["indices_path"], "br") as indices_fd:
+        indices_fd.seek(indices_bpr * index)
+        indices = np.fromfile(indices_fd, dtype=np.int32,
+                              count=config["kdtree_maximum_points"])
+    """
+
     y = np.array([data[ln][indices] for ln in config["predictor_label_names"]]).T
     
     if init is None:
@@ -85,64 +96,150 @@ def optimize_mixture_model(index, init=None):
 
 
 
-def swarm(*indices, queue=None):
+def swarm(*indices, max_random_starts=3, in_queue=None, candidate_queue=None, 
+    out_queue=None):
 
-    init = [None]
-    for index in indices:
+    def _random_index():
+        yield from np.random.choice(indices, max_random_starts, replace=False)
 
-        while True:
-            _, result, kdt_indices = optimize_mixture_model(index, init.pop(0))
+    _ri = _random_index()
+    random_start = lambda *_: (_ri.__next__(), None)
 
-            if queue is not None:
-                queue.put(1)
+    swarm = True
 
-            # Check to see which kdt_indices have not been done.
-            done[index] = True
+    while swarm:
 
-            if result is None:
-                # We will say this point is 'done', even though result failed.
-                break
+        for func in (in_queue.get_nowait, random_start):
+            try:
+                index, init = func()
 
-            results[index] = npm._pack_params(**result)
+            except mp.queues.Empty:
+                logging.info("Using a random index to start")
+                continue
 
-            if C > 0:
-                # Assign the closest points to have the same result.
-                B = sum(done[kdt_indices[:C + 1]])
[…]

### 2018-07-02 article/ms.tex
stuff

@@
 \begin{figure*}
-	\includegraphics[width=1.0\textwidth]{../figures/todo.png}
+	\includegraphics[width=1.0\textwidth]{../figures/sb2_rvs_completeness.pdf}
     \caption{Fraction of \Gaia\ sources with reported radial velocities
@@
 
-\begin{figure*}
-	\includegraphics[width=1.0\textwidth,height=3.0in]{../figures/todo.png}
+\begin{figure}
+	\includegraphics[width=0.4\textwidth]{../figures/sb2_histograms.pdf}
     \caption{Distribution of 
-    			(a) photometric \bp\ - \rp\ excess factor, 
-				(b) astrometric unit weight error, 
+    			(a) astrometric unit weight error (see Eq. \ref{eq:astrometric_unit_weight_error}),
+			%$u = \left(\chi_{al}^2/(N_{al} - 5)\right)^{1/2}$
+			%		where $\chi_{al}^2$ and $N_{al}$ is the astrometric goodness-of-fit and number 
+			%		of observations in the along-scan direction, respectively (columns 
+			%			\texttt{astrometric\_chi2\_al} and \texttt{astrometric\_n\_obs\_al}), 
+				(b) photometric \bp\ - \rp\ excess factor, 				
 	 				and 
-				(c) photometric variability
-			of candidate SB2 type systems relative to a control sample.}
+				(c) photometric \rp\ variability (see Eq. \ref{eq:photometric_variability})
+			of candidate SB2 type systems relative to a control sample of the same size.}
     \label{fig:sb2_histograms}
-\end{figure*}
+\end{figure}
 
 \begin{figure*}
-	\includegraphics[width=1.0\textwidth,height=3.0in]{../figures/todo.png}
-    \caption{Hertzsprung-Russell diagram of X \Gaia\ sources that are in the
-    		 ranges specified in Equation \ref{eq:rvs_parameter_ranges} (gray).
-		 	 Overlaid in colour we show the SB2 candidates we deduce by the
-			 absence of reported radial velocity.}
-    \label{fig:sb2_hrd}
+	\includegraphics[width=1.0\textwidth]{../figures/sb2_sky_structure.pdf}
+    \caption{Fraction of sources (per sky bin) without reported radial velocities
+    		 for all sources with $G \lesssim 13$ (top) and sources in the ranges
+		 	 specified by Eqs \ref{eq:sb2_ranges} (bottom), where we assert that a missing
+			 radial velocity measurement signifies a likely SB2 candidate . The color scale is 
+			 arbitrarily set to highlight structure, where black indicates a higher
+			 fraction of sources do not have reported radial velocities. Crowding
+			 in the galactic plane likely results in some sources not having radial
+			 velocities reported. The large scale structure visible in both axes is
+			 a combined effect of the initial \Gaia\ source list, the scanning law,
+			 and star forming regions (i.e., where emission in the \ion{Ca}{2} 
+			 triplet likely causes issues for radial velocity determination).}
+    \label{fig:sb2_sky_structure}
 \end{figure*}
@@
 \begin{eqnarray}
-	u = \sqrt{\frac{\chi^2_{al}}{N_{obs,al} - 5}}
+	u = \sqrt{\frac{\chi^2_{al}}{N_{obs,al} - 5}} \label{eq:astrometric_unit_weight_error}
 \end{eqnarray}
@@
 \subsection{Photometric binaries} \label{sec:method_photometric_binaries}
+
+\todo{bp rp excess factor}
+
+\begin{eqnarray}
+	\textrm{Amp} = \log_{10}\left(\sqrt{N_{obs,al}}\frac{\sigma_{\langle{}f_G\rangle}}{\langle{}f_G\rangle}\right) \label{eq:photometric_variability}
+\end{eqnarray}
 

### 2018-07-02 calculate_rv_completeness.py

@@ -1,139 +1,0 @@
-
-"""
-Calculate the completeness of radial velocity measurements as a function of
-various properties so that we can gauge when we should use the *lack* of radial
-velocity as an indicator of binarity.
-"""
-
-import pickle
-import numpy as np
-from astropy.io import fits
-from collections import Counter
-
-N_bins = 50
-
-semilogx = [True, False]
-equidensity = [True, False]
-
-completeness = {
-    "N_bins": N_bins,
-    "semilogx": np.repeat(semilogx, 2),
-    "equidensity": np.tile(equidensity, 2),
-    "completeness": dict()
-}
-
-image = fits.open("data/rv-completeness.fits")
-data = image[1].data
-
-has_rv = np.isfinite(data["radial_velocity"])
-
-label_names = (
- 'ra',
- 'dec',
- 'parallax',
- 'pmra',
- 'pmdec',
- 'astrometric_gof_al',
- 'astrometric_chi2_al',
- 'astrometric_weight_al',
- 'astrometric_pseudo_colour',
- 'mean_varpi_factor_al',
- 'astrometric_matched_observations',
- 'visibility_periods_used',
- 'astrometric_sigma5d_max',
- 'phot_g_mean_flux',
- 'phot_g_mean_mag',
- 'phot_bp_mean_flux',
- 'phot_bp_mean_mag',
- 'phot_rp_mean_flux',
- 'phot_rp_mean_mag',
- 'phot_bp_rp_excess_factor',
- 'bp_rp',
- 'bp_g',
- 'g_rp',
- 'radial_velocity',
- 'radial_velocity_error',
- 'rv_template_teff',
- 'rv_template_logg',
- 'rv_template_fe_h',
- 'l',
- 'b',
- 'ecl_lon',
- 'ecl_lat',
- 'teff_val',
- 'a_g_val',
- 'radius_val',
- 'lum_val'
-)
-
-for label_name in label_names:
-
-    print(label_name)
-    if label_name in completeness["completeness"]: continue
-
-    x = image[1].data[label_name]
-    x_min, x_max = (np.nanmin(x), np.nanmax(x))
-
-    plot_bins = np.zeros((4, N_bins), dtype=float)
-    plot_x = np.zeros((4, N_bins - 1), dtype=float)
-    plot_y = np.zeros_like(plot_x)
-    plot_y_err = np.zeros_like(plot_x)
-
-    completeness["completeness"][label_name] = dict()
-
[…]

### 2018-07-02 mpl_utils.py

@@ -33,7 +33,7 @@ mpl_style = {
 
     # Axes
     'axes.facecolor': '#FFFFFF',
-    'axes.edgecolor': '#AAAAAA',
+    'axes.edgecolor': '#333333',
     'axes.linewidth': 1.0,
     'axes.grid': False,
     'axes.titlesize': 'x-large',
@@ -46,13 +46,13 @@ mpl_style = {
     'xtick.minor.size': 4,
     'xtick.major.pad': 6,
     'xtick.minor.pad': 6,
-    'xtick.color': '#565656',
+    'xtick.color': '#333333',
     'xtick.direction': 'in',
     'ytick.major.size': 8,
     'ytick.minor.size': 4,
     'ytick.major.pad': 6,
     'ytick.minor.pad': 6,
-    'ytick.color': '#565656',
+    'ytick.color': '#333333',
     'ytick.direction': 'in',
     'xtick.labelsize': 'medium',
     'ytick.labelsize': 'medium',
@@ -74,7 +74,7 @@ mpl_style = {
 
 
 
-def plot_histogram_steps(ax, x_bins, y, y_err):
+def plot_histogram_steps(ax, x_bins, y, y_err, **kwargs):
 
     xx = np.array(x_bins).repeat(2)[1:]
     xstep = np.repeat((x_bins[1:] - x_bins[:-1]), 2)
@@ -86,7 +86,7 @@ def plot_histogram_steps(ax, x_bins, y, y_err):
 
     #xbc = x_bins[:-1] + 0.5 * np.diff(x_bins)
 
-    _ = ax.plot(xx, yy, '-')
+    _ = ax.plot(xx, yy, '-', **kwargs)
     ax.errorbar(x_bins, y, y_err, fmt=None, capsize=0, ecolor=_[0].get_color())
 
     return (xx, yy)

### 2018-07-02 plot_sb2.py

@@ -1,0 +1,420 @@
+
+"""
+Selection and analysis of SB2 type binary stars.
+
+See notebooks/sb2.ipynb for inspiration.
+
+"""
+
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.ticker import MaxNLocator
+from astropy.io import fits
+
+import velociraptor
+from mpl_utils import mpl_style
+plt.style.use(mpl_style)
+
+
+data = fits.open("data/gaia-sources-for-npm.fits")[1].data
+
+
+# Apply ranges
+valid_ranges = dict(
+    bp_rp=(1.2, 2.3),
+    phot_rp_mean_mag=(10, 12.2))
+
+
+def get_rv_completeness(x, semilogx, equidensity, N_bins,
+                        x_min=None, x_max=None):
+
+    y = data["radial_velocity"]
+    
+    x_finite = np.isfinite(x)
+    y_finite = np.isfinite(y)
+    if x_min is None:
+        x_min = np.min(x[x_finite])
+    if x_max is None:
+        x_max = np.max(x[x_finite])
+    
+    mask = (x_max >= x) * (x >= x_min)
+    x_finite *= mask
+    y_finite *= mask
+    
+    p = np.linspace(0, 100, N_bins)
+    if equidensity and semilogx:
+        bins = 10**np.percentile(np.log10(x[x_finite]), p)
+        
+    elif equidensity and not semilogx:
+        bins = np.percentile(x[x_finite], p)
+    
+    elif not equidensity and semilogx:
+        bins = np.logspace(np.log10(x_min), np.log10(x_max), N_bins)
+    
+    elif not equidensity and not semilogx:
+        bins = np.linspace(x_min, x_max, N_bins)
+    
+    numerator, _ = np.histogram(x[x_finite * y_finite], bins=bins)
+    denominator, _ = np.histogram(x[x_finite], bins=bins)
+    f = numerator/denominator.astype(float)
+    
+    # Pretty sure this ~has~ to be wrong.
+    f_err = f / np.diff(bins) * np.sqrt(
+            (np.sqrt(numerator)/numerator)**2 + \
+            (np.sqrt(denominator)/denominator)**2)
+
+    return (bins, f, f_err, numerator, denominator)
+
+
+def plot_rv_completeness(x, latex_label_name, semilogx, equidensity, N_bins, 
[…]

### 2018-07-03 kdt_min_density.py
things

@@ -1,0 +1,372 @@
+
+# coding: utf-8
+
+# ## Introduce minimum density constraints on the k-d ball
+# 
+
+# In[4]:
+
+
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+import pickle
+from astropy.io import fits
+from scipy.stats import binned_statistic_2d
+
+import sys
+sys.path.insert(0, "../")
+
+import velociraptor
+
+
+# In[5]:
+
+
+data = fits.open("data/gaia-sources-for-npm.fits")[1].data
+
+with open("results/rv_single_epoch_scatter.pkl", "rb") as fp:
+    rv_results = pickle.load(fp)
+    
+
+
+# In[6]:
+
+
+def plot_binned_statistic(x, y, z, bins=100, function=np.nanmedian,
+                          xlabel=None, ylabel=None, zlabel=None,
+                          ax=None, colorbar=False, figsize=(6, 6),
+                          zlog10=False,
+                          subsample=None, **kwargs):
+    
+    if ax is None:
+        fig, ax = plt.subplots(1, 1, figsize=figsize)
+    else:
+        fig = ax.figure
+    
+    if function not in ("count", "mean", "median"):
+        finite = np.isfinite(x * y * z)
+    else:
+        finite = np.arange(len(x))
+
+    if subsample is not None:
+        idx = np.where(finite)[0]
+        if subsample < 1:
+            subsample *= idx.size
+        finite = np.random.choice(idx, int(subsample), replace=False)
+    
+    H, xedges, yedges, binnumber = binned_statistic_2d(
+        x[finite], y[finite], z[finite],
+        statistic=function, bins=bins)
+    
+    if zlog10:
+        H = np.log10(H)
+    
+    imshow_kwds = dict(
+        aspect=np.ptp(xedges)/np.ptp(yedges), 
+        extent=(xedges[0], xedges[-1], yedges[-1], yedges[0]),
+        cmap="inferno")
+    imshow_kwds.update(kwargs)
+
+    image = ax.imshow(H.T, **imshow_kwds)
+    if colorbar:
+        cbar = plt.colorbar(image, ax=ax)
+        if zlabel is not None:
+            cbar.set_label(zlabel)
+
+    if xlabel is not None:
+        ax.set_xlabel(xlabel)
+    if ylabel is not None:
[…]

### 2018-07-03 npm_run.py

@@ -33,14 +33,16 @@ F = finite_indices.size
 L = 4 * len(config["predictor_label_names"]) + 1
 C = config.get("share_optimised_result_with_nearest", 0)
 
-kdt, scale, offset = npm.build_kdtree(X[finite], 
+kdt, scales, offsets = npm.build_kdtree(X[finite], 
     relative_scales=config.get("kdtree_relative_scales", None))
 
-kdt_kwds = dict(offset=offset, scale=scale, full_output=False)
+kdt_kwds = dict(offsets=offsets, scales=scales, full_output=False)
 kdt_kwds.update(
     minimum_radius=config.get("kdtree_minimum_radius", None), # DEFAULT
     minimum_points=config.get("kdtree_minimum_points", 1024), # DEFAULT
-    maximum_points=config.get("kdtree_maximum_points", 8192)) # DEFAULT
+    maximum_points=config.get("kdtree_maximum_points", 8192),
+    minimum_density=config.get("minimum_density", None)
+    ) # DEFAULT
 
 model = stan.load_stan_model(config["model_path"], verbose=False)
 
@@ -63,19 +65,13 @@ done = np.zeros(N, dtype=bool)
 queued = np.zeros(N, dtype=bool)
 results = np.nan * np.ones((N, L), dtype=float)
 
-indices_bpr = config["kdtree_maximum_points"] * np.dtype(np.int32).itemsize
+default_init = np.array([0.6985507, 3.0525788, 1.1474566, 1.8999833, 0.6495420])
 
 
 def optimize_mixture_model(index, init=None):
 
     # Select indices and get data.
     indices = finite_indices[npm.query_around_point(kdt, X[index], **kdt_kwds)]
-    """
-    with open(config["indices_path"], "br") as indices_fd:
-        indices_fd.seek(indices_bpr * index)
-        indices = np.fromfile(indices_fd, dtype=np.int32,
-                              count=config["kdtree_maximum_points"])
-    """
 
     y = np.array([data[ln][indices] for ln in config["predictor_label_names"]]).T
     
@@ -104,7 +100,7 @@ def swarm(*indices, max_random_starts=3, in_queue=None, candidate_queue=None,
         yield from np.random.choice(indices, max_random_starts, replace=False)
 
     _ri = _random_index()
-    random_start = lambda *_: (_ri.__next__(), None)
+    random_start = lambda *_: (_ri.__next__(), default_init)
 
[…]

### 2018-07-03 npm_utils.py

@@ -44,8 +44,11 @@ def build_kdtree(X, relative_scales=None,**kwargs):
     return (kdt, relative_scales, offset)
 
 
-def query_around_point(kdtree, point, offset=0, scale=1, minimum_radius=None, 
-    minimum_points=1, maximum_points=None, dualtree=False, full_output=False):
+
+
+def query_around_point(kdtree, point, offsets=0, scales=1, minimum_radius=None, 
+    minimum_points=1, maximum_points=None, minimum_density=None, dualtree=False,
+    full_output=False, **kwargs):
     """
     Query around a point in the KD-Tree until certain conditions are met (e.g.,
     the number of points in the ball, and the minimum radius that the ball
@@ -57,17 +60,28 @@ def query_around_point(kdtree, point, offset=0, scale=1, minimum_radius=None,
     :param point:
         The (unscaled) point to query around.
 
-    :param offset: [optional]
-        The offset to apply to the query point.
+    :param offsets: [optional]
+        The offsets to apply to the query point.
 
-    :param scale: [optional]
-        The scaling to apply to the query point, after subtracting the offset.
+    :param scales: [optional]
+        The scaling to apply to the query point, after subtracting the offsets.
+
+    :param minimum_radius: [optional]
+        The minimum radius (or radii) that the ball must extend to.
 
     :param minimum_points: [optional]
         The minimum number of points to return in the ball.
 
-    :param minimum_radius: [optional]
-        The minimum radius (or radii) that the ball must extend to.
+    :param maximum_points: [optional]
+        The maximum number of points to return in the ball. If the number of
+        points returned exceeds this value, then a random subset of the points
+        will be returned.
+
+    :param minimum_density: [optional]
+        The minimum average density of points per dimension for the ball. This
+        can be useful to ensure that points that are in the edge of the k-d tree
+        parameter space will be compared against points that are representative
+        of the underlying space, and not just compared against nearest outliers.
 
     :param dualtree: [optional]
         Use the dual tree formalism for the query: a tree is built for the query
@@ -79,67 +93,82 @@ def query_around_point(kdtree, point, offset=0, scale=1, minimum_radius=None,
         the indicies, otherwise just return the indices.
     """
 
-    point_orig = np.atleast_1d(point).reshape(1, -1)
-    point = (point_orig - offset)/scale
-    N, D = kdtree.data.shape
+    offsets = np.atleast_1d(offsets)
[…]

### 2018-07-03 article/ms.tex
bah

@@
 The properties and occurrence rate of stellar multiples (e.g., binaries, triples) 
-underpins much of astrophysics. Precisely measuring these quantities is challenging
-and observationally demanding. Here we use \Gaia\ data to detect and partially 
+underpins much of astrophysics. 
+%Precisely measuring these quantities is challenging and observationally demanding. 
+Here we use \Gaia\ data to detect and partially 
 characterise \NumberOfStellarMultiples\ stellar multiple systems based on excess 
-astrometric jitter, excess radial velocity jitter, the absence of reported radial
-velocities, photometry, and intrinsic luminosity.
-\todo{We show that astrometric and radial velocity jitter can be calibrated to
-estimate inclination angles (and directly constrain companion masses) for thousands 
-of systems.}
+jitter in astrometry and radial velocity.
 Comparisons with literature surveys on stellar multiplicity demonstrate that we 
 reliably detect binary systems with orbital periods up to $\approx$3.5\,yr. 
+We use eclipsing binary systems to calibrate excess astrometric jitter with
+tangential motion, allowing us to estimate inclination angles and directly constrain
+companion masses for millions of systems.
 \todo{A comment on stellar multiplicity with stellar metallicity.}
@@
 
-\subsection{Unresolved near-equal mass binaries: photometry}
-\label{sec:pb_methods}
+%\subsection{Unresolved near-equal mass binaries: photometry}
+%\label{sec:pb_methods}
 
-If a binary system is observed nearly face-on (at an inclination angle 
-$i \approx 0^\circ$) then there is will be no detectable excess radial 
-velocity variations. In principle there may be detectable astrometric
-variations, but most near-equal mass binaries would be detected more 
-reliably through intrinsic magnitudes that are anomalously lower (brighter)
-and bluer colours than what would be expected for a single star system.
-
-\todo{How the fuck am I going to model this, exactly?}
+%If a binary system is observed nearly face-on (at an inclination angle 
+%$i \approx 0^\circ$) then there is will be no detectable excess radial 
+%velocity variations. In principle there may be detectable astrometric
+%variations, but most near-equal mass binaries would be detected more 
+%reliably through intrinsic magnitudes that are anomalously lower (brighter)
+%and bluer colours than what would be expected for a single star system.
 

### 2018-07-03 kdt_min_density.py

@@ -1,372 +1,0 @@
-
-# coding: utf-8
-
-# ## Introduce minimum density constraints on the k-d ball
-# 
-
-# In[4]:
-
-
-import numpy as np
-import matplotlib
-import matplotlib.pyplot as plt
-import pickle
-from astropy.io import fits
-from scipy.stats import binned_statistic_2d
-
-import sys
-sys.path.insert(0, "../")
-
-import velociraptor
-
-
-# In[5]:
-
-
-data = fits.open("data/gaia-sources-for-npm.fits")[1].data
-
-with open("results/rv_single_epoch_scatter.pkl", "rb") as fp:
-    rv_results = pickle.load(fp)
-    
-
-
-# In[6]:
-
-
-def plot_binned_statistic(x, y, z, bins=100, function=np.nanmedian,
-                          xlabel=None, ylabel=None, zlabel=None,
-                          ax=None, colorbar=False, figsize=(6, 6),
-                          zlog10=False,
-                          subsample=None, **kwargs):
-    
-    if ax is None:
-        fig, ax = plt.subplots(1, 1, figsize=figsize)
-    else:
-        fig = ax.figure
-    
-    if function not in ("count", "mean", "median"):
-        finite = np.isfinite(x * y * z)
-    else:
-        finite = np.arange(len(x))
-
-    if subsample is not None:
-        idx = np.where(finite)[0]
-        if subsample < 1:
-            subsample *= idx.size
-        finite = np.random.choice(idx, int(subsample), replace=False)
-    
-    H, xedges, yedges, binnumber = binned_statistic_2d(
-        x[finite], y[finite], z[finite],
-        statistic=function, bins=bins)
-    
-    if zlog10:
-        H = np.log10(H)
-    
-    imshow_kwds = dict(
-        aspect=np.ptp(xedges)/np.ptp(yedges), 
-        extent=(xedges[0], xedges[-1], yedges[-1], yedges[0]),
-        cmap="inferno")
-    imshow_kwds.update(kwargs)
-
-    image = ax.imshow(H.T, **imshow_kwds)
-    if colorbar:
-        cbar = plt.colorbar(image, ax=ax)
-        if zlabel is not None:
-            cbar.set_label(zlabel)
-
-    if xlabel is not None:
-        ax.set_xlabel(xlabel)
-    if ylabel is not None:
[…]

### 2018-07-03 npm_utils.py

@@ -145,7 +145,7 @@ def query_around_point(kdtree, point, offsets=0, scales=1, minimum_radius=None,
             N, D = kdtree.data.shape
             left, right = (minimum_radius_norm, maximum_radius_norm)
 
-            Q = kwargs.get("Q", 1000) # MAGIC HACK
+            Q = kwargs.get("Q", 10) # MAGIC HACK
  
             # MAGIC HACK
             tolerance = maximum_points if maximum_points is not None \
@@ -153,6 +153,7 @@ def query_around_point(kdtree, point, offsets=0, scales=1, minimum_radius=None,
 
             while True:
                 # Shrink it.
+
                 ri = np.logspace(np.log10(left), np.log10(right), Q)
 
                 counts = kdtree.two_point_correlation(point, ri)
@@ -166,9 +167,8 @@ def query_around_point(kdtree, point, offsets=0, scales=1, minimum_radius=None,
                 left, right = (ri[indices[0] - 1], ri[indices[1]])
 
                 if np.diff(counts[indices]).max() < tolerance:
+                    radius_norm = ri[indices[0]]
                     break
-
-            radius_norm = left
 
         # two_point_correlation(point, minimum_radius_norm)
         #   is eequivalent to

### 2018-07-04 npm_run_elastic_ball_test.py
add results from elastic ball test

@@ -31,17 +31,20 @@ finite_indices = np.where(finite)[0]
 N, D = X.shape
 F = finite_indices.size
 L = 4 * len(config["predictor_label_names"]) + 1
-C = config["share_optimised_result_with_nearest"]
+C = config.get("share_optimised_result_with_nearest", 0)
 
-kdt, scales, offsets = npm.build_kdtree(
-    X[finite], relative_scales=config["kdtree_relative_scales"])
+kdt, scales, offsets = npm.build_kdtree(X[finite], 
+    relative_scales=config.get("kdtree_relative_scales", None))
 
 kdt_kwds = dict(offsets=offsets, scales=scales, full_output=False)
 kdt_kwds.update(
-    minimum_radius=config["kdtree_minimum_radius"],
-    minimum_points=config["kdtree_minimum_points"], 
-    maximum_points=config["kdtree_maximum_points"], 
-    minimum_density=config["kdtree_minimum_density"])
+    minimum_radius=config.get("kdtree_minimum_radius", None), # DEFAULT
+    minimum_points=config.get("kdtree_minimum_points", 1024), # DEFAULT
+    maximum_points=config.get("kdtree_maximum_points", 8192),
+    minimum_density=config.get("kdtree_minimum_density", None)
+    ) # DEFAULT
+
+logging.info("k-d tree keywords: {}".format(kdt_kwds))
 
 model = stan.load_stan_model(config["model_path"], verbose=False)
 
@@ -59,15 +62,22 @@ for key in ("tol_obj", "tol_grad", "tol_rel_grad", "tol_rel_obj"):
     if key in default_opt_kwds:
         default_opt_kwds[key] = float(default_opt_kwds[key])
 
-logging.info("k-d tree keywords: {}".format(kdt_kwds))
-logging.info("optimization keywords: {}"/format(default_opt_kwds))
-
 
 done = np.zeros(N, dtype=bool)
 queued = np.zeros(N, dtype=bool)
 results = np.nan * np.ones((N, L), dtype=float)
 
+# Set everything done except a box in parameter space.
+do = (3 > data["bp_rp"]) \
+   * (data["bp_rp"] > 2.5) \
+   * (data["absolute_rp_mag"] > -7.5) \
+   * (data["absolute_rp_mag"] < -2.5) \
+   * finite
+
+done[~do] = True
+
 default_init = np.array([0.6985507, 3.0525788, 1.1474566, 1.8999833, 0.6495420])
+
 
[…]

### 2018-07-04 npm_run.py
typo

@@ -60,7 +60,7 @@ for key in ("tol_obj", "tol_grad", "tol_rel_grad", "tol_rel_obj"):
         default_opt_kwds[key] = float(default_opt_kwds[key])
 
 logging.info("k-d tree keywords: {}".format(kdt_kwds))
-logging.info("optimization keywords: {}"/format(default_opt_kwds))
+logging.info("optimization keywords: {}".format(default_opt_kwds))
 
 
 done = np.zeros(N, dtype=bool)

### 2018-07-04 npm_run.py
typos

@@ -209,7 +209,7 @@ with mp.Pool(processes=P) as pool:
 
     j = []
     for _ in range(P):
-        j.append(pool.apply_async(swarm, do_indices, kwds=swarm_kwds))
+        j.append(pool.apply_async(mp_swarm, do_indices, kwds=swarm_kwds))
 
     # The swarm will just run at random initial points until we communicate
     # back that the candidates are good.

### 2018-08-20 scripts/npm_collect_the_battery_stars.py
add scripts

@@ -1,0 +1,38 @@
+import pickle
+import yaml
+import numpy as np
+from astropy.io import fits
+from astropy.table import Table
+
+
+with open("data/the-battery-stars.indices.pkl", "rb") as fp:
+    battery_star_indices = pickle.load(fp)
+
+
+with open("results/the-battery-stars.astrometric_unit_weight_error.pkl", "rb") as fp:
+    astrometric_results = pickle.load(fp)
+
+with open("results/the-battery-stars.rv_single_epoch_scatter.pkl", "rb") as fp:
+    rv_results = pickle.load(fp)
+
+
+with open("the-battery-stars.astrometry.yaml", "r") as fp:
+    config = yaml.load(fp)
+
+# Load in the data.
+data = fits.open(config["data_path"])[1].data
+
+
+subset = Table(data[battery_star_indices])
+
+keys = ("theta", "mu_single", "sigma_single", "mu_multi", "sigma_multi")
+
+for i, key in enumerate(keys):
+    subset["rv_{}".format(key)] = rv_results[battery_star_indices, i]
+    subset["astrometric_{}".format(key)] = astrometric_results[battery_star_indices, i]
+
+finite = np.isfinite(subset["astrometric_theta"] * subset["rv_theta"])
+subset = subset[finite]
+print(len(subset))
+
+subset.write("results/the-battery-stars.results.fits", overwrite=True)

### 2018-08-20 scripts/npm_collect_the_battery_stars_comparison.py

@@ -1,0 +1,41 @@
+import pickle
+import yaml
+import numpy as np
+from astropy.io import fits
+from astropy.table import Table
+
+
+
+with open("data/the-battery-stars-comparison.indices.pkl", "rb") as fp:
+    comp_indices = pickle.load(fp)
+
+
+
+
+with open("results/the-battery-stars-comparison.astrometric_unit_weight_error.pkl", "rb") as fp:
+    astrometric_results = pickle.load(fp)
+
+with open("results/the-battery-stars-comparison.rv_single_epoch_scatter.pkl", "rb") as fp:
+    rv_results = pickle.load(fp)
+
+
+with open("the-battery-stars.astrometry.yaml", "r") as fp:
+    config = yaml.load(fp)
+
+# Load in the data.
+data = fits.open(config["data_path"])[1].data
+
+
+subset = Table(data[comp_indices])
+
+keys = ("theta", "mu_single", "sigma_single", "mu_multi", "sigma_multi")
+
+for i, key in enumerate(keys):
+    subset["rv_{}".format(key)] = rv_results[comp_indices, i]
+    subset["astrometric_{}".format(key)] = astrometric_results[comp_indices, i]
+
+finite = np.isfinite(subset["astrometric_theta"] * subset["rv_theta"])
+subset = subset[finite]
+print(len(subset))
+
+subset.write("results/the-battery-stars-comparison.results.fits", overwrite=True)

### 2018-08-20 scripts/npm_run_battery_stars.py

@@ -1,0 +1,396 @@
+
+""" Set up and run the non-parametric model. """
+
+import numpy as np
+import os
+import multiprocessing as mp
+import pickle
+import yaml
+import tqdm
+import logging
+from time import sleep
+from astropy.io import fits
+
+import npm_utils as npm
+import stan_utils as stan
+
+with open("the-battery-stars.rv.yaml", "r") as fp:
+    config = yaml.load(fp)
+
+# Load in the data.
+data = fits.open(config["data_path"])[1].data
+
+all_label_names = list(config["kdtree_label_names"]) \
+                + list(config["predictor_label_names"])
+
+# Set up a KD-tree.
+X = np.vstack([data[ln] for ln in config["kdtree_label_names"]]).T
+finite = np.all([np.isfinite(data[ln]) for ln in all_label_names], axis=0)
+finite_indices = np.where(finite)[0]
+
+N, D = X.shape
+F = finite_indices.size
+L = 4 * len(config["predictor_label_names"]) + 1 # + 1 if using mu_multiple_uv
+C = config["share_optimised_result_with_nearest"]
+
+kdt, scales, offsets = npm.build_kdtree(
+    X[finite], relative_scales=config["kdtree_relative_scales"])
+
+kdt_kwds = dict(offsets=offsets, scales=scales, full_output=False)
+kdt_kwds.update(
+    minimum_radius=config["kdtree_minimum_radius"],
+    minimum_points=config["kdtree_minimum_points"],
+    maximum_points=config["kdtree_maximum_points"],
+    minimum_density=config.get("kdtree_minimum_density", None))
+
+model = stan.load_stan_model(config["model_path"], verbose=False)
+
+default_opt_kwds = config.get("optimisation_kwds", {})
+
+# Make sure that some entries have the right units.
+for key in ("tol_obj", "tol_grad", "tol_rel_grad", "tol_rel_obj"):
+    if key in default_opt_kwds:
+        default_opt_kwds[key] = float(default_opt_kwds[key])
+
+logging.info("k-d tree keywords: {}".format(kdt_kwds))
+logging.info("optimization keywords: {}".format(default_opt_kwds))
+
+
+done = np.zeros(N, dtype=bool)
+queued = np.zeros(N, dtype=bool)
+results = np.nan * np.ones((N, L), dtype=float)
+
+default_init = dict(zip(
[…]

### 2018-08-20 scripts/npm_run_battery_stars_comparison.py

@@ -1,0 +1,346 @@
+
+""" Set up and run the non-parametric model. """
+
+import numpy as np
+import os
+import multiprocessing as mp
+import pickle
+import yaml
+import tqdm
+import logging
+from time import sleep
+from astropy.io import fits
+
+import npm_utils as npm
+import stan_utils as stan
+
+with open("the-battery-stars.rv.yaml", "r") as fp:
+    config = yaml.load(fp)
+
+config["results_path"] = "results/the-battery-stars-comparison.rv_single_epoch_scatter.pkl"
+
+
+with open("the-battery-stars.astrometry.yaml", "r") as fp:
+    config = yaml.load(fp)
+config["results_path"] = "results/the-battery-stars-comparison.astrometric_unit_weight_error.pkl"
+
+# Load in the data.
+data = fits.open(config["data_path"])[1].data
+
+all_label_names = list(config["kdtree_label_names"]) \
+                + list(config["predictor_label_names"])
+
+# Set up a KD-tree.
+X = np.vstack([data[ln] for ln in config["kdtree_label_names"]]).T
+finite = np.all([np.isfinite(data[ln]) for ln in all_label_names], axis=0)
+finite_indices = np.where(finite)[0]
+
+N, D = X.shape
+F = finite_indices.size
+L = 4 * len(config["predictor_label_names"]) + 1 # + 1 if using mu_multiple_uv
+C = config["share_optimised_result_with_nearest"]
+
+kdt, scales, offsets = npm.build_kdtree(
+    X[finite], relative_scales=config["kdtree_relative_scales"])
+
+kdt_kwds = dict(offsets=offsets, scales=scales, full_output=False)
+kdt_kwds.update(
+    minimum_radius=config["kdtree_minimum_radius"],
+    minimum_points=config["kdtree_minimum_points"],
+    maximum_points=config["kdtree_maximum_points"],
+    minimum_density=config.get("kdtree_minimum_density", None))
+
+model = stan.load_stan_model(config["model_path"], verbose=False)
+
+default_opt_kwds = config.get("optimisation_kwds", {})
+
+# Make sure that some entries have the right units.
+for key in ("tol_obj", "tol_grad", "tol_rel_grad", "tol_rel_obj"):
+    if key in default_opt_kwds:
+        default_opt_kwds[key] = float(default_opt_kwds[key])
+
+logging.info("k-d tree keywords: {}".format(kdt_kwds))
[…]

### 2018-08-20 scripts/npm_run_carney_2003.py

@@ -1,0 +1,320 @@
+
+""" Set up and run the non-parametric model. """
+
+import numpy as np
+import os
+import multiprocessing as mp
+import pickle
+import yaml
+import tqdm
+import logging
+from time import sleep
+from astropy.io import fits
+
+import npm_utils as npm
+import stan_utils as stan
+
+with open("the-battery-stars.rv.yaml", "r") as fp:
+    config = yaml.load(fp)
+config["results_path"] = "results/carney-2003-singles.rv_single_epoch_scatter.pkl"
+
+
+with open("the-battery-stars.astrometry.yaml", "r") as fp:
+    config = yaml.load(fp)
+config["results_path"] = "results/carney-2003-singles.astrometric_unit_weight_error.pkl"
+
+
+# Load in the data.
+data = fits.open(config["data_path"])[1].data
+
+all_label_names = list(config["kdtree_label_names"]) \
+                + list(config["predictor_label_names"])
+
+# Set up a KD-tree.
+X = np.vstack([data[ln] for ln in config["kdtree_label_names"]]).T
+finite = np.all([np.isfinite(data[ln]) for ln in all_label_names], axis=0)
+finite_indices = np.where(finite)[0]
+
+N, D = X.shape
+F = finite_indices.size
+L = 4 * len(config["predictor_label_names"]) + 1 # + 1 if using mu_multiple_uv
+C = config["share_optimised_result_with_nearest"]
+
+kdt, scales, offsets = npm.build_kdtree(
+    X[finite], relative_scales=config["kdtree_relative_scales"])
+
+kdt_kwds = dict(offsets=offsets, scales=scales, full_output=False)
+kdt_kwds.update(
+    minimum_radius=config["kdtree_minimum_radius"],
+    minimum_points=config["kdtree_minimum_points"],
+    maximum_points=config["kdtree_maximum_points"],
+    minimum_density=config.get("kdtree_minimum_density", None))
+
+model = stan.load_stan_model(config["model_path"], verbose=False)
+
+default_opt_kwds = config.get("optimisation_kwds", {})
+
+# Make sure that some entries have the right units.
+for key in ("tol_obj", "tol_grad", "tol_rel_grad", "tol_rel_obj"):
+    if key in default_opt_kwds:
+        default_opt_kwds[key] = float(default_opt_kwds[key])
+
+logging.info("k-d tree keywords: {}".format(kdt_kwds))
[…]

### 2018-08-20 scripts/npm_run_soubiran_2013.py

@@ -1,0 +1,321 @@
+
+""" Set up and run the non-parametric model. """
+
+import numpy as np
+import os
+import multiprocessing as mp
+import pickle
+import yaml
+import tqdm
+import logging
+from time import sleep
+from astropy.io import fits
+
+import npm_utils as npm
+import stan_utils as stan
+
+
+with open("the-battery-stars.rv.yaml", "r") as fp:
+    config = yaml.load(fp)
+config["results_path"] = "results/Soubiran_2013.rv_single_epoch_scatter.pkl"
+
+with open("the-battery-stars.astrometry.yaml", "r") as fp:
+    config = yaml.load(fp)
+config["results_path"] = "results/Soubiran_2013.astrometric_unit_weight_error.pkl"
+
+# Load in the data.
+data = fits.open(config["data_path"])[1].data
+
+all_label_names = list(config["kdtree_label_names"]) \
+                + list(config["predictor_label_names"])
+
+# Set up a KD-tree.
+X = np.vstack([data[ln] for ln in config["kdtree_label_names"]]).T
+finite = np.all([np.isfinite(data[ln]) for ln in all_label_names], axis=0)
+finite_indices = np.where(finite)[0]
+
+N, D = X.shape
+F = finite_indices.size
+L = 4 * len(config["predictor_label_names"]) + 1 # + 1 if using mu_multiple_uv
+C = config["share_optimised_result_with_nearest"]
+
+kdt, scales, offsets = npm.build_kdtree(
+    X[finite], relative_scales=config["kdtree_relative_scales"])
+
+kdt_kwds = dict(offsets=offsets, scales=scales, full_output=False)
+kdt_kwds.update(
+    minimum_radius=config["kdtree_minimum_radius"],
+    minimum_points=config["kdtree_minimum_points"],
+    maximum_points=config["kdtree_maximum_points"],
+    minimum_density=config.get("kdtree_minimum_density", None))
+
+model = stan.load_stan_model(config["model_path"], verbose=False)
+
+default_opt_kwds = config.get("optimisation_kwds", {})
+
+# Make sure that some entries have the right units.
+for key in ("tol_obj", "tol_grad", "tol_rel_grad", "tol_rel_obj"):
+    if key in default_opt_kwds:
+        default_opt_kwds[key] = float(default_opt_kwds[key])
+
+logging.info("k-d tree keywords: {}".format(kdt_kwds))
[…]

### 2018-08-20 scripts/xmatch_soubiran_2013_with_gaia.py

@@ -1,0 +1,35 @@
+from astropy.table import Table
+from astroquery.gaia import Gaia
+
+
+soubiran_2013 = Table.read("data/Soubiran_2013.fits")
+N = len(soubiran_2013)
+
+# Get gaia source IDs and we can do the cross-match from there.
+source_ids = []
+for i, hip in enumerate(soubiran_2013["HIP"]):
+
+    job = Gaia.launch_job("""
+        SELECT gaia.source_id
+        FROM gaiadr2.gaia_source AS gaia,
+             gaiadr2.hipparcos2_best_neighbour AS hip
+        WHERE hip.original_ext_source_id = '{:.0f}'
+        AND   hip.source_id = gaia.source_id
+        """.format(hip))
+
+    results = job.get_results()
+    if len(results) == 0:
+        source_id = -1
+    else:
+        source_id = results["source_id"][0]
+
+    print("{}/{}: HIP {} = Gaia DR2 {}".format(i, N, hip, source_id))
+    source_ids.append(source_id)
+
+import numpy as np
+soubiran_2013["source_id"] = np.array(source_ids, dtype=np.int64)
+
+OK = np.isfinite(soubiran_2013["source_id"])
+soubiran_2013 = soubiran_2013[OK]
+
+soubiran_2013.write("data/Soubiran_2013-xm-Gaia.fits", overwrite=True)

### 2018-08-20 npm_run.py
updates

@@ -14,6 +14,10 @@ from astropy.io import fits
 import npm_utils as npm
 import stan_utils as stan
 
+DO_BALL_TEST = False
+
+np.random.seed(42)
+
 with open("npm-config.rv.yaml", "r") as fp:
     config = yaml.load(fp)
 
@@ -36,9 +40,10 @@ C = config["share_optimised_result_with_nearest"]
 kdt, scales, offsets = npm.build_kdtree(
     X[finite], relative_scales=config["kdtree_relative_scales"])
 
-kdt_kwds = dict(offsets=offsets, scales=scales, full_output=False)
+kdt_kwds = dict(offsets=offsets, scales=scales, full_output=True)
 kdt_kwds.update(
     minimum_radius=config["kdtree_minimum_radius"],
+    maximum_radius=config.get("kdtree_maximum_radius", None),
     minimum_points=config["kdtree_minimum_points"],
     maximum_points=config["kdtree_maximum_points"],
     minimum_density=config.get("kdtree_minimum_density", None))
@@ -60,12 +65,34 @@ done = np.zeros(N, dtype=bool)
 queued = np.zeros(N, dtype=bool)
 results = np.nan * np.ones((N, L), dtype=float)
 
+meta_names = ("N", "ball_medians", "ball_ptps", "y_percentiles", "max_log_y")
+M = 1 + 3 + 3 + 3 + 1
+meta_results = np.nan * np.ones((N, M), dtype=float)
+
 default_init = dict(zip(
     ("theta", "mu_single", "sigma_single", "mu_multiple", "sigma_multiple"),
     np.array([0.75, 1, 0.5, 1, 0.75])))
 default_init["mu_multiple_uv"] = 0.1
+
+"""
+# For the v8 test
+default_init = dict(zip(
+    ("theta", "mu_single", "sigma_single", "mu_multiple", "sigma_multiple"),
+    np.array([0.65, 2.25, 1.0, 1.6, 0.60])))
+default_init["mu_multiple_uv"] = 0.1 # not used
+"""
+
 default_init = npm._check_params_dict(default_init)
 bounds = config["parameter_bounds"]
+
+
+
+
+
+
+
+
+
 
 def optimize_mixture_model(index, init=None):
     # kdt, X, indices, kdt_kwds, data
@@ -75,23 +102,41 @@ def optimize_mixture_model(index, init=None):
     # bounds
     # model OR model_path
     # opt_kwds
-    
+
 
     # Select indices and get data.
[…]

### 2018-08-20 npm_utils.py

@@ -48,7 +48,8 @@ def build_kdtree(X, relative_scales=None,**kwargs):
 
 
 def query_around_point(kdtree, point, offsets=0, scales=1, minimum_radius=None, 
-    minimum_points=1, maximum_points=None, minimum_density=None, dualtree=False,
+    maximum_radius=None, minimum_points=1, maximum_points=None, 
+    minimum_density=None, dualtree=False,
     full_output=False, **kwargs):
     """
     Query around a point in the KD-Tree until certain conditions are met (e.g.,
@@ -94,6 +95,8 @@ def query_around_point(kdtree, point, offsets=0, scales=1, minimum_radius=None,
         the indicies, otherwise just return the indices.
     """
 
+    #print("querying k-d tree")
+
     offsets = np.atleast_1d(offsets)
     scales = np.atleast_1d(scales)
 
@@ -138,6 +141,9 @@ def query_around_point(kdtree, point, offsets=0, scales=1, minimum_radius=None,
             radius_norm = minimum_radius_norm
 
         else:
+            #print("Using k-dtree to step out in radius because we have {} points within {} but need {}".format(
+            #    K, minimum_radius_norm, K_min))
+
             # We need to use the k-d tree to step out until our constraints are
             # met.
             maximum_radius_norm = 2 * np.max(np.ptp(kdtree.data, axis=0))
@@ -157,7 +163,11 @@ def query_around_point(kdtree, point, offsets=0, scales=1, minimum_radius=None,
 
                 ri = np.logspace(np.log10(left), np.log10(right), Q)
 
+
                 counts = kdtree.two_point_correlation(point, ri)
+                #print("tolerance {}: {} {} {} {}".format(tolerance, left, right, ri, counts))
+
+
 
                 minimum_counts = np.clip(2 * np.max(np.dot(ri.reshape(-1, 1), 
                     (minimum_density * scales).reshape(1, -1)), axis=1),
@@ -165,7 +175,9 @@ def query_around_point(kdtree, point, offsets=0, scales=1, minimum_radius=None,
 
                 indices = np.arange(Q)[counts >= minimum_counts]
 
-                left, right = (ri[indices[0] - 1], ri[indices[1]])
+                left, right = (ri[indices[0] - 1], ri[indices[0] + 1])
+                #print("new: {} {}".format(left, right))
+
 
                 if np.diff(counts[indices]).max() < tolerance:
[…]

### 2018-08-20 plot_gp_models.py

@@ -1,6 +1,6 @@
-
 import matplotlib
 import matplotlib.pyplot as plt
+plt.switch_backend("agg")
 import numpy as np
 import os
 import pickle
@@ -142,7 +142,7 @@ in enumerate(zip(axes, gp_labels, gps)):
 
     kernel = construct_kernel(D)
 
-    #white_noise=-50, 
+    #white_noise=-50,
     gp = george.GP(kernel, mean=0.0, white_noise=0.0,
                    fit_mean=True, fit_white_noise=True)
     gp.set_parameter_vector(list(hp.values()))
@@ -188,7 +188,7 @@ for i, (hp, y) in enumerate(gps):
 
     kernel = construct_kernel(D)
 
-    #white_noise=-50, 
+    #white_noise=-50,
     gp = george.GP(kernel, mean=0.0, white_noise=0.0,
                    fit_mean=True, fit_white_noise=True)
     gp.set_parameter_vector(list(hp.values()))
@@ -249,8 +249,11 @@ tau_single, tau_multiple = np.exp(log_tau).T
 
 gpm_predictions[:, -1] = tau_single
 
+with open("tmp_gp.pkl", "wb") as fp:
+    pickle.dump(gpm_predictions, fp, -1)
 
-t = Table.read("data/gaia-sources-for-npm.fits")
+from astropy.table import Table
+t = Table.read("data/gaia-sources-for-npm-colsubset.fits")
 
 names = ("mu_single", "mu_single_var",
          "sigma_single", "sigma_single_var",

### 2018-08-20 stan_utils.py

@@ -78,8 +78,6 @@ def sampling_kwds(**kwargs):
     return kwds
 
 
-
-
 class suppress_output(object):
     """ Suppress all stdout and stderr. """
 
@@ -128,5 +126,4 @@ class suppress_output(object):
 
         for fd, p in self.null_fds:
             os.close(fd)
-
-            
+            os.unlink(p)

### 2018-08-20 .ci/install.sh
fix travis?

@@ -1,8 +1,8 @@
 #!/bin/bash -x
 
 # SSL workaround
-mkdir -p ~/.config/Tectonic
-cp .ci/tectonic.config.toml ~/.config/Tectonic/config.toml
+#mkdir -p ~/.config/Tectonic
+#cp .ci/tectonic.config.toml ~/.config/Tectonic/config.toml
 
 # Conda install
 wget https://repo.continuum.io/miniconda/Miniconda3-latest-Linux-x86_64.sh -O miniconda.sh;
@@ -12,5 +12,7 @@ hash -r
 conda config --set always_yes yes --set changeps1 no
 conda update -q conda
 conda info -a
-conda install -c conda-forge -c pkgw-forge tectonic
+conda config --add channels conda-forge
+conda update --all
+conda install tectonic
 tectonic --help

### 2018-08-21 article/figures/plot_rv_figures.py
add plotting script for rv figures

@@ -1,0 +1,878 @@
+import numpy as np
+import os
+import pickle
+from astropy.io import fits
+from astropy.table import Table
+
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.ticker import MaxNLocator
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+
+from mpl_utils import (mpl_style, plot_binned_statistic)
+
+plt.style.use(mpl_style)
+
+
+BASE_PATH = "../../"
+RELEASE_CANDIDATE_VERSION = 0
+
+MAKE_FIGURES = [
+    "rv_hrd_single_star_fraction_hist",
+    "rv_hrd_single_star_fraction_scatter",
+    "rv_sb9_kalpha_comparison",
+    "rv_sb9_k_comparison",
+    "rv_sb9_kalpha_comparison_period",
+    "rv_sb9_kp_corner",
+    "rv_apw_unimodal_k_comparison",
+    "rv_apw_unimodal_kalpha_comparison",
+    "rv_apw_percentiles_k_comparison",
+    "rv_gp_hrd",
+    "rv_gp_wrt_params",
+    "rv_soubiran_hist",
+    "rv_huang_hist"
+]
+
+DEFAULT_SEQUENTIAL_CMAP = "viridis"
+
+
+velociraptor = fits.open(os.path.join(
+    BASE_PATH,
+    "results", 
+    "velociraptor-catalog-rc.{:.0f}.fits".format(RELEASE_CANDIDATE_VERSION)))
+velociraptor = velociraptor[1].data
+
+latex_labels = dict(
+    bp_rp=r"\textrm{bp - rp}",
+    phot_rp_mean_mag=r"\textrm{apparent rp mag}",
+    absolute_rp_mag=r"\textrm{absolute rp mag}",
+    rv_mu_single=r"$\mu_s$ \textrm{/ km\,s}$^{-1}$",
+    rv_sigma_single=r"$\sigma_s$ \textrm{/ km\,s}$^{-1}$"
+)
+
+
+common_kwds = dict([
+    ("bp_rp.limits", (-0.25, 6.5)),
+    ("phot_rp_mean_mag.limits", (13.0, 2.1)),
+    ("absolute_rp_mag.limits", (10.5, -16)),
+])
+
+one_to_one_line_kwds = dict(c="#666666", linestyle=":", lw=1, zorder=-1)
+
+def savefig(fig, basename):
[…]

### 2018-08-21 article/ms.tex
filler figures

@@
 \usepackage{color}
+\usepackage{comment}
 
@@
 
+\author[0000-0003-3494-343X]{Carles Badenes}
+\affiliation{Department of Physics and Astronomy, 
+			 University of Pittsburgh, 
+			 3941 O'Hara Street, Pittsburgh, PA 15260, USA}
+
+\author{Rosemary Mardling}
+\author{John Lattanzio}
+
 \author[0000-0003-2866-9403]{David W. Hogg}
@@
 
-\author[0000-0003-3494-343X]{Carles Badenes}
-\affiliation{Department of Physics and Astronomy, 
-			 University of Pittsburgh, 
-			 3941 O'Hara Street, Pittsburgh, PA 15260, USA}
-
 \author[0000-0003-0872-7098]{Adrian M. Price-Whelan}
@@
 \begin{abstract}
-The properties and occurrence rate of stellar multiples (e.g., binaries, triples) 
-underpins much of astrophysics. 
-%Precisely measuring these quantities is challenging and observationally demanding. 
-Here we use \Gaia\ data to detect and partially 
-characterise \NumberOfStellarMultiples\ stellar multiple systems based on excess 
-jitter in astrometry and radial velocity.
-Comparisons with literature surveys on stellar multiplicity demonstrate that we 
-reliably detect binary systems with orbital periods up to $\approx$3.5\,yr. 
+The frequency and properties of stellar multiples (e.g., binaries, triples) underpin 
+much of astrophysics, and are difficult to estimate.  
+Here we use \Gaia\ data to detect and partially characterise \NumberOfStellarMultiples\
+stellar multiple systems based on excess jitter in astrometry and radial velocity.
+We reliably detect binary systems with orbital periods up to $\approx$3.5\,yr. % and systems with what kind of mass ratios, semi-amplitudes,  etc. 
 We use eclipsing binary systems to calibrate excess astrometric jitter with
@@
 companion masses for millions of systems.
+\todo{We find that the distribution of inclination angles is isotropic.}
 \todo{A comment on stellar multiplicity with stellar metallicity.}
@@
 
-Higher order stellar systems describes any stars that have at least one stellar 
+A higher order stellar system describes any star that has at least one stellar 
 companion (e.g., binaries, trinaries). The presence of stellar companions
-introduces a number of complications to the inference of stellar and galactic
-properties, but these factors are often ignored.
+complicates inferences on stellar and galactic properties, but these factors are
+often ignored because it is extremely challenging to detect stellar multiplicity
+for an unresolved system, and even harder to separate the contributions from individual
+sources.
 
 The \Gaia\ space telescope provides exquisite astrometry, photometry, and radial
-velocity measurements over many years for million of point stars in our galaxy.
-%These data provides an excellent opportunity to self-calibrate the errors in those
-%measurements. 
+velocity measurements over many years for million of point sources in our galaxy
+\citep{GaiaDR2}.
 When a point source has an unresolved stellar companion within some range of orbital
-and stellar parameters, it is expected that there will be an anomalous excess in 
-these measured properties relative to a single star of a similar type. This is
-nothing new: astronomers have used radial velocity \citep{Butler:1996} and 
-astrometric variations \citep{Muterspaugh:2010} and photometry \citep{someone} to infer the 
-presence of stellar and exoplanet companions forever. However, the opportunity
-presented by \Gaia\ is to infer the presence of stellar companions for millions
-of point sources, across both hemispheres. No ground- or space-based instrument
-has ever provided such an opportunity.
+and stellar parameters, it is expected that there will be an excess in motion jitter
+relative to a single star of a similar type. This is nothing new: astronomers have 
+used radial velocity \citep{Butler:1996} and astrometric variations \citep{Muterspaugh:2010}
+to infer the presence of stellar and exoplanet companions for decades. However, 
+the \Gaia\ data presents an opportunity to infer the presence of stellar companions
+for millions of point sources across both hemispheres. No ground- or space-based 
+instrument has ever provided such data.
 
-In this work we make use of photometry and astrometry (including radial motion)
-measurements in the second \Gaia\ data release to infer the presence of stellar
-companions for millions of point sources, and provide some basic characterisations
-of these systems. In Section \ref{sec:method} we describe our methods, and in
-Section \ref{sec:results} we outline our results in context of other stellar
-multiplicity surveys. We include comparisons of binary properties between
-clusters and the field, as a function of stellar properties (e.g., stellar
-metallicity), and highlight some particularly noteworthy systems that would
-benefit from immediate follow-up spectroscopic observations.
+\todo{Literature review on what we do and don't know about  the multiplicity fraction}
+
+In this work we make use of astrometry (including radial motion) \todo{and photometry?}
+from the second \Gaia\ data release to infer the presence of stellar companions 
+for millions of point sources. We provide partial characterisations of these systems,
+based on the data available for each source. In Section \ref{sec:method} we 
+describe our methods, and in Section \ref{sec:results} we outline our results 
+in context of other stellar multiplicity surveys. We include comparisons of 
+binary properties between clusters and the field, as a function of stellar 
+properties (e.g., stellar metallicity), and highlight some particularly 
+noteworthy systems that would benefit from additional observations.
+
+\section{Data} \label{sec:data}
+
+There are many ways that binaries can be reliably identified using \Gaia\
+data, including radial velocity measurements, astrometry, and photometry.
+These data are sensitive to characterising stellar multiplicity in 
+different ways. Here we describe the data that we use from the second
+\Gaia\ data release \citep{Lindegren:2018} before defining our model.
+
+%Here we define a model that we will use to detect and 
+%characterise stellar multiples from radial velocity and astrometric data,
+%before describing how the absence of \Gaia\ radial velocities can also be
+%used to infer stellar multiplicity.
+
+%Here we describe the absence of reported radial velocity in
+%\Gaia\ data can be used to deduce stellar multiples, before we describe
+%our model that incorporates measurements of radial velocity, astrometry, 
+%and photometry, to detect and characterise stellar multiples.
+
+\subsection{Radial velocities}
+\label{sec:data_vrad}
+
+Radial velocity measurements are not available for individual epochs in the 
+second \Gaia\ data release \citep{Lindegren:2018,Cropper:2018}, but a point estimate
+and an error in radial velocity is available for 7,224,631 sources \citep{Cropper:2018}.
+% TODO: That's the value from the paper, but I think the true value is lower. Check.
+The reported radial velocity error ($\GaiaRVE$; column name \texttt{radial\_velocity\_error}) 
+is a function of the number of transits $N$ (i.e., the number of observations; 
+given by column \texttt{rv\_nb\_transits}) and the standard deviation among 
+those measurements. This allows us to calculate the standard deviation in radial 
+velocity for each source, independent of the number of measurements,
+which we will refer to as the \emph{radial velocity jitter} $\RVJitter$,
+\begin{eqnarray}
+\RVJitter = \sqrt{\frac{2N}{\pi}}\GaiaRVE \quad .
+\end{eqnarray}
+
+For a single star, or a star without a significant mass companion, the source radial
+velocity jitter represents the minimum uncertainty with which \Gaia\ can measure 
+radial velocity for a star of that colour, apparent magnitude, and absolute magnitude.
+Stars with bluer colours have fewer absorption lines with deeper wings, and fainter stars
+have on average lower signal-to-noise (S/N) ratios, both of which result in noisier radial
+velocity measurements. The absolute magnitude is similarly important, as giant stars have
+narrow absorption lines than main-sequence stars of the same temperature and colour.
+For stellar multiples, the source radial velocity jitter will be the quadrature sum of the expected jitter for the primary (if it were a single star) and a contribution based on the orbital parameters  and the observation epochs.
+
+% todo: lack of RV measurements?
+
+\subsection{Astrometry} \label{sec:data_astrometry}
+
+Many higher-order star systems are unresolved point sources in \Gaia, but the
+stellar companion(s) may have sufficient mass to measurably affect the tangential 
+velocity of the primary and produce a detectable jitter in the astrometric 
+position. There are many other sources of astrometric jitter, but here we show 
+that the principle effect of astrometric excess noise is due to stellar 
+companions. Note that due to the astrometry `degree of freedom bug' in the 
+processing of the second \Gaia\ data release \citep[e.g., see Appendix A1 of ][]{Lindegren:2018}, 
+that resulted in 80\% of sources having zero astrometric excess noise,
+we instead make use of the astrometric unit weight error $u$
+\begin{eqnarray}
+	u = \sqrt{\frac{\chi^2_{al}}{N_{obs,al} - 5}} \label{eq:astrometric_unit_weight_error}
+\end{eqnarray}
+\noindent{}where $\chi^2_{al}$ is the astrometric $\chi^2$ value
+in the along-scan direction (column \texttt{astrometric\_chi2\_al}) and 
+$N_{obs,al}$ is the number of astrometric observations in the along-scan
+direction (column \texttt{astrometric\_n\_obs\_al}). The values of this
+quantity can be interpreted similarly to how the astrometric excess noise
+would be interpreted (e.g., large values are less consistent with a
+single star model), although like the radial velocity and astrometric
+jitter, it must be considered in context with stars of similar colour,
+apparent magnitude, and preferably, absolute magnitude.
+
+% todo: what ADQL queries did we do?
+% what will our subset be?
 
@@
 
-There are many ways that binaries can be reliably identified using \Gaia\
-data, including radial velocity measurements, photometry, and astrometry.
-These detection methods are sensitive to binary systems with different 
-properties. Here we describe the absence of reported radial velocity in
-\Gaia\ data can be used to deduce stellar multiples, before we describe
-our model that incorporates measurements of radial velocity, astrometry, 
-and photometry, to detect and characterise stellar multiples.
+We assume that for sources with similar \bp\ - \rp\ colours, apparent \rp\ magnitudes, and absolute \rp\ magnitudes, the jitter distribution (in radial velocity or astrometry) is a mixture of two populations:
+
+\begin{enumerate}
+\item \emph{single-star systems}, where the jitter represents the minimum noise that 
+	   \Gaia\ can measure for a star of that
+      \bp\ - \rp\ colour, apparent \rp\ magnitude, and absolute \rp\ magnitude;
+
+\item \emph{higher-order star systems}, where the jitter is
+      significantly above that for single-star systems of a similar colour, 
+      apparent magnitude, and absolute magnitude.
+\end{enumerate}
+
+We also assume that the mean single-star jitter will change
+depending on the \bp\ - \rp\ colour, the apparent magnitude, and the absolute magnitude.
+For example, the mean radial velocity jitter of single stars with $\teff \approx 6000\,\textrm{K}$
+will be higher than the jitter of single stars with $\teff \approx 5000\,\textrm{K}$, 
+with all else being (nearly) equal. We similarly assume that the mean 
+jitter of stellar \emph{multiples} will depend on the source colour and its position
+in the Hertzsprung-Russell diagram. For these reasons, in order to evaluate whether a
+source is more likely to be a single star or a stellar multiple, we must consider the 
+jitter in context with other sources that have similar: \bp\ - \rp\
+colour, apparent \rp\ magnitude, and absolute \rp\ magnitude. In principle if we knew
+the distribution stellar multiple properties (or assumed one) then we could model the
+observability of binary (and higher order) systems across the Hertszprung-Russell
+diagram, but this is beyond the scope of this work.
+
+% ADDITIONAL ASSUMPTIONS.
+\todo{additional assumptions}
+
+Given these assumptions, we chose to adopt a non-parametric model for stellar multiplicity,
+primarily because we do not know the distribution of properties of stellar multiples.
+Our approach is illustrated in Figure \ref{fig:npm_schematic} and can be qualitatively 
+described by the following steps.\\
+
+\noindent{}For each \Gaia\ source:
+\begin{enumerate}
+
+	\item Select $N$ sources that have a similar \bp\ - \rp\ colour, apparent \rp\
+		  magnitude, and absolute \rp\ magnitude.
+
+	\item Fit the distribution of radial velocity jitter for sources selected in
+		  Step 1 using a mixture of a normal and a log-normal distribution.
+
+	\item Evaluate whether the source of interest is more consistent with being
+		  a single star component (drawn from the normal distribution) or a
+		  stellar multiple (drawn from the log-normal distribution).
+
+	\item If the source is more likely to be a stellar multiple, use the estimated
+		  properties of single stars with similar properties to estimate the
+		  excess radial velocity jitter due to the stellar companion(s).
+
+\end{enumerate}
+
+\begin{figure*}
+	\includegraphics[width=1.0\textwidth]{../figures/todo.png}
+    \caption{Illustration of the method we adopt for identifying and characterising
+    		 stellar multiplicity across the Hertsprung-Russell diagram. For each
+		 	 \Gaia\ source we select between 128 and 1024 sources that have
+			 similar \bp\ - \rp\ colours, apparent \rp\ magnitudes, and absolute \rp\
+			 magnitudes (Step 1). With these similar sources we fit a two-component
+			 model to distinguish single star sources from stellar multiples (Step 2).
+			 With the model parameters optimised from Step 2, we evaluate probability
+			 of stellar multiplicity and partially characterise the higher order
+			 systems by assuming binarity and estimating the system semi-amplitude (Step 3).}
+    \label{fig:npm_schematic}
+\end{figure*}
+
+% todo: joining sentence(s) here?
+
+\subsection*{Step 1: Selecting similar sources}
+
+We construct a $k$-dimensional tree \citep[k-d tree;][]{someone} for all sources in the
+dimensions of \bp\ - \rp\ colour, apparent \rp\ magnitude, and absolute \rp\ magnitude.
+When doing so we scale the dimensions such that \todo{0.1}\,mags in \bp\ - \rp\ colour is
+approximately equal to \todo{1} magnitude in apparent \rp\ magnitude and \todo{1} magnitude in 
+absolute \rp\ magnitude. We use this $k$-d tree for selecting a `ball' of similar
+stars for a given source.
+
+We enforce a number of constraints when selecting similar stars. For each source
+we require the multi-dimensional `ball' to include at least \todo{128} sources.
+We further require that the radius of the ball extends to at least \todo{0.05}\,mags in 
+\bp\ - \rp\ colour, \todo{0.5}\,mags in apparent \rp\ magnitude, and \todo{0.5}\,mags in absolute
+\rp\ magnitude. If the initial query of \todo{128} points does not extend wide enough to
+meet our constraints on minimum radius, we extend the radius (linearly in each scaled)
+dimension until the minimum constraints on radii are met. Our minimum radii constraints
+imply that there will be many (e.g., $>>\todo{128}$) points returned in dense regions of the
+Hertszprung-Russell diagram. In these situations we then subsample the selected points
+to return a random \todo{1024} sources in any ball, while enforcing our constraints on
+minimum radii. As a result, the number of `similar' sources selected will range from
+a minimum of \todo{128} to a maximum of \todo{1024}, and the minimum radii constraints on various
+dimensions are always maintained.
+
+% maximum ball radius
+
+\subsection*{Step 2: Fitting the two-component mixture}
+
+We fit a two-component mixture model to the distribution of radial velocity jitter for
+all similar sources selected through Step 1. This mixture model includes a normal
+distribution to represent the intrinsic jitter for single stars, and a log-normal
+distribution to represent stellar multiples. Specifically the model parameters include
+a mixing weight parameter $\theta$ that is constrained between $(0, 1)$, the mean 
+$\mu_{s}$ and standard deviation $\sigma_{s}$ of the normal distribution,
+and the mean $\mu_{m}$ and standard deviation $\sigma_{m}$ of the 
+log-normal distribution.
+
+We assume a beta prior on $\theta$  such that $\theta \sim \mathcal{B}\left(5, 5\right)$
+and require that $\sigma_{s}$ and $\sigma_{m}$ are larger than zero.
+Depending on the distribution of the radial velocity jitter, we often found this mixture
+challenging to optimize because the log-normal distribution could easily account for 
+most of the observed distribution, which tended the mixing weight to zero and implied
+that all selected sources were stellar multiples. To avoid this situation we placed a
+constraint on the mean of the log-normal distribution to ensure that the \emph{mode}
+of the log-normal distribution is larger than $\mu_{s} + 1\sigma_{s}$ such that
+\begin{eqnarray}
+%	\mu_{s} + 5\sigma_{s} & > & \exp\left(\mu_{m} - \sigma_{m}^2\right) \label{eq:rvm_lower_bound} \\
+	\mu_{s} + 1\sigma_{s} & < & \exp\left(\mu_{m} - \sigma_{m}^2\right) \label{eq:rvm_upper_bound} \quad .
+\end{eqnarray}
+
+\begin{comment}
+However, we found that this constraint alone was insufficient. Often the mode would
+be very close to the $\mu_{s} + 1\sigma_{s}$ lower bound, which implies that
+the log-normal distribution is providing a considerable fraction of probability support
+to sources with radial velocity jitter near the mean of the \emph{normal} distribution.
+For this reason we chose to alter our lower bound constraint in that we also required
+that at the mean of the normal distribution, the cumulative probability distribution
+of the log-normal must be less than 10\% (let $q = 0.10$). The cumulative distribution 
+function for the log-normal distribution is
+\begin{eqnarray}
+\textrm{CDF}(x|\mu_{m},\sigma_{m}) = \frac{1}{2} + \frac{1}{2}\textrm{erf}\left(\frac{\log{x} - \mu_{m}}{\sqrt{2}\sigma_{m}}\right) 
+\end{eqnarray}
+
+\noindent{}where \textrm{erf} is the standard error function. We use the approximation 
+to the error function 
+\begin{eqnarray}
+	\textrm{erf}\left(x\right) \approx \tanh\left(\sqrt{\pi}\log{2}x\right)
+\end{eqnarray}
+
+\noindent{}and use the identity
+\begin{eqnarray}
+	\tanh^{-1}\left(x\right) = \frac{1}{2}\left[\log{\left(1 + x\right)} - \log{\left(1 - x\right)}\right]
+\end{eqnarray}
+
+\noindent{}to write our constraint on $\mu_{m}$ as
+\begin{eqnarray}
+	\mu_{m} & > & \log\mu_{s} + \frac{\sigma_{m}}{2\log{2}}\sqrt{\frac{2}{\pi}}\log{\left(\frac{1}{q} - 1\right)} \nonumber \\
+	\mu_{m} & > & \log\mu_{s} + \frac{\sigma_{m}}{2\log{2}}\sqrt{\frac{2}{\pi}}\log{9} \label{eq:rvm_lower_bound2} \quad .
+\end{eqnarray}
+
+Incorporating the bounds from Equations \ref{eq:rvm_lower_bound}, \ref{eq:rvm_upper_bound},
+and \ref{eq:rvm_lower_bound2} leads to the prior
+\begin{eqnarray}
+	\mu_{m} \sim \mathcal{U}\left(\alpha, \beta\right)
+\end{eqnarray}
+
+\noindent{}where:
+\begin{eqnarray}
+  \alpha & = & \max\left(\!
+	\begin{aligned}
+		\log{\left(\mu_{s} + 1\sigma_{s}\right)} + \sigma_{m}^2\\
+		\log\mu_{s} + \frac{\sigma_{m}}{2\log{2}}\sqrt{\frac{2}{\pi}}\log{9}
+	\end{aligned}\right) \\
+  \beta & = & \log{\left(\mu_{s} + 5\sigma_{s}\right)} + \sigma_{m}^2 \quad .
+\end{eqnarray}
+
+We found these priors were sufficient to ensure numerical stability and efficiency.
+\end{comment}
+
+Wherever possible we initialised the model parameters from nearby points (within the
+ball selected by the k-d tree) where an optimised result already exists. This was primarily to minimise
+computational cost; the initialisation did not have an impact on the result. 
+We required the log probability and the gradient of the log
+probability to reach machine precision before considering the optimisation
+converged. After convergence, we queried whether any points within the ball of
+similar stars had not been optimized. If no result existed, we would next move to that
+point. We refer to this process as \emph{swarm optimisation}, which we conducted in 
+parallel, and illustrate in Figure \ref{fig:swarm_optimisation}.
+
+\todo{To minimise computational cost we assigned the closest 8 stars with the result
+from others}
+
+\subsubsection*{Step 3: Evaluating probability and system properties}
+
+\todo{Bayes factors for each dimension?}
+
+\todo{source properties: semi-amplitude (easy)}
+
+\todo{Can we recover period and eccentricity using The Joker?}
+
+%For higher-order star systems with longer periods, we can still find that a
+%stellar multiple system is more likely than a single star scenario, but our
+%estimate of the orbital properties (e.g., the semi-amplitude $K$) will be biased
+%to lower values because we have not fully sampled half of the orbital period.
+%At even longer orbital periods, or for stellar multiples with mass ratios that
+%produce a small velocity semi-amplitude $K$, under our assumptions these systems 
+%may be classified as single-star systems because the intrinsic radial velocity variation is within the expected 
+%variations for single stars of a similar apparent magnitude and colour. We 
+%revisit this issue in Section \ref{sec:sb_limits}, where we explicitly define 
+%the probability that a \Gaia\ source is a SB1-type system $p(\textrm{SB1}|y)$ 
+%within defined limits of orbital period (and other orbital and source parameters). 
+%Outside of this parameter range, we are insensitive to distinguishing single-star 
+%systems from higher-order SB1-type systems.
+
+%\subsection{Unresolved near-equal mass binaries: photometry}
+%\label{sec:pb_methods}
+
+%If a binary system is observed nearly face-on (at an inclination angle 
+%$i \approx 0^\circ$) then there is will be no detectable excess radial 
+%velocity variations. In principle there may be detectable astrometric
+%variations, but most near-equal mass binaries would be detected more 
+%reliably through intrinsic magnitudes that are anomalously lower (brighter)
+%and bluer colours than what would be expected for a single star system.
 
@@
 \begin{figure*}
-	\includegraphics[width=1.0\textwidth]{../figures/sb2_rvs_completeness.pdf}
+%	\includegraphics[width=1.0\textwidth]{../figures/sb2_rvs_completeness.pdf}
+	\includegraphics[width=1.0\textwidth]{../figures/todo.png}
     \caption{Fraction of \Gaia\ sources with reported radial velocities
@@
 \begin{figure}
-	\includegraphics[width=0.4\textwidth]{../figures/sb2_histograms.pdf}
+%	\includegraphics[width=0.4\textwidth]{../figures/sb2_histograms.pdf}
+	\includegraphics[width=0.4\textwidth]{../figures/todo.png}
     \caption{Distribution of 
@@
 \begin{figure*}
-	\includegraphics[width=1.0\textwidth]{../figures/sb2_sky_structure.pdf}
+%	\includegraphics[width=1.0\textwidth]{../figures/sb2_sky_structure.pdf}
+	\includegraphics[width=1.0\textwidth]{../figures/todo.png}
+
     \caption{Fraction of sources (per sky bin) without reported radial velocities
@@
 
-\subsection{SB1 Systems: Single-lined spectroscopic binaries}
-\label{sec:sb1_methods}
+%\subsection{Photometric binaries} \label{sec:method_photometric_binaries}
 
-Radial velocity measurements are not available for individual epochs in the 
-second \Gaia\ data release \citep{Lindegren:2018,Cropper:2018}, but a point estimate
-and error in radial velocity is available for 7,224,631 sources \cite{Cropper:2018}.
-% TODO: That's the value from the paper, but I think the true value is lower. Check.
-The reported radial velocity error ($\GaiaRVE$; column name \texttt{radial\_velocity\_error}) 
-is a function of the number of transits $N$ (i.e., the number of observations; 
-represented by column \texttt{rv\_nb\_transits}) and the standard deviation among 
-those $N$ measurements of radial velocity. This allows us to calculate the standard
-deviation in radial velocity for each source, independent of the number of measurements,
-which we will refer to as the \emph{radial velocity jitter} $\RVJitter$,
-\begin{eqnarray}
-\RVJitter = \sqrt{\frac{2N}{\pi}}\GaiaRVE \quad .
-\end{eqnarray}
+%\begin{eqnarray}
+%	\textrm{Amp} = \log_{10}\left(\sqrt{N_{obs,al}}\frac{\sigma_{\langle{}f_G\rangle}}{\langle{}f_G\rangle}\right) \label{eq:photometric_variability}
+%\end{eqnarray}
 
-For a single star (without a significant mass companion) the source radial velocity
-jitter represents the minimum uncertainty with which \Gaia\ can measure 
-radial velocity for a star of that colour, apparent brightness, and absolute brightness.
-However for stars with stellar companions and no evidence of double lines or binarity in the
-cross-correlation-function, so-called SB1 type systems \citep{someone}, the
-radial velocity jitter is a function of the orbital parameters and observation
-epochs. We assume that the distribution of radial velocity jitter in the second
-\Gaia\ data release is a mixture of two populations:
-
-\begin{enumerate}
-\item \emph{single-star systems}, where the radial velocity jitter represents the
-      minimum radial velocity noise that \Gaia\ can measure for a star of that
-      \bp\ - \rp\ colour, apparent \rp\ magnitude, and absolute \rp\ magnitude;
-
-\item \emph{higher-order star systems}, where the radial velocity jitter is
-      significantly above that for single-star systems of a similar colour, 
-      apparent magnitude, and absolute magnitude.
-\end{enumerate}
-
-The mean radial velocity jitter of a single star will vary depending on the 
-\bp\ - \rp\ colour of the source and the apparent magnitude. Similarly, the mean
-radial velocity jitter of stellar \emph{multiples} will depend on the source 
-colour and its position in the Hertzsprung-Russell diagram. For these reasons,
-we must consider the radial velocity jitter of a source in context with other
-sources that have a similar \bp\ - \rp\ colour, a similar apparent \rp\ magnitude,
-and a similar absolute \rp\ magnitude. In principle if we knew the underlying
-distribution and properties of stellar multiples then we could model how different
-binary (and higher order) systems would be observable across the Hertszprung-Russell
-diagram.
-
-We chose to adopt a non-parametric model for stellar multiplicity across the
-Hertsprung-Russell diagram because we do not know the distribution of properties
-of stellar multiples. Specifically, for each source we take the following steps:
-
-\begin{enumerate}
-
-	\item Select sources with similar \bp\ - \rp\ colours, a similar apparent \rp\
-		  magnitude, and a similar absolute \rp\ magnitude.
-
-	\item We fit a two-component mixture model to the distribution of radial
-		  velocity jitter for the subset of similar sources selected in Step 1.
-
-	\item Evaluate whether the source of interest is more consistent with being
-		  drawn from the single star component (the normal distribution) or the
-		  stellar multiple component (the log-normal distribution).
-
-	\item For sources that are more consistent with being drawn from the stellar
-		  multiple component, provide a partial characterisation of the system by
-		  calculating the excess radial velocity jitter (above what is expected 
-		  for a single star with similar observed properties), among other 
-		  quantities.
-
-\end{enumerate}
-
-We illustrate this procedure in Figure \ref{fig:npm_schematic}, where we demonstrate
-how the distribution of source properties (e.g., radial velocity jitter) varies across
-the Hertsprung-Russell diagram. While this model is flexible enough to describe varying
-properties of single stars and stellar multiples, there are a number of explicit 
-assumptions that go into this model:
-
-\todo{assumptions}
-
-\begin{figure*}
-	\includegraphics[width=1.0\textwidth]{../figures/todo.png}
-    \caption{Illustration of the method we adopt for identifying and characterising
-    		 stellar multiplicity across the Hertsprung-Russell diagram. For each
-		 	 \Gaia\ source we select between 128 and 1024 sources that have
-			 similar \bp\ - \rp\ colours, apparent \rp\ magnitudes, and absolute \rp\
-			 magnitudes (Step 1). With these similar sources we fit a two-component
-			 model to distinguish single star sources from stellar multiples (Step 2).
-			 With the model parameters optimised from Step 2, we evaluate probability
-			 of stellar multiplicity and partially characterise the higher order
-			 systems by assuming binarity and estimating the system semi-amplitude (Step 3).}
-    \label{fig:npm_schematic}
-\end{figure*}
-
-\subsubsection*{Step 1: Selecting similar sources}
-
-We construct a $k$-dimensional tree \citep[k-d tree;][]{} for all sources in the
-dimensions of \bp\ - \rp\ colour, apparent \rp\ magnitude, and absolute \rp\ magnitude.
-When doing so we scale the dimensions such that 0.1\,mags in \bp\ - \rp\ colour is
-approximately equal to 1 magnitude in apparent \rp\ magnitude and 1 magnitude in 
-absolute \rp\ magnitude. We use this k-d tree for constructing the `ball' of similar
-stars for every source.
-
-We enforce a number of constraints when querying for similar stars. For each source
-we require the `ball' in multi-dimensional space to include at least 128 sources.
-We further require that the radius of the ball extends to at least 0.05\,mags in 
-\bp\ - \rp\ colour, 0.5\,mags in apparent \rp\ magnitude, and 0.5\,mags in absolute
-\rp\ magnitude. If the initial query of 128 points does not extend wide enough to
-meet our constraints on minimum radius, we extend the radius (linearly in each scaled)
-dimension until the minimum constraints on radii are met. Our minimum radii constraints
-imply that there will be many (e.g., $>>128$) points returned in dense regions of the
-Hertsprung-Russell diagram. In these situations we then subsample the selected points
-to return a random 1024 sources in any ball, while enforcing our constraints on
-minimum radii. As a result, the number of `similar' sources selected will range from
-a minimum of 128 to a maximum of 1024, and the minimum radii constraints on various
-dimensions are always maintained.
-
-\subsubsection*{Step 2: Optimizing the two-component mixture}
-
-We fit a two-component mixture model to the distribution of radial velocity jitter for
-all similar sources selected through Step 1. This mixture model includes a normal
-distribution to represent the intrinsic jitter for single stars, and a log-normal
-distribution to represent stellar multiples. Specifically the model parameters include
-a mixing weight parameter $\theta$ that is constrained between $(0, 1)$, the mean 
-$\mu_{s}$ and standard deviation $\sigma_{s}$ of the normal distribution,
-and the mean $\mu_{m}$ and standard deviation $\sigma_{m}$ of the 
-log-normal distribution. For brevity we
-
-We assume a beta prior on $\theta$  such that $\theta \sim \mathcal{B}\left(5, 5\right)$
-and we require $\sigma_{s}$ and $\sigma_{m}$ to be larger than zero.
-Depending on the distribution of the radial velocity jitter, we often found this mixture
-challenging to optimize because the log-normal distribution could easily account for 
-most of the observed distribution, which tended the mixing weight to zero and implied
-that all selected sources were stellar multiples. To avoid this situation we placed a
-constraint on the mean of the log-normal distribution to ensure that the \emph{mode}
-of the log-normal distribution is between $\mu_{s} + 1\sigma_{s}$ and
-$\mu_{s} + 5\sigma_{s}$. The mode of the log-normal distribution is
-$\exp\left(\mu_{m} - \sigma_{m}^2\right)$ such that
-\begin{eqnarray}
-	\mu_{s} + 5\sigma_{s} & > & \exp\left(\mu_{m} - \sigma_{m}^2\right) \label{eq:rvm_lower_bound} \\
-	\mu_{s} + 1\sigma_{s} & < & \exp\left(\mu_{m} - \sigma_{m}^2\right) \label{eq:rvm_upper_bound} \quad .
-\end{eqnarray}
-
-However, we found that this constraint alone was insufficient. Often the mode would
-be very close to the $\mu_{s} + 1\sigma_{s}$ lower bound, which implies that
-the log-normal distribution is providing a considerable fraction of probability support
-to sources with radial velocity jitter near the mean of the \emph{normal} distribution.
-For this reason we chose to alter our lower bound constraint in that we also required
-that at the mean of the normal distribution, the cumulative probability distribution
-of the log-normal must be less than 10\% (let $q = 0.10$). The cumulative distribution 
-function for the log-normal distribution is
-\begin{eqnarray}
-\textrm{CDF}(x|\mu_{m},\sigma_{m}) = \frac{1}{2} + \frac{1}{2}\textrm{erf}\left(\frac{\log{x} - \mu_{m}}{\sqrt{2}\sigma_{m}}\right) 
-\end{eqnarray}
-
-\noindent{}where \textrm{erf} is the standard error function. We use the approximation 
-to the error function 
-\begin{eqnarray}
-	\textrm{erf}\left(x\right) \approx \tanh\left(\sqrt{\pi}\log{2}x\right)
-\end{eqnarray}
-
-\noindent{}and use the identity
-\begin{eqnarray}
-	\tanh^{-1}\left(x\right) = \frac{1}{2}\left[\log{\left(1 + x\right)} - \log{\left(1 - x\right)}\right]
-\end{eqnarray}
-
-\noindent{}to write our constraint on $\mu_{m}$ as
-\begin{eqnarray}
-	\mu_{m} & > & \log\mu_{s} + \frac{\sigma_{m}}{2\log{2}}\sqrt{\frac{2}{\pi}}\log{\left(\frac{1}{q} - 1\right)} \nonumber \\
-	\mu_{m} & > & \log\mu_{s} + \frac{\sigma_{m}}{2\log{2}}\sqrt{\frac{2}{\pi}}\log{9} \label{eq:rvm_lower_bound2} \quad .
-\end{eqnarray}
-
-Incorporating the bounds from Equations \ref{eq:rvm_lower_bound}, \ref{eq:rvm_upper_bound},
-and \ref{eq:rvm_lower_bound2} leads to the prior
-\begin{eqnarray}
-	\mu_{m} \sim \mathcal{U}\left(\alpha, \beta\right)
-\end{eqnarray}
-
-\noindent{}where:
-\begin{eqnarray}
-  \alpha & = & \max\left(\!
-	\begin{aligned}
-		\log{\left(\mu_{s} + 1\sigma_{s}\right)} + \sigma_{m}^2\\
-		\log\mu_{s} + \frac{\sigma_{m}}{2\log{2}}\sqrt{\frac{2}{\pi}}\log{9}
-	\end{aligned}\right) \\
-  \beta & = & \log{\left(\mu_{s} + 5\sigma_{s}\right)} + \sigma_{m}^2 \quad .
-\end{eqnarray}
-
-We found these priors were sufficient to ensure numerical stability and efficiency.
-
-Wherever possible we initialised the model parameters from nearby points (within the
-ball selected by the k-d tree) where an optimised result already exists. This was primarily to minimise
-computational cost; the initialisation did not have a noticeable impact on the
-optimised result. We required the log probability and the gradient of the log
-probability to reach machine precision before considering the optimisation
-converged. After convergence, we queried to see if any points within the ball of
-similar stars had not been optimized. If no result existed, we would next move to that
-point. We refer to this process as \emph{swarm optimisation}, which we conducted in 
-parallel, and illustrate in Figure \ref{fig:swarm_optimisation}.
-
-\todo{To minimise computational cost we assigned the closest 8 stars with the result
-from others}
-
-\subsubsection*{Step 3: Evaluating probability and system properties}
-
-\todo{Bayes factors for each dimension?}
-
-\todo{source properties: semi-amplitude (easy)}
-
-\todo{Can we recover period and eccentricity using The Joker?}
-
-%For higher-order star systems with longer periods, we can still find that a
-%stellar multiple system is more likely than a single star scenario, but our
-%estimate of the orbital properties (e.g., the semi-amplitude $K$) will be biased
-%to lower values because we have not fully sampled half of the orbital period.
-%At even longer orbital periods, or for stellar multiples with mass ratios that
-%produce a small velocity semi-amplitude $K$, under our assumptions these systems 
-%may be classified as single-star systems because the intrinsic radial velocity variation is within the expected 
-%variations for single stars of a similar apparent magnitude and colour. We 
-%revisit this issue in Section \ref{sec:sb_limits}, where we explicitly define 
-%the probability that a \Gaia\ source is a SB1-type system $p(\textrm{SB1}|y)$ 
-%within defined limits of orbital period (and other orbital and source parameters). 
-%Outside of this parameter range, we are insensitive to distinguishing single-star 
-%systems from higher-order SB1-type systems.
-
-%\subsection{Unresolved near-equal mass binaries: photometry}
-%\label{sec:pb_methods}
-
-%If a binary system is observed nearly face-on (at an inclination angle 
-%$i \approx 0^\circ$) then there is will be no detectable excess radial 
-%velocity variations. In principle there may be detectable astrometric
-%variations, but most near-equal mass binaries would be detected more 
-%reliably through intrinsic magnitudes that are anomalously lower (brighter)
-%and bluer colours than what would be expected for a single star system.
-
-\subsection{Astrometric binaries} \label{sec:method_astrometric_binaries}
-
-Many higher-order star systems are unresolved point sources in \Gaia, but the
-stellar companion(s) may have sufficient mass to measurably affect the tangential 
-velocity of the primary and produce a detectable jitter in the astrometric 
-positions. There are many other sources of astrometric jitter, but here we show 
-that the principle effect of astrometric excess noise is due to stellar 
-companions. Note that due to the astrometry `degree of freedom bug' in the 
-processing of the second \Gaia\ data release \citep[e.g., see Appendix A1 of ][]{Lindegren:2018}, 
-that resulted in 80\% of sources having zero astrometric excess noise,
-we instead make use of the astrometric unit weight error $u$
-\begin{eqnarray}
-	u = \sqrt{\frac{\chi^2_{al}}{N_{obs,al} - 5}} \label{eq:astrometric_unit_weight_error}
-\end{eqnarray}
-\noindent{}where $\chi^2_{al}$ is the astrometric $\chi^2$ value
-in the along-scan direction (column \texttt{astrometric\_chi2\_al}) and 
-$N_{obs,al}$ is the number of astrometric observations in the along-scan
-direction (column \texttt{astrometric\_n\_obs\_al}). The values of this
-quantity can be interpreted similarly to how the astrometric excess noise
-would be interpreted (e.g., large values are less consistent with a
-single star model), although like the radial velocity and astrometric
-jitter, it must be considered in context with stars of similar colour,
-apparent magnitude (and preferably, absolute magnitude).
-
-\subsection{Photometric binaries} \label{sec:method_photometric_binaries}
-
-\todo{bp rp excess factor}
-
-\begin{eqnarray}
-	\textrm{Amp} = \log_{10}\left(\sqrt{N_{obs,al}}\frac{\sigma_{\langle{}f_G\rangle}}{\langle{}f_G\rangle}\right) \label{eq:photometric_variability}
-\end{eqnarray}
-
-\section{Results}
+\section{Results} \label{sec:results}
 

### 2018-08-21 scripts/download_literature_catalogs.sh
add scripts

@@ -1,0 +1,1 @@
+wget -nc -O GALAH_DR2.1_catalog.fits https://datacentral.aao.gov.au/teamdata/GALAH/public/GALAH_DR2.1_catalog.fits

### 2018-08-21 .ci/install.sh
trigger travis

@@ -1,9 +1,5 @@
 #!/bin/bash -x
 
-# SSL workaround
-#mkdir -p ~/.config/Tectonic
-#cp .ci/tectonic.config.toml ~/.config/Tectonic/config.toml
-
 # Conda install
 wget https://repo.continuum.io/miniconda/Miniconda3-latest-Linux-x86_64.sh -O miniconda.sh;
 bash miniconda.sh -b -p $HOME/miniconda

### 2018-08-21 .ci/upload.sh

@@ -1,11 +1,9 @@
 #!/bin/bash -x
 
 # Push to GitHub
-if [ -n "$GITHUB_API_KEY" ]; then
-  cd $TRAVIS_BUILD_DIR
-  git checkout --orphan pdf
-  git rm -rf .
-  git add -f ms.pdf
-  git -c user.name='travis' -c user.email='travis' commit -m "building the paper"
-  git push -q -f https://andycasey:$GITHUB_API_KEY@github.com/andycasey/velociraptor pdf
-fi
+cd $TRAVIS_BUILD_DIR
+git checkout --orphan pdf
+git rm -rf .
+git add -f article/ms.pdf
+git -c user.name='travis' -c user.email='travis' commit -m "building the paper"
+git push -q -f https://andycasey:$GITHUB_API_KEY@github.com/andycasey/velociraptor pdf

### 2018-08-22 create_release_candidate.py
things

@@ -1,0 +1,464 @@
+
+"""
+Fit a non-parametric model to the jitter in radial velocity error and the
+astrometric unit weight error, and then use a gaussian process to enforce
+smoothness and make predictions for 10^7 sources.
+"""
+
+import george
+import logging as logger
+import multiprocessing as mp
+import numpy as np
+import os
+import pickle
+import tqdm
+import yaml
+from astropy.io import fits
+from george import kernels
+from scipy import optimize as op
+from time import (sleep, time) # yea it is
+
+
+import npm_utils as npm
+import stan_utils as stan
+
+
+with open("model.yaml", "r") as fp:
+    config = yaml.load(fp)
+
+results_path = config["results_path"]
+if os.path.exists(results_path) and not config.get("overwrite_results", False):
+    raise IOError("{} exists and we will not overwrite".format(results_path))
+
+np.random.seed(config.get("random_seed", None))
+
+# Load the data.
+data = fits.open(config["data_path"])[config.get("data_hdu", 1)].data
+
+catalog_results = dict()
+
+# Select the sources that will be used by the non-parametric model and the
+# gaussian process. These sources should have:
+# - finite values in the predictor label names of all non-parametric models
+# - finite values in the kdtree label names of all non-parametric models
+
+relevant_label_names = list(set(sum([
+    (m["predictor_label_names"], m["kdtree_label_names"]) \
+    for descr, m in config["non_parametric_models"].items()])))
+
+is_suitable_source = np.all(np.isfinite(np.array(
+    [data[ln] for ln in relevant_label_names])), axis=0)
+
+logger.info("There are {} suitable sources".format(sum(is_suitable_source)))
+
+# We will use all indices relative to data_subset, because if a source does not
+# have finite properties of some label then we cannot use it for GP predictions
+# anyways.
+data_subset = data[is_suitable_source]
+
+# Select the sources that will be used for fitting.
+S = len(data_subset)
+N = min(config["number_of_sources_to_fit"], S)
+
+finite_indices = np.arange(S)
+data_indices = np.random.choice(finite_indices, size=N, replace=False)
+
+
+def construct_kernel(D):
+    return kernels.ExpKernel(np.ones(D), ndim=D) \
+         + kernels.Matern32Kernel(np.ones(D), ndim=D)
+
+
+
+def get_reference_indices(indices):
+    indices = np.atleast_1d(indices)
[…]

### 2018-08-22 npm_run_astrometry.py

@@ -1,0 +1,520 @@
+
+""" Set up and run the non-parametric model. """
+
+import numpy as np
+import os
+import multiprocessing as mp
+import pickle
+import yaml
+import tqdm
+import logging
+from time import sleep
+from astropy.io import fits
+
+import npm_utils as npm
+import stan_utils as stan
+
+DO_BALL_TEST = False
+
+np.random.seed(42)
+
+with open("npm-config.astrometry.yaml", "r") as fp:
+    config = yaml.load(fp)
+
+# Load in the data.
+data = fits.open(config["data_path"])[1].data
+
+all_label_names = list(config["kdtree_label_names"]) \
+                + list(config["predictor_label_names"])
+
+# Set up a KD-tree.
+X = np.vstack([data[ln] for ln in config["kdtree_label_names"]]).T
+finite = np.all([np.isfinite(data[ln]) for ln in all_label_names], axis=0)
+finite_indices = np.where(finite)[0]
+
+N, D = X.shape
+F = finite_indices.size
+L = 4 * len(config["predictor_label_names"]) + 1 # + 1 if using mu_multiple_uv
+C = config["share_optimised_result_with_nearest"]
+
+kdt, scales, offsets = npm.build_kdtree(
+    X[finite], relative_scales=config["kdtree_relative_scales"])
+
+kdt_kwds = dict(offsets=offsets, scales=scales, full_output=True)
+kdt_kwds.update(
+    minimum_radius=config["kdtree_minimum_radius"],
+    maximum_radius=config.get("kdtree_maximum_radius", None),
+    minimum_points=config["kdtree_minimum_points"],
+    maximum_points=config["kdtree_maximum_points"],
+    minimum_density=config.get("kdtree_minimum_density", None))
+
+model = stan.load_stan_model(config["model_path"], verbose=False)
+
+default_opt_kwds = config.get("optimisation_kwds", {})
+
+# Make sure that some entries have the right units.
+for key in ("tol_obj", "tol_grad", "tol_rel_grad", "tol_rel_obj"):
+    if key in default_opt_kwds:
+        default_opt_kwds[key] = float(default_opt_kwds[key])
+
+logging.info("k-d tree keywords: {}".format(kdt_kwds))
+logging.info("optimization keywords: {}".format(default_opt_kwds))
+
+
+done = np.zeros(N, dtype=bool)
+queued = np.zeros(N, dtype=bool)
[…]

### 2018-08-22 npm_utils.py

@@ -214,7 +214,7 @@ def query_around_point(kdtree, point, offsets=0, scales=1, minimum_radius=None,
 
 
     elif maximum_radius is not None:
-        
+
         L = np.where(np.all(np.abs(point[0] - np.asarray(kdtree.data)[indices]) <= maximum_radius, axis=1))[0]
         maximum_points = min(maximum_points, L.size)
 
@@ -384,12 +384,24 @@ def _pack_params(theta, mu_single, sigma_single, mu_multiple, sigma_multiple, mu
         return np.hstack([theta, mu_single, sigma_single, mu_multiple, sigma_multiple, mu_multiple_uv])
 
 
-def _check_params_dict(d):
+def _check_params_dict(d, bounds_dict=None, fail_on_bounds=True, tolerance=0.01):
     if d is None: return d
     
     dc = {**d}
     for k in ("mu_single", "sigma_single", "mu_multiple", "sigma_multiple"):
         dc[k] = np.atleast_1d(dc[k]).flatten()[0]
+        if bounds_dict is not None and k in bounds_dict:
+            lower, upper = bounds_dict[k]
+            if (not np.all(upper >= dc[k]) or not np.all(dc[k] >= lower)):
+                if fail_on_bounds:
+                    raise ValueError("bounds not met: {} = {} not within ({} {})"\
+                                     .format(k, dc[k], lower, upper))
+                else:
+                    print("Clipping initial {} to be within bounds ({}, {}): {}"\
+                        .format(k, lower, upper, dc[k]))
+                    dc[k] = np.clip(dc[k], lower + tolerance, upper - tolerance)
+
+
     return dc
 
 

### 2018-08-22 plot_gp_models_astrometry.py

@@ -1,6 +1,5 @@
 import matplotlib
 import matplotlib.pyplot as plt
-plt.switch_backend("agg")
 import numpy as np
 import os
 import pickle
@@ -17,24 +16,21 @@ import george.kernels
 from mpl_utils import mpl_style
 
 
-seed = 123
+seed = 42
 np.random.seed(seed)
 matplotlib.style.use(mpl_style)
 
 gp_labels = (
-    r"$\mu_\textrm{s}$ \textrm{/ km\,s}$^{-1}$",
-    r"$\sigma_\textrm{s}$ \textrm{/ km\,s}$^{-1}$",
-    r"$\mu_\textrm{m}$ \textrm{/ km\,s}$^{-1}$",
-    r"$\sigma_\textrm{m}$ \textrm{/ km\,s}$^{-1}$",
+    r"$\mu_\textrm{s}$",
+    r"$\sigma_\textrm{s}$",
+    r"$\mu_\textrm{m}$",
+    r"$\sigma_\textrm{m}$",
 )
 G = len(gp_labels)
 
 
 
-def x_for_gp(x):
-    x_ = np.copy(x)
-    x_[:, 0] = 1.0/x_[:,0]
-    return x_
+x_for_gp = lambda x: x
 
 
 def construct_kernel(D):
@@ -43,10 +39,10 @@ def construct_kernel(D):
 
 
 # Load the data.
-with open("npm-config.rv.yaml", "r") as fp:
+with open("npm-config.astrometry.yaml", "r") as fp:
     config = yaml.load(fp)
 
-with open("results/rv_single_epoch_scatter.v3.pkl", "rb") as fp:
+with open("results/astrometric_unit_weight_error.v4.pkl", "rb") as fp:
     npm_results = pickle.load(fp)
 
 
@@ -55,15 +51,16 @@ data = fits.open(config["data_path"])[1].data
 X = np.vstack([data[ln] for ln in config["kdtree_label_names"]]).T
 
 # Number of points to fit the GPs.
-M = 1000
+M = 10000
 
 
 N, D = X.shape
-finite_indices = np.arange(N)[np.all(np.isfinite(npm_results), axis=1)]
+use_for_fit = np.all(np.isfinite(npm_results), axis=1) 
+finite_indices = np.arange(N)[use_for_fit]
 
 fit_indices = np.random.choice(finite_indices, M, replace=False)
 
-
+"""
 gps = []
 x = X[fit_indices]
 
@@ -105,35 +102,39 @@ for npm_index in range(1, 5):
[…]

### 2018-08-22 create_release_candidate.py
fucking

@@ -14,10 +14,12 @@ import pickle
 import tqdm
 import yaml
 from astropy.io import fits
+from astropy.table import Table
+from collections import OrderedDict
 from george import kernels
 from scipy import optimize as op
+from scipy.special import logsumexp
 from time import (sleep, time) # yea it is
-
 
 import npm_utils as npm
 import stan_utils as stan
@@ -35,7 +37,7 @@ np.random.seed(config.get("random_seed", None))
 # Load the data.
 data = fits.open(config["data_path"])[config.get("data_hdu", 1)].data
 
-catalog_results = dict()
+catalog_results = OrderedDict()
 
 # Select the sources that will be used by the non-parametric model and the
 # gaussian process. These sources should have:
@@ -67,7 +69,6 @@ data_indices = np.random.choice(finite_indices, size=N, replace=False)
 def construct_kernel(D):
     return kernels.ExpKernel(np.ones(D), ndim=D) \
          + kernels.Matern32Kernel(np.ones(D), ndim=D)
-
 
 
 def get_reference_indices(indices):
@@ -363,7 +364,19 @@ for description, npm_config in config["non_parametric_models"].items():
         # Serial.
         serial_process(*data_indices)
 
-    npm_results[description] = results
+    prefix = npm_config["descriptive_prefix"]
+    gp_labels = (
+    	"npm_{prefix}_theta",
+        "npm_{prefix}_mu_single",
+        "npm_{prefix}_sigma_single",
+        "npm_{prefix}_mu_multiple",
+        "npm_{prefix}_sigma_multiple",
+    )
+
+    for i, gp_label_format in enumerate(gp_labels):
+    	_ = np.nan * np.ones(len(data))
+    	_[data_indices] = results[:, i]
+    	catalog_results[gp_label_format.format(prefix=prefix)] = _
 
     # Run gaussian process on results.
     use_in_gp = np.all(np.isfinite(results), axis=1)
@@ -378,10 +391,10 @@ for description, npm_config in config["non_parametric_models"].items():
     gp_results[description] = []
 
     npm_labels = (
-        "{descriptive_prefix}_mu_single",
-        "{descriptive_prefix}_sigma_single",
-        "{descriptive_prefix}_mu_multiple",
-        "{descriptive_prefix}_sigma_multiple",
+        "{prefix}_mu_single",
[…]

### 2018-08-22 create_release_candidate.py
now it's just down to what you can afford

@@ -437,7 +437,7 @@ for description, npm_config in config["non_parametric_models"].items():
         gp_result = op.minimize(nll, p0, jac=grad_nll, method="L-BFGS-B")
         t_opt = time() - t_init
 
-        logger.info("Result: {}".format(result))
+        logger.info("Result: {}".format(gp_result))
         logger.info("Final \log{{L}} = {:.2f}".format(gp.log_likelihood(y)))
         logger.info("Took {:.0f} seconds to optimize".format(t_opt))
 
@@ -492,14 +492,22 @@ for description, npm_config in config["non_parametric_models"].items():
 
     catalog_results[f"{prefix}_tau_single"] = np.exp(log_tau_single)
 
+# TODO: Calculate joint likelihoods and tau.
+metadata_path = config["metadata_path"]
+logger.info("Saving metadata to {}".format(metadata_path))
 
-with open(config["metadata_path"], "wb") as fp:
+with open(metadata_path, "wb") as fp:
     pickle.dump((config, gp_results), fp)
 
 # Save the catalog.
+results_path = config["results_path"]
+logger.info("Saving catalog to {}".format(results_path))
+
 del data
-data = Table.read(config["data_path"])
+data = Table.read(results_path)
 for k, v in catalog_results.items():
     data[k] = v
 
-data.write(config["results_path"], overwrite=True)
+data.write(results_path, overwrite=True)
+
+

### 2018-08-23 article/ms.tex
trigger travis
minted etc

@@
 \usepackage{comment}
+\usepackage{minted}
 
@@
 \newcommand{\acronym}[1]{{\small{#1}}}
+\newcommand{\ESA}{\acronym{ESA}}
 \newcommand{\Gaia}{\project{Gaia}}
@@
 \newcommand{\NumberOfStellarSingles}{XXX}
+\newcommand{\NumberOfSourcesInSubset}{8,911,018} % check
 
@@
 \accepted{2018 XX XX}
+
+%%% This file is generated by the Makefile.
+
+\newcommand{\vcpath}{vc.tex}
+
+\IfFileExists{\vcpath}{\input{\vcpath}}{
+	\newcommand{\giturl}{UNKNOWN}
+	\newcommand{\gitslug}{UNKNOWN}
+	\newcommand{\githash}{UNKNOWN}
+	\newcommand{\gitdate}{UNKNOWN}
+	\newcommand{\gitauthor}{UNKNOWN}
+}
 
@@
 narrow absorption lines than main-sequence stars of the same temperature and colour.
-For stellar multiples, the source radial velocity jitter will be the quadrature sum of the expected jitter for the primary (if it were a single star) and a contribution based on the orbital parameters  and the observation epochs.
+For stellar multiples, the source radial velocity jitter will be the quadrature sum of the expected jitter for the primary (if it were a single star) and a contribution from the radial velocity semi-amplitude of the orbit.
 
-% todo: lack of RV measurements?
+Some sources in \Gaia\ do not have reported radial velocities even though they are
+apparently bright enough and within a suitable \bp\ - \rp\ colour range.  The likely
+explanation for these situations is that the \Gaia\  team identified
+the source to be a double-lined spectroscopic binary, and there is no utility in
+reporting a point estimate of the \emph{source} radial velocity and associated error.
+Other explanations are also possible, which implies that we cannot use the \emph{lack}
+of radial velocity as a robust identifier of stellar multiplicity. \todo{In the following
+Sections}
+% Show it ehre?
+we demonstrate that for sources that do not have reported radial velocities -- and 
+sources of similar apparent magnitude and colour \emph{do} have radial velocities --
+their astrometric jitter is usually sufficient to distinguish single stars from stellar
+multiples.
+
+% todo:  Figure showing distribution of astrometric unit weight error.
 
@@
 velocity of the primary and produce a detectable jitter in the astrometric 
-position. There are many other sources of astrometric jitter, but here we show 
-that the principle effect of astrometric excess noise is due to stellar 
-companions. Note that due to the astrometry `degree of freedom bug' in the 
-processing of the second \Gaia\ data release \citep[e.g., see Appendix A1 of ][]{Lindegren:2018}, 
-that resulted in 80\% of sources having zero astrometric excess noise,
-we instead make use of the astrometric unit weight error $u$
+position. 
+%There are many other sources of astrometric jitter, but here we show 
+%that the principle effect of astrometric excess noise is due to stellar 
+%companions.
+We make use of the astrometric unit weight error $u$ to measure astrometric
+jitter
 \begin{eqnarray}
@@
 $N_{obs,al}$ is the number of astrometric observations in the along-scan
-direction (column \texttt{astrometric\_n\_obs\_al}). The values of this
-quantity can be interpreted similarly to how the astrometric excess noise
-would be interpreted (e.g., large values are less consistent with a
-single star model), although like the radial velocity and astrometric
-jitter, it must be considered in context with stars of similar colour,
-apparent magnitude, and preferably, absolute magnitude.
+direction (column \texttt{astrometric\_n\_obs\_al}). 
+The astrometric unit weight error is unaffected by the `degree of freedom bug'
+that occurred during the processing of the the second \Gaia\ data release
+\citep[e.g., see Appendix A1 of ][]{Lindegren:2018}, which resulted in 80\% of
+sources having zero astrometric excess noise. 
+The astrometric unit weight error can be interpreted similarly to how the
+astrometric excess noise would be interpreted (e.g., large values are less
+consistent with a single star model), but using the astrometric unit weight
+error allows us to distinguish astrometric binaries in a much larger sample.
+And like the radial velocity jitter, the unit weight astrometric error of a
+single source must be considered in context with stars of similar colour 
+and apparent magnitude (but not necessarily absolute magnitude).
 
@@
 % what will our subset be?
+\subsection{\texttt{ADQL} query}
+
+We restrict our work to the brightest $\sim{}10^7$ sources in \Gaia\ DR2.
+About 70\% of sources in this subset have reported radial velocities,
+allowing us to identify spectroscopic binaries, and this subset provides
+a good sample of sources to test the precision to which we can identify
+astrometric binaries. Specifically, we executed the following Astronomical Data Query
+Language \citep[\texttt{ADQL};][]{someone} query on \ESA\ \Gaia\ archive
+\citep{esa_gaia_archive}:
+
+\begin{minted}[style=friendly]{postgresql}
+SELECT *
+  FROM gaiadr2.gaia_source
+ WHERE radial_velocity IS NOT NULL
+    OR phot_rp_mean_mag <= 13
+\end{minted}
+\noindent{}This query returned \NumberOfSourcesInSubset\ sources.
 
@@
 
-We assume that for sources with similar \bp\ - \rp\ colours, apparent \rp\ magnitudes, and absolute \rp\ magnitudes, the jitter distribution (in radial velocity or astrometry) is a mixture of two populations:
+We assume that for a sample of sources with similar \bp\ - \rp\ colours
+and apparent \rp\ magnitudes -- and absolute \rp\ magnitudes, if radial velocities
+are being considered -- the distribution of jitter is a mixture of two components:
 
@@
 \item \emph{single-star systems}, where the jitter represents the minimum noise that 
-	   \Gaia\ can measure for a star of that
-      \bp\ - \rp\ colour, apparent \rp\ magnitude, and absolute \rp\ magnitude;
+	   \Gaia\ can measure for a source with those properties (e.g., a source with
+	   that \bp\ - \rp\ colour, apparent \rp\ magnitude, and perhaps absolute \rp\
+	   magnitude);
 
-\item \emph{higher-order star systems}, where the jitter is
-      significantly above that for single-star systems of a similar colour, 
-      apparent magnitude, and absolute magnitude.
+\item \emph{higher-order star systems}, where the jitter is significantly higher 
+	  than what is measured for single-star systems of similar properties.
 \end{enumerate}
 
-We also assume that the mean single-star jitter will change
-depending on the \bp\ - \rp\ colour, the apparent magnitude, and the absolute magnitude.
-For example, the mean radial velocity jitter of single stars with $\teff \approx 6000\,\textrm{K}$
+We further assume that the jitter for a single star will change depending on the
+\bp\ - \rp\ colour, the apparent magnitude, and the absolute magnitude. For example, 
+the mean radial velocity jitter of single stars with $\teff \approx 6000\,\textrm{K}$
 will be higher than the jitter of single stars with $\teff \approx 5000\,\textrm{K}$, 
-with all else being (nearly) equal. We similarly assume that the mean 
-jitter of stellar \emph{multiples} will depend on the source colour and its position
-in the Hertzsprung-Russell diagram. For these reasons, in order to evaluate whether a
-source is more likely to be a single star or a stellar multiple, we must consider the 
-jitter in context with other sources that have similar: \bp\ - \rp\
-colour, apparent \rp\ magnitude, and absolute \rp\ magnitude. In principle if we knew
-the distribution stellar multiple properties (or assumed one) then we could model the
-observability of binary (and higher order) systems across the Hertszprung-Russell
-diagram, but this is beyond the scope of this work.
+with all else being (nearly) equal. Similarly, the minimum radial velocity jitter
+for giant stars is expected to be lower on average than for main-sequence stars of
+the same colour, as giant stars have narrower absorption lines than main-sequence
+stars of the same colour, allowing for a more precise measurement of the radial
+velocity. However, this effect will not affect the astrometric jitter: the minimum
+astrometric jitter for a main-sequence star and a giant star of the same temperature
+is likely to be indistinguishable. For these reasons, in order to evaluate whether 
+a source is more likely to be a single star or a stellar multiple, we must consider 
+the jitter in context with other sources that have a similar \bp\ - \rp\ colour, 
+apparent \rp\ magnitude, and in the case of radial velocity jitter, the absolute
+\rp\ magnitude.\footnote{The magnitude in any band would be sufficient. Throughout
+this work we chose the \rp\ band because its wavelength range most closely overlaps 
+with that of the Radial Velocity Spectrograph on board \Gaia.} However, we cannot reliably describe a quantitative
+relationship for \emph{how} exactly the astrometric and radial velocity jitter is
+expected to vary for a single star given some source properties. For these reasons
+we chose to adopt a flexible non-parametric model for stellar multiplicity, primarily
+conditioned on the assumption that if we select enough sources with \emph{similar}
+properties then within that subset of similar stars there exists \emph{some} single stars 
+and \emph{some} stellar multiples.
 
-% ADDITIONAL ASSUMPTIONS.
-\todo{additional assumptions}
+We illustrate our approach in Figure \ref{fig:npm_schematic}, which can be qualitatively
+described by the following steps. When considering the radial velocity jitter for a
+\Gaia\ source of interest, we select $N$ sources that have a similar \bp\ - \rp\ colour,
+apparent \rp\ magnitude, and absolute \rp\ magnitude. We then fit the distribution of
+radial velocity jitter $\RVJitter$ as a mixture of two components: a normal distribution
+to represent single stars, and a log-normal distribution to represent stellar multiples.
+We perform these steps for $\approx10^4$ sources. While this process approximately maps
+how the radial velocity jitter of single (and multiple) stars varies across the Hertszprung-Russell
+diagram, each source of interest is considered independent from all others. For example,
+for three sources that are all indistinguishable from each other in \bp\ - \rp\ colour, 
+apparent \rp\ magnitude, and absolute \rp\ magnitude, the procedure we described above
+treats the fitting of these three mixture models as entirely independent of each other,
+even though we suspect that the parameters of the three mixture models ought to be 
+identical. In other words: the procedure as described does not enforce any smoothness.
+Given enough data this is not a problem for densely-populated regions of the 
+Hertszprung-Russell diagram as the optimized results \emph{will} be (statistically)
+indistinguishable for similar sources. In practice this poses a problem for poorly-sampled
+regions of the Hertzsprung-Russell diagram, and raises subjective questions on what
+constitutes \emph{the most} similar type of source.
 
-Given these assumptions, we chose to adopt a non-parametric model for stellar multiplicity,
-primarily because we do not know the distribution of properties of stellar multiples.
-Our approach is illustrated in Figure \ref{fig:npm_schematic} and can be qualitatively 
-described by the following steps.\\
-
-\noindent{}For each \Gaia\ source:
-\begin{enumerate}
-
-	\item Select $N$ sources that have a similar \bp\ - \rp\ colour, apparent \rp\
-		  magnitude, and absolute \rp\ magnitude.
-
-	\item Fit the distribution of radial velocity jitter for sources selected in
-		  Step 1 using a mixture of a normal and a log-normal distribution.
-
-	\item Evaluate whether the source of interest is more consistent with being
-		  a single star component (drawn from the normal distribution) or a
-		  stellar multiple (drawn from the log-normal distribution).
-
-	\item If the source is more likely to be a stellar multiple, use the estimated
-		  properties of single stars with similar properties to estimate the
-		  excess radial velocity jitter due to the stellar companion(s).
-
-\end{enumerate}
+\todo{In the third step we apply a Gaussian process.}
 
@@
 This project was developed in a \texttt{git} repository that is publicly
-accessible at \url{https://github.com/andycasey/velociraptor}. The repository
-includes notebooks that detail our work, \LaTeX\ to compile this manuscript, 
-and scripts to reproduce our analysis in its entirety. Reproducing this work in 
-full will require at least \todo{X}\,Gb of disk space and \todo{Y}\,compute hours.
-To reproduce this work (including data retrieval, analysis, production of figures,
-and manuscript compilation) use the \todo{following commands on a modern terminal}:
+accessible at \giturl. The repository includes notebooks that detail our work,  
+\LaTeX\ to compile this manuscript, and scripts to reproduce the results presented in
+this manuscript. The results presented here can be reproduced in full (including data
+retrieval, analysis, creation of figures, and manuscript compilation) using these
+commands on a modern terminal:
+% this is not python, but latex fails to evaluate \githash if I set bash environment.
+\begin{minted}[
+style=friendly,
+escapeinside=||
+]{python} 
+git clone https://github.com/andycasey/velociraptor.git velociraptor
+cd velociraptor
+git checkout |\githash|
+./reproduce
+\end{minted}
 
-\section{Relevant Astronomical Data Query Language (ADQL) queries}
-\todo{Include relevant ADQL queries}
+Reproducing these results will require at least \todo{X}\,Gb of free disk space 
+and \todo{Y}\,hours of compute time. The settings in \texttt{model.yaml} can be 
+adjusted to reduce the sample size of the data and shorten the compute time, but
+varying these settings does not guarantee exact replication of the results shown
+here.
 
@@
 
+% astroquery
+% minted
 \software{
-	\package{Astropy} \citep{astropy},
+	\package{Astropy} \citep{astropy:v1,astropy:v2},
     \package{IPython} \citep{ipython},

### 2018-08-23 create_release_candidate.py
bug fixes

@@ -28,6 +28,10 @@ import stan_utils as stan
 with open("model.yaml", "r") as fp:
     config = yaml.load(fp)
 
+# This is real meta.
+with open(__file__, "r") as fp:
+    this_code = fp.read()
+
 results_path = config["results_path"]
 if os.path.exists(results_path) and not config.get("overwrite_results", False):
     raise IOError("{} exists and we will not overwrite".format(results_path))
@@ -44,9 +48,15 @@ catalog_results = OrderedDict()
 # - finite values in the predictor label names of all non-parametric models
 # - finite values in the kdtree label names of all non-parametric models
 
-relevant_label_names = list(set(sum([
-    (m["predictor_label_names"], m["kdtree_label_names"]) \
-    for descr, m in config["non_parametric_models"].items()])))
+#relevant_label_names = list(set(sum([
+#    (m["predictor_label_names"], m["kdtree_label_names"]) \
+#    for descr, m in config["non_parametric_models"].items()])))
+relevant_label_names = []
+for descr, m in config["non_parametric_models"].items():
+    for k in ("predictor_label_names", "kdtree_label_names"):
+        relevant_label_names.extend(m[k])
+
+relevant_label_names = list(set(relevant_label_names))
 
 is_suitable_source = np.all(np.isfinite(np.array(
     [data[ln] for ln in relevant_label_names])), axis=0)
@@ -95,7 +105,7 @@ for description, npm_config in config["non_parametric_models"].items():
         npm.build_kdtree(X, relative_scales=npm_config["kdtree_relative_scales"])
 
     kdt_kwds = dict(
-        offsets=kdt_offsets, 
+        offsets=kdt_offsets,
         scales=kdt_scales,
         full_output=True,
         minimum_radius=npm_config.get("kdtree_minimum_radius", None),
@@ -109,7 +119,7 @@ for description, npm_config in config["non_parametric_models"].items():
     for opt_key in ("tol_obj", "tol_grad", "tol_rel_grad", "tol_rel_obj"):
         if opt_key in opt_kwds:
             opt_kwds[opt_key] = float(opt_kwds[opt_key])
-    
+
     # Initialisation point.
     init_dict = npm._check_params_dict(npm_config.get("initialisation", None))
 
@@ -117,7 +127,7 @@ for description, npm_config in config["non_parametric_models"].items():
 
     # Arrays for results.
[…]

### 2018-08-29 article/ms.tex
some words

@@
 
-There are many ways that binaries can be reliably identified using \Gaia\
-data, including radial velocity measurements, astrometry, and photometry.
-These data are sensitive to characterising stellar multiplicity in 
-different ways. Here we describe the data that we use from the second
-\Gaia\ data release \citep{Lindegren:2018} before defining our model.
+Stellar multiplicity can be inferred from \Gaia\ data using radial velocity
+measurements, astrometry, and photometry. These data are sensitive in different 
+ways. Here we describe the data that we use from the second \Gaia\ data 
+release \citep{Lindegren:2018} before we define our methods.
 
@@
 have on average lower signal-to-noise (S/N) ratios, both of which result in noisier radial
-velocity measurements. The absolute magnitude is similarly important, as giant stars have
-narrow absorption lines than main-sequence stars of the same temperature and colour.
-For stellar multiples, the source radial velocity jitter will be the quadrature sum of the expected jitter for the primary (if it were a single star) and a contribution from the radial velocity semi-amplitude of the orbit.
+velocity measurements. The absolute magnitude is similarly important, as giant stars have narrow absorption lines than main-sequence stars of the same temperature and colour. 
+For sources that are stellar multiples, the radial velocity jitter will be the quadrature sum
+of the expected jitter for the primary (if it were a single star), and a contribution from the
+radial velocity semi-amplitude of the orbit.
 
-Some sources in \Gaia\ do not have reported radial velocities even though they are
-apparently bright enough and within a suitable \bp\ - \rp\ colour range.  The likely
-explanation for these situations is that the \Gaia\  team identified
-the source to be a double-lined spectroscopic binary, and there is no utility in
-reporting a point estimate of the \emph{source} radial velocity and associated error.
-Other explanations are also possible, which implies that we cannot use the \emph{lack}
-of radial velocity as a robust identifier of stellar multiplicity. \todo{In the following
-Sections}
-% Show it ehre?
-we demonstrate that for sources that do not have reported radial velocities -- and 
-sources of similar apparent magnitude and colour \emph{do} have radial velocities --
-their astrometric jitter is usually sufficient to distinguish single stars from stellar
-multiples.
+Some sources in \Gaia\ DR2 do not have reported radial velocities even though they are
+apparently bright enough and within a suitable \bp\ - \rp\ colour range. The two most
+likely explanations are that either the \Gaia\ team identified the source to be a
+double-lined spectroscopic binary, or the radial velocity error was larger than 20\,km\,s$^{-1}$,
+and so the source radial velocity was removed. While it may be appealing to use the \emph{lack}
+of radial velocity measurement as an indicator of stellar multiplicity, we show in Appendix A
+that such an approach is problematic due to completeness, and other complications arising from
+the initial \Gaia\ source catalog. However, in the following sections we show that for 
+sources that do not have reported radial velocities -- and it has an apparent magnitude 
+and colour that suggests it ought to have a radial velocity reported -- the astrometric 
+jitter is often sufficient to identify stellar multiples.
 
@@
 %companions.
-We make use of the astrometric unit weight error $u$ to measure astrometric
-jitter
+We make use of the astrometric unit weight error $u$ 
 \begin{eqnarray}
@@
 single source must be considered in context with stars of similar colour 
-and apparent magnitude (but not necessarily absolute magnitude).
+and apparent magnitude (but not necessarily absolute magnitude). For lack of
+a more descriptive term, throughout this work we will refer to the astrometric
+unit weight error as the \emph{astrometric jitter}, similar to how we refer
+to the standard deviation of radial velocities as the \emph{radial velocity jitter}.
 
@@
 We restrict our work to the brightest $\sim{}10^7$ sources in \Gaia\ DR2.
-About 70\% of sources in this subset have reported radial velocities,
-allowing us to identify spectroscopic binaries, and this subset provides
-a good sample of sources to test the precision to which we can identify
-astrometric binaries. Specifically, we executed the following Astronomical Data Query
-Language \citep[\texttt{ADQL};][]{someone} query on \ESA\ \Gaia\ archive
-\citep{esa_gaia_archive}:
+About 70\% of sources in this subset have reported radial velocities, and
+this subset is relatively close to the Sun. Specifically, we executed the
+following Astronomical Data Query Language \citep[\texttt{ADQL};][]{someone} 
+query on \ESA\ \Gaia\ archive \citep{esa_gaia_archive}:
 
@@
 
-We assume that for a sample of sources with similar \bp\ - \rp\ colours
-and apparent \rp\ magnitudes -- and absolute \rp\ magnitudes, if radial velocities
-are being considered -- the distribution of jitter is a mixture of two components:
+We assume that for a sample of sources with similar properties:
+\begin{itemize}
+	\item \bp\ - \rp\ colours;
+	\item apparent \rp\ magnitude; and
+	\item absolute \rp\ magnitude,
+\end{itemize}
+\noindent{}the distribution of jitter (in astrometry or radial velocity)
+is a mixture of two components:
 
@@
 \item \emph{single-star systems}, where the jitter represents the minimum noise that 
-	   \Gaia\ can measure for a source with those properties (e.g., a source with
-	   that \bp\ - \rp\ colour, apparent \rp\ magnitude, and perhaps absolute \rp\
-	   magnitude);
+	   \Gaia\ can measure for a source with those properties,
 
@@
 We further assume that the jitter for a single star will change depending on the
-\bp\ - \rp\ colour, the apparent magnitude, and the absolute magnitude. For example, 
-the mean radial velocity jitter of single stars with $\teff \approx 6000\,\textrm{K}$
-will be higher than the jitter of single stars with $\teff \approx 5000\,\textrm{K}$, 
-with all else being (nearly) equal. Similarly, the minimum radial velocity jitter
-for giant stars is expected to be lower on average than for main-sequence stars of
-the same colour, as giant stars have narrower absorption lines than main-sequence
-stars of the same colour, allowing for a more precise measurement of the radial
-velocity. However, this effect will not affect the astrometric jitter: the minimum
-astrometric jitter for a main-sequence star and a giant star of the same temperature
-is likely to be indistinguishable. For these reasons, in order to evaluate whether 
+source properties. For example, the mean radial velocity jitter of single stars 
+with $\teff \approx 6000\,\textrm{K}$ will be higher than the jitter of single stars with
+$\teff \approx 5000\,\textrm{K}$, all else being equal. For these reasons, in order to evaluate whether 
 a source is more likely to be a single star or a stellar multiple, we must consider 
-the jitter in context with other sources that have a similar \bp\ - \rp\ colour, 
-apparent \rp\ magnitude, and in the case of radial velocity jitter, the absolute
-\rp\ magnitude.\footnote{The magnitude in any band would be sufficient. Throughout
+the jitter in context with other sources that have a similar properties (\bp\ - \rp\ colour,
+apparent \rp\ magnitude, and absolute \rp\ magnitude).\footnote{The magnitude in any band would be sufficient. Throughout
 this work we chose the \rp\ band because its wavelength range most closely overlaps 
 with that of the Radial Velocity Spectrograph on board \Gaia.} However, we cannot reliably describe a quantitative
-relationship for \emph{how} exactly the astrometric and radial velocity jitter is
-expected to vary for a single star given some source properties. For these reasons
+relationship for \emph{how} exactly the jitter of a single star is
+expected to vary given some source properties. For these reasons
 we chose to adopt a flexible non-parametric model for stellar multiplicity, primarily
@@
 properties then within that subset of similar stars there exists \emph{some} single stars 
-and \emph{some} stellar multiples.
+and \emph{some} stellar multiples -- even if we do not necessarily know \emph{a priori}
+which sources are single stars, and which are stellar multiples.
 
-We illustrate our approach in Figure \ref{fig:npm_schematic}, which can be qualitatively
-described by the following steps. When considering the radial velocity jitter for a
-\Gaia\ source of interest, we select $N$ sources that have a similar \bp\ - \rp\ colour,
-apparent \rp\ magnitude, and absolute \rp\ magnitude. We then fit the distribution of
-radial velocity jitter $\RVJitter$ as a mixture of two components: a normal distribution
-to represent single stars, and a log-normal distribution to represent stellar multiples.
-We perform these steps for $\approx10^4$ sources. While this process approximately maps
-how the radial velocity jitter of single (and multiple) stars varies across the Hertszprung-Russell
-diagram, each source of interest is considered independent from all others. For example,
-for three sources that are all indistinguishable from each other in \bp\ - \rp\ colour, 
-apparent \rp\ magnitude, and absolute \rp\ magnitude, the procedure we described above
-treats the fitting of these three mixture models as entirely independent of each other,
-even though we suspect that the parameters of the three mixture models ought to be 
-identical. In other words: the procedure as described does not enforce any smoothness.
-Given enough data this is not a problem for densely-populated regions of the 
-Hertszprung-Russell diagram as the optimized results \emph{will} be (statistically)
-indistinguishable for similar sources. In practice this poses a problem for poorly-sampled
-regions of the Hertzsprung-Russell diagram, and raises subjective questions on what
-constitutes \emph{the most} similar type of source.
+We adopt a two-step procedure for detecting stellar multiplicity. In the first step, for a
+given \Gaia\ source of interest, we consider the jitter in context of similar sources  in 
+order to measure the jitter of single stars and evaluate whether this source is more consistent with being a single star or stellar multiple. While we can repeat this process 
+for every source, this procedure is largely independent between two sources: there is no
+constraint requiring the estimated jitter to be smooth across source parameter space. At the second step we use a Gaussian process to model the jitter for single stars across
+source parameter space, and enforce smoothness. We illustrate our procedure in Figure \ref{fig:schematic} and describe it in more detail below.
 
-\todo{In the third step we apply a Gaussian process.}
+\subsection{Non-parametric model}
+
+When considering the (radial velocity or astrometric) jitter  for a \Gaia\ source of interest,
+we select $N$ sources in a `\emph{ball}' that have similar \bp\ - \rp\ colour, apparent \rp\ magnitude, and absolute \rp\ magnitudes as the source of interest.
+We chose $N = 1024$ in order to ensure that the sample is large enough such that
+it will include \emph{some} single stars and some stellar multiples, while remaining small enough
+such that all sources have very similar colours and magnitudes.
+There are other subtle constraints that we enforce on the sources in the `\emph{ball}'.
+Specifically we scale the relative dimensions such that 0.05\,mags in \bp\ - \rp\
+colour is equivalent to 0.5\,mags in absolute \rp\ magnitude, and 0.25\,mags in apparent magnitude.
+We further require that the ball has a minimum radius of 0.1\,mags in \bp\ - \rp\ colour and 0.5\,mags in apparent and absolute magnitude.
+In dense regions of parameter space there are often more than $N$ sources that are within a multi-dimensional ball of that radius.
+In those situations we randomly chose $N$ sources within the ball such that the randomly chosen sources approximately span the minimum radius constraint.
+
+Given that sample of $N$ similar sources, we fit a two-component mixture model to the
+distribution of jitter values. We treat the astrometric and radial velocity jitter separately.  This mixture model includes a normal (Gaussian) distribution to represent the intrinsic
+jitter for single stars, and a log-normal distribution to represent stellar multiples.
+The model parameters include the relative mixing weight $\theta$, the mean $\mu_\textrm{s}$
+and standard deviation $\sigma_\textrm{s}$ of the normal distribution, and the mean $\mu_\textrm{m}$ and standard deviation $\sigma_\textrm{m}$ of the log-normal distribution. We assume the priors
+\begin{eqnarray}
+\theta 				& \sim & \textrm{Uniform}(0, 1)    \\
+\mu_\textrm{s} 		& \sim & \textrm{Uniform}(0.5, 10) \\
+\sigma_\textrm{s} 	& \sim & \textrm{Uniform}(0, 10)   \\
+\mu_\textrm{m} 		& \sim & \textrm{Uniform}(\log[\mu_\textrm{s} + \sigma_\textrm{s}] + \sigma_\textrm{m}^2, \infty) \label{eq:mode_prior} \\
+\sigma_\textrm{m} 	& \sim & \textrm{Uniform}(0, 10)   \quad .
+\end{eqnarray}
+
+Where Eq. \ref{eq:mode_prior} ensures that the \emph{mode} of the log-normal
+distribution must be greater than $\mu_\textrm{s} + 1\sigma_\textrm{s}$, thereby
+requiring that stellar multiples must have higher jitter than single stars.
+
+%We illustrate our approach in Figure \ref{fig:npm_schematic}, which can be qualitatively
+%described by the following steps. When considering the radial velocity jitter for a
+%\Gaia\ source of interest, we select $N$ sources that have a similar \bp\ - \rp\ colour,
+%apparent \rp\ magnitude, and absolute \rp\ magnitude. We then fit the distribution of
+%radial velocity jitter $\RVJitter$ as a mixture of two components: a normal distribution
+%to represent single stars, and a log-normal distribution to represent stellar multiples.
+%We perform these steps for $\approx10^4$ sources. While this process approximately maps
+%how the radial velocity jitter of single (and multiple) stars varies across the Hertszprung-Russell
+%diagram, each source of interest is considered independent from all others. For example,
+%for three sources that are all indistinguishable from each other in \bp\ - \rp\ colour, 
+%apparent \rp\ magnitude, and absolute \rp\ magnitude, the procedure we described above
+%treats the fitting of these three mixture models as entirely independent of each other,
+%even though we suspect that the parameters of the three mixture models ought to be 
+%identical. In other words: the procedure as described does not enforce any smoothness.
+%Given enough data this is not a problem for densely-populated regions of the 
+%Hertszprung-Russell diagram as the optimized results \emph{will} be (statistically)
+%indistinguishable for similar sources. In practice this poses a problem for poorly-sampled
+%regions of the Hertzsprung-Russell diagram, and raises subjective questions on what
+%constitutes \emph{the most} similar type of source.
 
@@
 
-\subsection*{Step 1: Selecting similar sources}
+%\subsection*{Step 1: Selecting similar sources}
 
-We construct a $k$-dimensional tree \citep[k-d tree;][]{someone} for all sources in the
-dimensions of \bp\ - \rp\ colour, apparent \rp\ magnitude, and absolute \rp\ magnitude.
-When doing so we scale the dimensions such that \todo{0.1}\,mags in \bp\ - \rp\ colour is
-approximately equal to \todo{1} magnitude in apparent \rp\ magnitude and \todo{1} magnitude in 
-absolute \rp\ magnitude. We use this $k$-d tree for selecting a `ball' of similar
-stars for a given source.
+%We construct a $k$-dimensional tree \citep[k-d tree;][]{someone} for all sources in the
+%dimensions of \bp\ - \rp\ colour, apparent \rp\ magnitude, and absolute \rp\ magnitude.
+%When doing so we scale the dimensions such that \todo{0.1}\,mags in \bp\ - \rp\ colour is
+%approximately equal to \todo{1} magnitude in apparent \rp\ magnitude and \todo{1} magnitude in 
+%absolute \rp\ magnitude. We use this $k$-d tree for selecting a `ball' of similar
+%stars for a given source.
 
-We enforce a number of constraints when selecting similar stars. For each source
-we require the multi-dimensional `ball' to include at least \todo{128} sources.
-We further require that the radius of the ball extends to at least \todo{0.05}\,mags in 
-\bp\ - \rp\ colour, \todo{0.5}\,mags in apparent \rp\ magnitude, and \todo{0.5}\,mags in absolute
-\rp\ magnitude. If the initial query of \todo{128} points does not extend wide enough to
-meet our constraints on minimum radius, we extend the radius (linearly in each scaled)
-dimension until the minimum constraints on radii are met. Our minimum radii constraints
-imply that there will be many (e.g., $>>\todo{128}$) points returned in dense regions of the
-Hertszprung-Russell diagram. In these situations we then subsample the selected points
-to return a random \todo{1024} sources in any ball, while enforcing our constraints on
-minimum radii. As a result, the number of `similar' sources selected will range from
-a minimum of \todo{128} to a maximum of \todo{1024}, and the minimum radii constraints on various
-dimensions are always maintained.
+%We enforce a number of constraints when selecting similar stars. For each source
+%we require the multi-dimensional `ball' to include at least \todo{128} sources.
+%We further require that the radius of the ball extends to at least \todo{0.05}\,mags in 
+%\bp\ - \rp\ colour, \todo{0.5}\,mags in apparent \rp\ magnitude, and \todo{0.5}\,mags in absolute
+%\rp\ magnitude. If the initial query of \todo{128} points does not extend wide enough to
+%meet our constraints on minimum radius, we extend the radius (linearly in each scaled)
+%dimension until the minimum constraints on radii are met. Our minimum radii constraints
+%imply that there will be many (e.g., $>>\todo{128}$) points returned in dense regions of the
+%Hertszprung-Russell diagram. In these situations we then subsample the selected points
+%to return a random \todo{1024} sources in any ball, while enforcing our constraints on
+%minimum radii. As a result, the number of `similar' sources selected will range from
+%a minimum of \todo{128} to a maximum of \todo{1024}, and the minimum radii constraints on various
+%dimensions are always maintained.
 
@@
 
-\subsection*{Step 2: Fitting the two-component mixture}
-
-We fit a two-component mixture model to the distribution of radial velocity jitter for
-all similar sources selected through Step 1. This mixture model includes a normal
-distribution to represent the intrinsic jitter for single stars, and a log-normal
-distribution to represent stellar multiples. Specifically the model parameters include
-a mixing weight parameter $\theta$ that is constrained between $(0, 1)$, the mean 
-$\mu_{s}$ and standard deviation $\sigma_{s}$ of the normal distribution,
-and the mean $\mu_{m}$ and standard deviation $\sigma_{m}$ of the 
-log-normal distribution.
-
-We assume a beta prior on $\theta$  such that $\theta \sim \mathcal{B}\left(5, 5\right)$
-and require that $\sigma_{s}$ and $\sigma_{m}$ are larger than zero.
-Depending on the distribution of the radial velocity jitter, we often found this mixture
-challenging to optimize because the log-normal distribution could easily account for 
-most of the observed distribution, which tended the mixing weight to zero and implied
-that all selected sources were stellar multiples. To avoid this situation we placed a
-constraint on the mean of the log-normal distribution to ensure that the \emph{mode}
-of the log-normal distribution is larger than $\mu_{s} + 1\sigma_{s}$ such that
-\begin{eqnarray}
-%	\mu_{s} + 5\sigma_{s} & > & \exp\left(\mu_{m} - \sigma_{m}^2\right) \label{eq:rvm_lower_bound} \\
-	\mu_{s} + 1\sigma_{s} & < & \exp\left(\mu_{m} - \sigma_{m}^2\right) \label{eq:rvm_upper_bound} \quad .
-\end{eqnarray}
+%\subsection*{Step 2: Fitting the two-component mixture}
 
@@
 
+\begin{comment}
 Wherever possible we initialised the model parameters from nearby points (within the
@@
 \subsubsection*{Step 3: Evaluating probability and system properties}
+\end{comment}
 
-\todo{Bayes factors for each dimension?}
+We performed this procedure and optimized this mixture model for $K = 5000$ \todo{randomly}
+chosen sources, giving us noisy estimates of the astrometric and radial velocity jitter
+for sources across the Hertszprung-Russell diagram. In theory we could have repeated the
+procedure above for all sources, but we were computationally limited  by the number of
+sources that we could use in a Gaussian process.
 
-\todo{source properties: semi-amplitude (easy)}
-
-\todo{Can we recover period and eccentricity using The Joker?}
+\subsection{Gaussian processes}
 
@@
 %and bluer colours than what would be expected for a single star system.
-
-\subsection{SB2 Systems: Double-lined spectroscopic binaries}
-\label{sec:sb2_methods}
-
-Many sources in the second \Gaia\ data release do not have a reported radial 
-velocity, despite being bright enough ($G \lesssim 13$) and in a suitable 
-temperature range (e.g., between $\approx4000\,\textrm{K}$ and $\approx6500\,\textrm{K}$) 
-for radial velocities to be measured \citep{Cropper:2018}. The principle reason 
-why radial velocity measurements are \emph{not} reported for these stars is 
-because the \Gaia/DPAC team have identified the source to be a double-lined 
-spectroscopic binary (a so-called SB2-type system), either through a 
-composition of two stellar sources present in the spectra, or from multiple 
-(significant) modes in the cross-correlation function. In these situations it
-is not sensible to report a point estimate of the radial velocity of the point 
-source. For fainter or bluer sources, however, the radial velocity 
-may not be reported simply because it is too faint, blue, or red.
-
-We calculated the completeness of radial velocity measurements (e.g., the
-fraction of sources with reported radial velocities) as a function of all 
-available source properties that might affect whether the radial velocity may
-be reported or not. This included position ($\alpha$, $\delta$, $l$, $b$),
-parallax, proper motions, apparent magnitudes, \bp\ - \rp\ colour, 
-properties of the radial velocity templates, stellar parameters ($\teff$, $\radius$, $\luminosity$),
-and other properties. The full list of properties investigated is described in
-Appendix \ref{appendix:sb2}, with corresponding figures.
-
-In Figure \ref{fig:radial_velocity_completeness} we show the completeness as 
-a function of some pertinent properties, which demonstrate that the radial 
-velocity completeness is relatively flat until a source becomes too faint 
-(low \rp\ flux), or is either too blue or too red. We adopt conservative 
-limits for when the radial velocity completeness starts to drop with these 
-properties, and we assume that any source within the following range of 
-source parameters
-\begin{eqnarray}
-    \mathrm{phot\,\,rp\,\,mean\,\,flux} & \in & [10^{4}, 10^8] \nonumber \\
-    \mathrm{bp-rp} & \in & [X, Y] 
-    \label{eq:sb2_criteria}
-\end{eqnarray}
-\noindent{}is likely to be a double-lined spectroscopic binary (SB2) if no 
-radial velocity is reported. That is to say that we are explicitly assuming
-that if a point source meets the criteria in Equation \ref{eq:sb2_criteria} 
-and does not have a reported radial velocity measurement, then the point 
-source is an unresolved double-lined spectroscopic binary. In principle we 
-could make more realistic attempts to model the radial velocity completeness as
-a function of stellar properties rather than simply stating ``sources within
-this parameter range should have radial velocities unless they are SB2 systems'',
-but the radial velocity completeness within our specified range is reasonably
-flat.
-
-\todo{Some confirmation of this? The properties of SB2 candidate systems relative to others. Figure \ref{fig:sb2_histograms}.}
-
-We stress that our sample of SB2 type systems is calculated (or rather, deduced)
-separately to our calculations of multiplicity probability given the 
-\emph{measured} quantities from \Gaia\ like radial velocity, astrometry, and
-photometry. In Section \ref{sec:discussion} we detail some of the immediate
-limitations or interpretations of this assumption, and we emphasise that the 
-main conclusions of this work do not depend on SB2 systems. Nevertheless, we 
-\emph{strongly} 
-caution that our deductive inference of SB2 systems will be far more uncertain 
-than the binary probabilities we derive from other information. Our indicator 
-whether a star is likely a SB2 system likely suffers from more contamination
-and lower completeness compared to the probabilities we derive from other information
-(e.g., the radial velocity jitter, astrometric excess noise, and photometric 
-colours), and the contamination and completeness likely vary over some
-complex (unknown) function of parameter space.
-
-\begin{figure*}
-%	\includegraphics[width=1.0\textwidth]{../figures/sb2_rvs_completeness.pdf}
-	\includegraphics[width=1.0\textwidth]{../figures/todo.png}
-    \caption{Fraction of \Gaia\ sources with reported radial velocities
-		     as a function of source properties. Within the adopted source
-		     parameter ranges (gray; see Section \ref{sec:sb2_methods})
-		     the radial velocity completeness is approximately flat.
-		     We flag sources within this range that do not have reported 
-		     radial velocities to be candidate SB2 systems.}
-    \label{fig:sb2_rvs_completeness}
-\end{figure*}
 
@@
 \section{Radial velocity completeness as a function of \Gaia\ source properties}
-\todo{Figures}
+
+\subsection{SB2 Systems: Double-lined spectroscopic binaries}
+\label{sec:sb2_methods}
+
+Many sources in the second \Gaia\ data release do not have a reported radial 
+velocity, despite being bright enough ($G \lesssim 13$) and in a suitable 
+temperature range (e.g., between $\approx4000\,\textrm{K}$ and $\approx6500\,\textrm{K}$) 
+for radial velocities to be measured \citep{Cropper:2018}. The principle reason 
+why radial velocity measurements are \emph{not} reported for these stars is 
+because the \Gaia/DPAC team have identified the source to be a double-lined 
+spectroscopic binary (a so-called SB2-type system), either through a 
+composition of two stellar sources present in the spectra, or from multiple 
+(significant) modes in the cross-correlation function. In these situations it
+is not sensible to report a point estimate of the radial velocity of the point 
+source. For fainter or bluer sources, however, the radial velocity 
+may not be reported simply because it is too faint, blue, or red.
+
+We calculated the completeness of radial velocity measurements (e.g., the
+fraction of sources with reported radial velocities) as a function of all 
+available source properties that might affect whether the radial velocity may
+be reported or not. This included position ($\alpha$, $\delta$, $l$, $b$),
+parallax, proper motions, apparent magnitudes, \bp\ - \rp\ colour, 
+properties of the radial velocity templates, stellar parameters ($\teff$, $\radius$, $\luminosity$),
+and other properties. The full list of properties investigated is described in
+Appendix \ref{appendix:sb2}, with corresponding figures.
+
+In Figure \ref{fig:radial_velocity_completeness} we show the completeness as 
+a function of some pertinent properties, which demonstrate that the radial 
+velocity completeness is relatively flat until a source becomes too faint 
+(low \rp\ flux), or is either too blue or too red. We adopt conservative 
+limits for when the radial velocity completeness starts to drop with these 
+properties, and we assume that any source within the following range of 
+source parameters
+\begin{eqnarray}
+    \mathrm{phot\,\,rp\,\,mean\,\,flux} & \in & [10^{4}, 10^8] \nonumber \\
+    \mathrm{bp-rp} & \in & [X, Y] 
+    \label{eq:sb2_criteria}
+\end{eqnarray}
+\noindent{}is likely to be a double-lined spectroscopic binary (SB2) if no 
+radial velocity is reported. That is to say that we are explicitly assuming
+that if a point source meets the criteria in Equation \ref{eq:sb2_criteria} 
+and does not have a reported radial velocity measurement, then the point 
+source is an unresolved double-lined spectroscopic binary. In principle we 
+could make more realistic attempts to model the radial velocity completeness as
+a function of stellar properties rather than simply stating ``sources within
+this parameter range should have radial velocities unless they are SB2 systems'',
+but the radial velocity completeness within our specified range is reasonably
+flat.
+
+\todo{Some confirmation of this? The properties of SB2 candidate systems relative to others. Figure \ref{fig:sb2_histograms}.}
+
+We stress that our sample of SB2 type systems is calculated (or rather, deduced)
+separately to our calculations of multiplicity probability given the 
+\emph{measured} quantities from \Gaia\ like radial velocity, astrometry, and
+photometry. In Section \ref{sec:discussion} we detail some of the immediate
+limitations or interpretations of this assumption, and we emphasise that the 
+main conclusions of this work do not depend on SB2 systems. Nevertheless, we 
+\emph{strongly} 
+caution that our deductive inference of SB2 systems will be far more uncertain 
+than the binary probabilities we derive from other information. Our indicator 
+whether a star is likely a SB2 system likely suffers from more contamination
+and lower completeness compared to the probabilities we derive from other information
+(e.g., the radial velocity jitter, astrometric excess noise, and photometric 
+colours), and the contamination and completeness likely vary over some
+complex (unknown) function of parameter space.
+
+\begin{figure*}
+%	\includegraphics[width=1.0\textwidth]{../figures/sb2_rvs_completeness.pdf}
+	\includegraphics[width=1.0\textwidth]{../figures/todo.png}
+    \caption{Fraction of \Gaia\ sources with reported radial velocities
+		     as a function of source properties. Within the adopted source
+		     parameter ranges (gray; see Section \ref{sec:sb2_methods})
+		     the radial velocity completeness is approximately flat.
+		     We flag sources within this range that do not have reported 
+		     radial velocities to be candidate SB2 systems.}
+    \label{fig:sb2_rvs_completeness}
+\end{figure*}
 

### 2018-08-29 article/ms.tex
words

@@
 
-\section{Method} \label{sec:method}
+\section{Methods} \label{sec:method}
 
@@
 the jitter in context with other sources that have a similar properties (\bp\ - \rp\ colour,
-apparent \rp\ magnitude, and absolute \rp\ magnitude).\footnote{The magnitude in any band would be sufficient. Throughout
-this work we chose the \rp\ band because its wavelength range most closely overlaps 
-with that of the Radial Velocity Spectrograph on board \Gaia.} However, we cannot reliably describe a quantitative
+apparent magnitude, and absolute magnitude).\footnote{The magnitude in any band would be sufficient.
+Throughout this work we chose the \rp\ band for radial velocity data because the wavelength of the \rp\ band most closely overlaps with that of the Radial Velocity Spectrograph on board \Gaia.
+For similar reasons we use the $G$ band magnitudes for astrometry data.}
+However, we cannot reliably describe a quantitative
 relationship for \emph{how} exactly the jitter of a single star is
@@
 
-\subsection{Non-parametric model}
+\subsection{Two component mixture model}
 
@@
 
-We performed this procedure and optimized this mixture model for $K = 5000$ \todo{randomly}
-chosen sources, giving us noisy estimates of the astrometric and radial velocity jitter
-for sources across the Hertszprung-Russell diagram. In theory we could have repeated the
-procedure above for all sources, but we were computationally limited  by the number of
-sources that we could use in a Gaussian process.
+We \todo{randomly} selected $K = 5000$ \Gaia\ sources of interest and performed the procedure
+described above, where we select a ball of similar sources, and then fit the jitter 
+distributions with a two-component mixture model. We numerically optimised this model 
+using the \texttt{L-BFGS-B} optimisation algorithm \citep{lbfgsb}. In theory we could repeat this procedure
+for all \Gaia\ sources as the problem is embarrassingly parallel. In practice we are
+computationally limited by the number of sources that we can include in our Gaussian process
+model.
 
-\subsection{Gaussian processes}
+\subsection{Gaussian process model}
+
+The procedure of selecting comparison sources and optimising a two-component mixture model
+is largely independent between similar sources. That is to say that even if two sources have very
+similar properties, there is no constraint stating that the estimated single-star jitter
+between those two sources should be similar. There is similarly no constraint that they 
+should vary smoothly with respect to the source properties, even though we expect that they
+ought to vary smoothly, albeit by some complex relationship that is non-trivial to describe.
+For these reasons we chose to use a Gaussian process model to parameterise the covariance
+between data points, where here we will treat our noisy estimates of the jitter parameters
+$\{\mu_\textrm{s},\sigma_\textrm{s},\mu_\textrm{m},\sigma_\textrm{m}\}$ as the data.
+
+Gaussian processes are specified by the use of kernels \citep[see][for an thorough introduction]{Rasmussen:2005}. These kernel functions model
+the covariance structure between data points (e.g., $\mu_\textrm{s}$) instead of directly
+modelling the data, providing a flexible generative model for data typically with only very
+few hyperparameters. A detailed discussion of kernel choice is beyond the scope of this work \citep[see Chapter 4 of][]{Rasmussen:2005}.
+We chose to use the sum of a squared-exponential kernel function \citep{someone} and a Mat\'ern-3/2
+function \citep{someone}, with a white noise component $\sigma$ such that the combined kernel is
+\begin{equation}
+k(r) = k_1(r) + k_2(r)
+\end{equation}
+\noindent{}where
+\begin{eqnarray}
+k_1(r^2) & = & \exp\left(-\sqrt{r^2}\right) \\
+k_2(r^2) & = & \left(1 + \sqrt{3r^2}\right)\exp\left(-\sqrt{3r^2}\right) \quad .
+\end{eqnarray}
+
+We allow a different scale length in each dimension, and there are three dimensions of
+$\textbf{x}$ (\bp\ - \rp\ colour, absolute magnitude and apparent magnitude),
+leading to 6 hyperparameters to describe the kernel functions. The addition of a white
+noise component and a mean model brings the total number of unknown hyperparameters to 8
+for each Gaussian process model. We construct a different Gaussian process for each 
+jitter parameter (e.g., $\mu_\textrm{s}$, $\sigma_\textrm{s}$).
+
+We constructed this Gaussian process model and optimized the hyperparameters using
+\texttt{george}  \citep{george, others}. With these models we then made posterior
+predictions of the jitter properties for all \todo{X} \Gaia\ sources in our subset.
+With these predictions we evaluated the log-likelihood of each \Gaia\ source being
+drawn from a single star distribution $\mathcal{L}_\textrm{s}$ and a stellar
+multiple $\mathcal{L}_\textrm{m}$
+
+\todo{log-L functions}
+
+\noindent{}and evaluated the probability of each \Gaia\ source being a single star,
+given the radial velocity jitter or astrometric jitter:
+
+\todo{probabilityy functions}
+
+We assume the astrometric observations are independent from the radial velocity
+observations, and combine this information to evaluate the probability of a
+\Gaia\ source being a single star
+\begin{equation}
+	\tau_\textrm{single} = \todo{etc}
+\end{equation}
+\noindent{}where a $\tau_\textrm{single}$ value of 1 indicates the source is
+fully consistent with being a single star, and a value of 0 indicates the
+source is a stellar multiple (e.g., a binary or trinary system). We provide
+estimates of $\tau_\textrm{rv}$ just based on radial velocity data,
+$\tau_\textrm{ast}$ based on only astrometric data, and $\tau_\textrm{single}$
+based on all radial velocity and astrometric data.
+
+\todo{estimates of uncertainty on $\tau$, and some brief cautionary  notes on using  $\tau$, while we have the readers' attention}
+
+\subsection{Partial characterisation of stellar multiples}
+
+$\tau_\textrm{single}$ provides an estimate on whether a source is a single star. 
+If a source is not consistent with being a single star (i.e., $\tau_\textrm{single} \approx 0$),
+then we do not attempt to make any inferences as to whether the source is a binary
+system, a trinary, or higher stellar multiple: we simply assume that the source must
+be a binary system. This -- and other assumptions -- have implications for our
+partial characterisation of binary systems.
+
+We adopt the following conservative criteria for establishing a source as being a
+binary system:
+\begin{eqnarray}
+\todo{\textrm{sum hard cuts}} \quad .
+\end{eqnarray}
+
+There are \todo{X} sources in our subset that meet this criteria (\todo{X\%} of \todo{Y} sources).
+
+\subsubsection{The radial velocity semi-amplitude, $K_1$}
+
+For these systems with radial velocity measurements we estimate the radial velocity semi-amplitude
+$K_1$ of the primary from the radial velocity jitter that is in excess of what we would expect for
+a single star with the same properties
+\begin{equation}
+	K_1 \approx \RVJitter - \mathcal{GP}_{\mu,s} \quad .
+\end{equation}
+\todo{best way to express this}
+
+\noindent{}and estimate the error in $K_1$ by propagating all sources of parameter uncertainty
+\todo{from the GP and the sigma etc,...}.
+
+\subsubsection{Orbital period $P$}
+
+A comparison with \cite{SB9} reveals that we reliably detect RV binaries with periods between 
+X and Y days, consistent with our expectations: the second \Gaia\ data release includes observations
+spanning Z months (half of Y), and only half of the orbital period is required to characterise the
+radial-velocity semi-amplitude of the system. However, for systems with longer periods that we
+identify as a binary, our estimate of the radial velocity semi-amplitude will be underestimated 
+approximately by the fraction of the observing span relative to the true orbital period.
+These are coarse approximations, because most \Gaia\ sources are only observed about once a month,
+but they  give qualitative intuition for what is the shortest and longest orbital period that
+is detectable by RV variations.
+
+\subsubsection{Eccentricity $e$}
+
+\todo{Periodicity, eccentricity from The Joker}
+
+\subsubsection{Inclination angle $i$}
+
+- Estimate astrometric excess for systems that are `clearly' astrometric  binaries
+
+%- \todo{is the astrometric \textbf{excess} comparable between photometric magnitudes -- eg can that be used to translate to tangential velocities}
+% OMG YES IT IS
+
+- Eclipsing binaries to translate the excess astrometric jitter to tangential velocities
+
+- Inclination angles for some subset
+
+%- Photometric variability of sources that are astrometric binaries and appear to be eclipsing (e.g., should have RVs but don't)?
+Or just EBs compared to a sample of single stars with similar properties?
+Or just EBs?
+Or nothing at all?
 
@@
 \section{Results} \label{sec:results}
+
+\todo{We confirm that most stars are in binaries.}
+
+- First convince the reader that we have measured something
+
+- Comparisons: SB9, APW, et cetera
+
+- Binary fraction as a  function  of everything
+
+- Number of binaries detected as a function of things
 

### 2018-08-29 article/ms.tex
minor

@@
 
-\begin{figure*}
-%	\includegraphics[width=1.0\textwidth]{../figures/sb2_sky_structure.pdf}
-	\includegraphics[width=1.0\textwidth]{../figures/todo.png}
-
-    \caption{Fraction of sources (per sky bin) without reported radial velocities
-    		 for all sources with $G \lesssim 13$ (top) and sources in the ranges
-		 	 specified by Eqs \ref{eq:sb2_ranges} (bottom), where we assert that a missing
-			 radial velocity measurement signifies a likely SB2 candidate . The color scale is 
-			 arbitrarily set to highlight structure, where black indicates a higher
-			 fraction of sources do not have reported radial velocities. Crowding
-			 in the galactic plane likely results in some sources not having radial
-			 velocities reported. The large scale structure visible in both axes is
-			 a combined effect of the initial \Gaia\ source list, the scanning law,
-			 and star forming regions (i.e., where emission in the \ion{Ca}{2} 
-			 triplet likely causes issues for radial velocity determination).}
-    \label{fig:sb2_sky_structure}
-\end{figure*}
-
 % Figure: cross-match against a catalog of SB2 systems,... do we find all of
@@
 \section{Results} \label{sec:results}
+
+\begin{figure}
+	\includegraphics[width=0.5\textwidth]{figures/rv_sb9_k_comparison_v5.png}
+    \caption{}
+    \label{fig:rv_sb9_k_comparison}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{figures/ast_gp_hrd_v5.png}
+\caption{}
+\label{fig:ast_gp_hrd_v5}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{figures/hrd_single_star_fraction_hist.png}
+\caption{}
+\label{fig:hrd_single_star_fraction_hist}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{figures/rv_apw_unimodal_k_comparison_v5.png}
+\caption{}
+\label{fig:rv_apw_unimodal_k_comparison_v5}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{figures/rv_apw_unimodal_kalpha_comparison_v5.png}
+\caption{}
+\label{fig:rv_apw_unimodal_kalpha_comparison_v5}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{figures/rv_hrd_single_star_fraction_hist_v5.png}
+\caption{}
+\label{fig:rv_hrd_single_star_fraction_hist_v5}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{figures/rv_sb9_k_comparison_v5.png}
+\caption{}
+\label{fig:rv_sb9_k_comparison_v5}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{figures/rv_sb9_kalpha_comparison_period.png}
+\caption{}
+\label{fig:rv_sb9_kalpha_comparison_period}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{figures/rv_sb9_kalpha_comparison_tau.png}
+\caption{}
+\label{fig:rv_sb9_kalpha_comparison_tau}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{figures/rv_soubiran_hist_v5.png}
+\caption{}
+\label{fig:rv_soubiran_hist_v5}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{figures/tau_single_wrt_kdt_labels_v5.png}
+\caption{}
+\label{fig:tau_single_wrt_kdt_labels_v5}
+\end{figure}
 
@@
 
+\begin{figure*}
+%	\includegraphics[width=1.0\textwidth]{../figures/sb2_sky_structure.pdf}
+	\includegraphics[width=1.0\textwidth]{../figures/todo.png}
+
+    \caption{Fraction of sources (per sky bin) without reported radial velocities
+    		 for all sources with $G \lesssim 13$ (top) and sources in the ranges
+		 	 specified by Eqs \ref{eq:sb2_ranges} (bottom), where we assert that a missing
+			 radial velocity measurement signifies a likely SB2 candidate . The color scale is 
+			 arbitrarily set to highlight structure, where black indicates a higher
+			 fraction of sources do not have reported radial velocities. Crowding
+			 in the galactic plane likely results in some sources not having radial
+			 velocities reported. The large scale structure visible in both axes is
+			 a combined effect of the initial \Gaia\ source list, the scanning law,
+			 and star forming regions (i.e., where emission in the \ion{Ca}{2} 
+			 triplet likely causes issues for radial velocity determination).}
+    \label{fig:sb2_sky_structure}
+\end{figure*}
+
 % astroquery

### 2018-08-29 article/figures/plot_rv_figures.py
thing

@@ -3,35 +3,49 @@ import os
 import pickle
 from astropy.io import fits
 from astropy.table import Table
+from scipy.stats import binned_statistic
 
 import matplotlib
 import matplotlib.pyplot as plt
 from matplotlib.ticker import MaxNLocator
 from mpl_toolkits.axes_grid1 import make_axes_locatable
 
-from mpl_utils import (mpl_style, plot_binned_statistic)
+from mpl_utils import (mpl_style, plot_binned_statistic, plot_histogram_steps)
 
 plt.style.use(mpl_style)
 
 
 BASE_PATH = "../../"
-RELEASE_CANDIDATE_VERSION = 0
+RELEASE_CANDIDATE_VERSION = 5
 
 MAKE_FIGURES = [
     "rv_hrd_single_star_fraction_hist",
     "rv_hrd_single_star_fraction_scatter",
-    "rv_sb9_kalpha_comparison",
-    "rv_sb9_k_comparison",
-    "rv_sb9_kalpha_comparison_period",
-    "rv_sb9_kp_corner",
     "rv_apw_unimodal_k_comparison",
     "rv_apw_unimodal_kalpha_comparison",
     "rv_apw_percentiles_k_comparison",
     "rv_gp_hrd",
     "rv_gp_wrt_params",
     "rv_soubiran_hist",
-    "rv_huang_hist"
+    "rv_huang_hist",
+    "rv_mean_tau_single_wrt_kdt_labels",
+    "ast_mean_tau_single_wrt_kdt_labels",
+    "mean_tau_single_wrt_kdt_labels",
+    "all_tau_single_wrt_kdt_labels",
+    "rv_sb9_k_comparison",
+    "hrd_single_star_fraction_hist",
+    "tau_single_wrt_kdt_labels",
+    "rv_huang_hist2",
+
 ]
+
+#MAKE_FIGURES = [
+#    "ast_gp_hrd"
+#]
+MAKE_FIGURES = [
+    "tau_single_hrd"
+]
+
 
 DEFAULT_SEQUENTIAL_CMAP = "viridis"
 
@@ -39,17 +53,24 @@ DEFAULT_SEQUENTIAL_CMAP = "viridis"
 velociraptor = fits.open(os.path.join(
     BASE_PATH,
     "results", 
-    "velociraptor-catalog-rc.{:.0f}.fits".format(RELEASE_CANDIDATE_VERSION)))
[…]

### 2018-08-30 article/figures/mpl_utils.py
ready for v7

@@ -1,0 +1,157 @@
+
+# A matplotlib style based on the gala package by @adrn:
+# github.com/adrn/gala
+
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.stats import binned_statistic_2d
+
+mpl_style = {
+
+    # Lines
+    'lines.linewidth': 1.7,
+    'lines.antialiased': True,
+    'lines.marker': '.',
+    'lines.markersize': 5.,
+
+    # Patches
+    'patch.linewidth': 1.0,
+    'patch.facecolor': '#348ABD',
+    'patch.edgecolor': '#CCCCCC',
+    'patch.antialiased': True,
+
+    # images
+    'image.origin': 'upper',
+
+    # colormap
+    'image.cmap': 'viridis',
+
+    # Font
+    'font.size': 12.0,
+    'text.usetex': True,
+    'text.latex.preamble': r'\usepackage{amsmath}',
+    'text.latex.preview': True,
+    'axes.unicode_minus': False,
+
+    # Axes
+    'axes.facecolor': '#FFFFFF',
+    'axes.edgecolor': '#333333',
+    'axes.linewidth': 1.0,
+    'axes.grid': False,
+    'axes.titlesize': 'x-large',
+    'axes.labelsize': 'large',
+    'axes.labelcolor': 'k',
+    'axes.axisbelow': True,
+
+    # Ticks
+    'xtick.major.size': 8,
+    'xtick.minor.size': 4,
+    'xtick.major.pad': 6,
+    'xtick.minor.pad': 6,
+    'xtick.color': '#333333',
+    'xtick.direction': 'in',
+    'ytick.major.size': 8,
+    'ytick.minor.size': 4,
+    'ytick.major.pad': 6,
+    'ytick.minor.pad': 6,
+    'ytick.color': '#333333',
+    'ytick.direction': 'in',
+    'xtick.labelsize': 'medium',
+    'ytick.labelsize': 'medium',
+
+    # Legend
+    'legend.fancybox': True,
+    'legend.loc': 'best',
+
+    # Figure
+    'figure.figsize': [6, 6],
+    'figure.facecolor': '1.0',
+    'figure.edgecolor': '0.50',
+    'figure.subplot.hspace': 0.5,
+
+    # Other
+    'savefig.dpi': 300,
+}
+
+
+
+
+def plot_histogram_steps(ax, x_bins, y, y_err, **kwargs):
+
[…]

### 2018-09-03 article/ms.tex
changes

@@
 
-\title{Detection and partial characterisation of stellar multiplicity with Gaia}
+\title{Detection and partial characterisation of stellar multiplicity}
 
@@
 \author{Rosemary Mardling}
+\affiliation{School of Physics \& Astronomy, 
+			 Monash University,
+			 Wellington Rd, Clayton 3800, Victoria, Australia}
+
 \author{John Lattanzio}
+\affiliation{School of Physics \& Astronomy, 
+			 Monash University,
+			 Wellington Rd, Clayton 3800, Victoria, Australia}
 
-\author[0000-0003-2866-9403]{David W. Hogg}
-\affiliation{Center for Cosmology and Particle Physics, Department of Physics,
-    		 New York University, 
-		  	 726 Broadway, New York, NY 10003, USA}
-\affiliation{Center for Data Science, 
-			 New York University, 60 Fifth Ave, New York, NY 10011, USA}
-\affiliation{Max-Planck-Institut f\"ur Astronomie, 
-			 K\"onigstuhl 17, D-69117 Heidelberg}
-\affiliation{Flatiron Institute,
-			 162 Fifth Ave, New York, NY 10010, USA}
+%\author[0000-0003-2866-9403]{David W. Hogg}
+%\affiliation{Center for Cosmology and Particle Physics, Department of Physics,
+%    		 New York University, 
+%		  	 726 Broadway, New York, NY 10003, USA}
+%\affiliation{Center for Data Science, 
+%			 New York University, 60 Fifth Ave, New York, NY 10011, USA}
+%\affiliation{Max-Planck-Institut f\"ur Astronomie, 
+%			 K\"onigstuhl 17, D-69117 Heidelberg}
+%\affiliation{Flatiron Institute,
+%			 162 Fifth Ave, New York, NY 10010, USA}
 
-\author[0000-0003-0872-7098]{Adrian M. Price-Whelan}
-\affiliation{Department of Astrophysical Sciences, 
-			 Princeton University, 
-			 Princeton, NJ 08544, USA}
+%\author[0000-0003-0872-7098]{Adrian M. Price-Whelan}
+%\affiliation{Department of Astrophysical Sciences, 
+%			 Princeton University, 
+%			 Princeton, NJ 08544, USA}
 
@@
 likely explanations are that either the \Gaia\ team identified the source to be a
-double-lined spectroscopic binary, or the radial velocity error was larger than 20\,km\,s$^{-1}$,
+double-lined spectroscopic binary, or the radial velocity error $\sigma_{\textrm{V}_\textrm{R}}^\textrm{MTA}$ exceeded 20\,km\,s$^{-1}$ (irrespective of the number of transits),
 and so the source radial velocity was removed. While it may be appealing to use the \emph{lack}
-of radial velocity measurement as an indicator of stellar multiplicity, we show in Appendix A
+of radial velocity measurement as an indicator of stellar multiplicity, we show in Appendix \ref{sec:appendix-missing-rvs}
 that such an approach is problematic due to completeness, and other complications arising from
-the initial \Gaia\ source catalog. However, in the following sections we show that for 
-sources that do not have reported radial velocities -- and it has an apparent magnitude 
-and colour that suggests it ought to have a radial velocity reported -- the astrometric 
-jitter is often sufficient to identify stellar multiples.
-
-% todo:  Figure showing distribution of astrometric unit weight error.
+the initial \Gaia\ source catalog. However, in later sections we also show that for 
+sources without reported radial velocities -- and those sources have apparent magnitudes
+and colours suggesting it ought to have a radial velocity measurement -- the astrometric 
+jitter is often sufficient to identify stellar multiples (see also Figure \ref{fig:sb2_histograms}).
 
@@
 We restrict our work to the brightest $\sim{}10^7$ sources in \Gaia\ DR2.
-About 70\% of sources in this subset have reported radial velocities, and
-this subset is relatively close to the Sun. Specifically, we executed the
+Specifically, we executed the
 following Astronomical Data Query Language \citep[\texttt{ADQL};][]{someone} 
@@
 \end{minted}
-\noindent{}This query returned \NumberOfSourcesInSubset\ sources.
+\noindent{}This query returned \NumberOfSourcesInSubset\ sources, of which
+about 70\% have radial velocities in \Gaia\ DR2.
 
@@
 
-We assume that for a sample of sources with similar properties:
-\begin{itemize}
-	\item \bp\ - \rp\ colours;
-	\item apparent \rp\ magnitude; and
-	\item absolute \rp\ magnitude,
-\end{itemize}
-\noindent{}the distribution of jitter (in astrometry or radial velocity)
+We assume that for a sample of sources with a similar \bp\ - \rp\ colour,
+apparent magnitude\footnote{The magnitude in any band is sufficient.
+Throughout this work we chose the \rp\ band for radial velocity data because the wavelength of the \rp\ band most closely overlaps with that of the Radial Velocity Spectrograph on board \Gaia.
+For similar reasons we use the G-band magnitudes for astrometry data.}, and absolute magnitude,
+the distribution of jitter (in astrometry or radial velocity)
 is a mixture of two components:
@@
 the jitter in context with other sources that have a similar properties (\bp\ - \rp\ colour,
-apparent magnitude, and absolute magnitude).\footnote{The magnitude in any band would be sufficient.
-Throughout this work we chose the \rp\ band for radial velocity data because the wavelength of the \rp\ band most closely overlaps with that of the Radial Velocity Spectrograph on board \Gaia.
-For similar reasons we use the $G$ band magnitudes for astrometry data.}
+apparent magnitude, and absolute magnitude).
+
 However, we cannot reliably describe a quantitative
@@
 
+\begin{figure*}
+	\includegraphics[width=\textwidth]{../figures/todo.png}
+%	\includegraphics[width=\textwidth]{figures/tau_single_hrd.png}
+    \caption{Two-dimensional histogram of \todo{X} showing the mean single star probability $\tau$ per bin.
+The three panels show probabilities calculated from different sources of information: (a) using only radial velocity jitter; (b) using only astrometric jitter; and (c) using the joint information of radial velocity and astrometric jitter.
+All panels share the same colour range.}
+    \label{fig:tau_single_hrd_v5}
+\end{figure*}
+
 %For higher-order star systems with longer periods, we can still find that a
@@
 \begin{figure}
-%	\includegraphics[width=0.4\textwidth]{../figures/sb2_histograms.pdf}
-	\includegraphics[width=0.4\textwidth]{../figures/todo.png}
+	\centering
+	\includegraphics[width=0.35\textwidth]{../figures/sb2_histograms.pdf}
+%	\includegraphics[width=0.4\textwidth]{../figures/todo.png}
     \caption{Distribution of 
@@
 				(c) photometric \rp\ variability (see Eq. \ref{eq:photometric_variability})
-			of candidate SB2 type systems relative to a control sample of the same size.}
+			of sources that lack radial velocities but their properties suggest they should have
+			a radial velocity measurement (red; candidate SB2 systems), and an equal-size control sample of sources
+			with similar \bp\ - \rp\ colours, apparent magnitudes, absolute magnitudes,
+			which \emph{do} have radial velocity measurements (grey). It is clear that the
+			candidate SB2 systems have systematically higher astrometric unit weight error
+			than the control sample, and a larger variance in photometric \bp\ - \rp\ excess
+			factor and photometric variability, both consistent with a secondary source.}
     \label{fig:sb2_histograms}
@@
 %\subsection{Photometric binaries} \label{sec:method_photometric_binaries}
-
-%\begin{eqnarray}
-%	\textrm{Amp} = \log_{10}\left(\sqrt{N_{obs,al}}\frac{\sigma_{\langle{}f_G\rangle}}{\langle{}f_G\rangle}\right) \label{eq:photometric_variability}
-%\end{eqnarray}
 
@@
 
-\section{Radial velocity completeness as a function of \Gaia\ source properties}
+\section{Radial velocity completeness as a function of \Gaia\ source properties} \label{sec:appendix-missing-rvs}
 
@@
 
+\todo{\begin{eqnarray}
+	\textrm{Amp} = \log_{10}\left(\sqrt{N_{obs,al}}\frac{\sigma_{\langle{}f_G\rangle}}{\langle{}f_G\rangle}\right) \label{eq:photometric_variability}
+\end{eqnarray}}
+
 \begin{figure*}
